### 2016-12-05 manuscript.Rmd
initial commit

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

author_note: >
  Complete departmental affiliations for each author (note the indentation, if you start a new paragraph).

  Enter author note here.

abstract: >
  Enter abstract here (note the indentation, if you start a new paragraph).

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_word
---

```{r message = FALSE, warning = FALSE}
library("papaja")
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

### 2016-12-05 manuscript.Rmd
added APA manuscript

@@
 ---
-title             : "The title"
-shorttitle        : "Title"
+title             : "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
+shorttitle        : "A Computational Perspective"
 
 author: 
-  - name          : "First Author"
+  - name          : "Andrew W. Ellis"
     affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
-    email         : "my@email.com"
-  - name          : "Ernst-August Doelle"
-    affiliation   : "1,2"
+    address       : "Fabrikstrasse 8, 3012 Bern, Switzerland"
+    email         : "andrew.ellis@psy.unibe.ch"
+  - name          : "Fred W. Mast"
+    affiliation   : "1"
 
@@
   - id            : "1"
-    institution   : "Wilhelm-Wundt-University"
-  - id            : "2"
-    institution   : "Konstanz Business School"
+    institution   : "Department of Psychology, University of Bern"
 
 author_note: >
-  Complete departmental affiliations for each author (note the indentation, if you start a new paragraph).
+  Andrew W. Ellis, Fred W. Mast, Department of Psychology.
 
-  Enter author note here.
+  AE and FM planned and conducted the research; AE and FM analyzed the data and wrote the manuscript.
+
+  Raw data and analysis scripts are available at https://osf.io/u26mq.
 
 abstract: >
-  Enter abstract here (note the indentation, if you start a new paragraph).
+    Recent research showed the involvement of vestibular information in cognitive processes.
+We provide a computational framework for vestibular cognition and thereby focus on perspective transformations, which require a simulated rotation of the self.
+We explore how the ability to model the effects of body rotation using vestibular input might enable the ability to imagine body rotations.
+We approach this problem by considering a simple agent endowed with a complex statistical model of the hidden causes of its sensory input.
+We explain how this statistical model enables the agent to predict its sensory input and hidden sensory variables.
+Based on a probabilistic graphical model, we formalize qualitative concepts of mental simulation.
+Predicting the sensory input is key to understanding imagined rotations.
+Mental imagery goes beyond prediction, rather, it involves the use of counterfactual queries in a probabilistic model.
 
-keywords          : "keywords"
+keywords          : Probabilistic graphical models, Particle filter, Embodied cognition, Vestibular cognition, Spatial perspective taking, Mental imagery
 wordcount         : "X"
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["references.bib"]
 
-figsintext        : no
+figsintext        : yes
 figurelist        : no
@@
 class             : "man"
-output            : papaja::apa6_word
+# output            : papaja::apa6_word
+output            : papaja::apa6_pdf
 ---
@@
 ```
-
-# Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-## Participants
-
-## Material
-
-## Procedure
-
-## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
 

### 2016-12-08 manuscript.Rmd
added loads of new stuff

@@
 title             : "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
+# title             : "Probabilistic Computations Underlying Simulated Self-Rotations"
 shorttitle        : "A Computational Perspective"
@@
 
-```{r message = FALSE, warning = FALSE}
+```{r preliminary, message = FALSE, warning = FALSE}
 library("papaja")
+library(pander)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(mvtnorm)
+library(ggthemes)
+
+source('~/thesis/papers/pgm-spt/R/utils.R')
+source('~/thesis/papers/pgm-spt/R/2d-particle-filter.R')
+source('~/thesis/papers/pgm-spt/simulations/rbiips-model.R')
+
+knitr::opts_chunk$set(cache = TRUE,
+                      fig.width = 12, fig.height = 8,
+                      dpi = 600, fig.path = "../generated-figures/",
+                      dev=c('pdf', 'png', 'cairo_ps'))
 ```
 
-# Results
+# Preliminary
+
+generate some data:
+
+```{r generate-data-1, echo=FALSE, message=FALSE, warning=FALSE}
+dt <- 0.1
+motiondata <- generate_data(T = 2, amplitude = 20, sensor_sd = 2.0)
+plot_trajectories(motiondata = motiondata)
+```
+
+# Introduction
+
+## Notes
+We need to simulate the behaviour of a complex dynamical system.
+
+Few (if any) attempts have been made to connect higher-level vestibular cognition to the lower-level computations performed in the service of sensory inference.
+THis is rather surprising, since the vestibular system has some very interesting properties with direct relevance for cognitive systems, and it is comparatively well understood in terms of its dynamics and the sensory afferent signals
+
+# Simulating a dynamical system
+A dynamical system can be described by a set of differential equations; these differential equation relate the first order derivatives of the state to the state variables.
+
+Actually, the brain must solve __stochastic differential equations__.
+
+Write equations for position, velocity and acceleration. __Important:__ $\omega$, the rate of change (of angular position $\theta$), depends on the current position $\theta$.
+
+These are the _forward kinematic equations_ (that take us from inputs (acceleration) to velocity and position). Given the kinematic equations, we can simulate entire trajectories through space.
+
+Why simulate? Movement through 3 dimensional space is complex (non-linear). It is 'easier' to simulate this kind of system using knowledge that is already available.
+
+Simulatinf a movement through space is akin to planning a movement before it is carried out. This is a very challenging problem
+
+> Euler’s method is probably the most simple and popular method for solving initial value problems for ODEs.
+
+# Vestibular Cognition
+Two tasks thought to constitute vestibular cognition are : SPT and imagined self-motion.
+We show that both processes are thought to involve a simulated rotation about the body's vertical (yaw) axis.
+Describe both processes and show figure.
+Both can be interpreted as mentally simulated rotations.
+We refer to both as simulation: in this case we show that the simulation consists of simulating a dynamical system using a generative model.
+Simulation can also be thought of as a kind of internal 'replay' [@Takac:2012jc] or working memory representation
+
+![SPT and imagined self-motion involve common computations](diagrams/fig-1-common-process.pdf)
+But what are these computations? The brain could either solve the equation for the final position, or could run foward simulation of the dynamic system.
+
+## Spatial Perspective Taking
+
+cite Grabherr vestibular patients paper
+
+## Imagined Self-Motion
+
+We investigate one specific type of head movement, a rotation of the head about the yaw axis, and discuss this in the context of the probabilistic computations that are performed when the goal is to:
+
+1) infer motion caused by external events
+2) infer the consequences of one's own actions.
+
+## Common Computations
+
+Head direction cells are thought to integrate signals of vestibular origin to maintain a signal of cumulative rotation (reviewed in @Taube:2007fk and are found in many brain areas, including the postsubiculum, retrosplenial cortex, thalamus, lateral mammillary nucleus, dorsal tegmental nucleus, striatum and entorhinal cortex [@Cullen:2014gx].
+This may be necessary in order to mentally simulate being in another position, i.e.
+the brain apparently needs to simulate the n-step dynamics of its motion through space.
+Why is this?
+Because vestibular connections make a major contribution to spatial orientation; information about the position in space obtained through the vestibular sensory system is computed by integrating velocity over time.
+
+The computations performed for SPT require a generative model. Further, we can identify two main topics:
+
+1) the probabilistic computations performed during sensory inference
+2) the computations carried out by the brain in order to distinguish between sensory consequences of voluntary actions and externally produced actions.
+The cerebellum is thought to be the site of a forward model that predicts the sensory consequences of self-generated action [@Brooks:2013cf], which is then 'compared' to the actual sensory stimulation in order to compute a prediction error.
+
+> The brain has a generative model for simulation. Where does this model come from? We show that this is precisely what the brain uses in order to perform sensory inference.
+
+> Tasks performed by a moving agent: sense our movements in space.
+This is made challenging by the fact the we move, and our sensory data may be due to our own movements, or externally applied forces.
+THese need to be distinguished, as they have to be handled differently.
+Exteranlly applied forces need to be compensated for, whereas this is often not the case for volitional movements.
+
+# Vestibular System
+
+# Dynamic Probabilistic Model
+The generative model consists of several components (Process model, observation model), some of which (process model) are re-used for simulation, while the observation model is not needed for simulation.
+It might however, provide a useful function (play a role) in simulation, by indicating whether the data are fit well or not (usually they are not).
+
+# Inference
+
+this is a reference to \autoref{figurelabel}
+
+![This is the caption \label{figurelabel}](diagrams/figure-particle-filter-explained.pdf)
+# Model description
+
+# Simulations
+
+We simulate various scenarios of model mis-spefication
+
+```{r}
+library(tables, quietly = TRUE)
+tab <- tabular( (Species + 1) ~ (n=1) + Format(digits=2)*
+         (Sepal.Length + Sepal.Width)*(mean + sd), data=iris )
+pander(tab)
+```
 
 # Discussion
+
+A simple model for perceptual decision-making: evidence accumulation via log-likelihood difference test.
 
@@
 ```{r create_r-references}
-r_refs(file = "r-references.bib")
+# r_refs(file = "r-references.bib")
 ```

### 2016-12-08 simulations/sim-data.Rmd
added loads of new stuff

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

### 2016-12-08 fix_apa_citations.R

@@ -1,0 +1,23 @@
+library("stringr")
+
+fix_inline_citations <- function(text) {
+    re_inline_year <- "[(]\\d{4}[)]"
+    re_author <- "[[:alpha:]- ]+"
+    re_author_year <- paste(re_author, re_inline_year)
+    re_ampersand <- " & "
+    re_ampersand_author_year <- sprintf("%s(?=%s)", re_ampersand, re_author_year)
+    str_replace_all(text, re_ampersand_author_year, " and ")
+}
+
+do_fix <- c(
+    "Jones & Name (2005) found...",
+    "Jones & Hyphen-Name (2005) found...",
+    "Jones & Space Name (2005) found...",
+    "Marge, Maggie, & Lisa (2005) found...")
+
+
+do_not_fix <- c(
+    "...have been found (Jones & Name, 2005)",
+    "...have been found (Jones & Hyphen-Name, 2005)",
+    "...have been found (Jones & Space Name, 2005)",
+    "...have been found (Marge, Maggie, & Lisa, 2005)")

### 2017-01-22 draft-manuscript.Rmd
added new content and bibliography

--- 2016-12-08 manuscript.Rmd
@@
 # title             : "Probabilistic Computations Underlying Simulated Self-Rotations"
+# title: Probabilistic Computations Underlying Vestibular Cognition
 shorttitle        : "A Computational Perspective"
@@
     - name          : "Andrew W. Ellis"
-    affiliation   : "1"
+# affiliation   : "1"
 corresponding : yes    # Define only one corresponding author
@@
 - name          : "Fred W. Mast"
-    affiliation   : "1"
+# affiliation   : "1"
 
@@
 
-bibliography      : ["references.bib"]
+bibliography      : ["bibliography.bib"]
 
 figsintext        : yes
-figurelist        : no
-tablelist         : no
+figurelist        : yes
+tablelist         : yes
 footnotelist      : no
 lineno            : yes
+citeproc          : yes
+csl               : apa.csl
 
@@
 class             : "man"
-# output            : papaja::apa6_word
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
+# output            : papaja::apa6_pdf
 ---
@@
 
+> Perspective taking: ability to project our body image onto other actors.
+
+- higher level variables generate an entire sequence of control commands
+- show models:
+    + passive model for low-level sensory inference
++ active model with control input for anticipating sensory consequences of active movement
++ the active model can be used for:
+    - estimating ex-afference
+- "interpreting" senosry signals when these are very noisy/difficult.
+In other words, when the stimulus is difficult, we need to interpret the data.
+We claim that this makes use of the control input model.
+- intereference as the result of
++ inappropriately applying active control model for sensory estimation, leadin to mis-specified model
++ wrong prior during model selection
+- show various scenarios for interference:
+    + decision making for motion direction discrimination
++ imaginining motion may result in inadvertently using an active motion model for inference in a subsequent trial.
+IN a discrimination trial, participants need to perfom decision-making.
+THerefore, we need to understand decision-making.
+We present a model of how decision-making in a 2afc experiment may occur (Bayesian model selection -> sequential Bayes factor)
++ if imagery was in a specific direction, this may result in an expectation. Define prior, expectation, anticipation, etc.
+
 generate some data:
@@
 
-# Introduction
-
 ## Notes
 We need to simulate the behaviour of a complex dynamical system.
+
+Why does the brain need to simulate?
+Because in the case of complex, non-linear transformations, the brain does not have acces to analytical solutions, and must therefore rely on simulations.
+Intuitively, the brain uses the 'cheap' solution of simulating a dyanmic system (i.e.
+movement of the body through space), because it already posseses the machinery to perform these simulations, and because analytic solutions to complex mulit-dimensional rotations and translation are hard to obtain.
 
@@
 ![SPT and imagined self-motion involve common computations](diagrams/fig-1-common-process.pdf)
+
 But what are these computations? The brain could either solve the equation for the final position, or could run foward simulation of the dynamic system.
+
+> SPT and mental imagery as active construction, not just memory recall
 
@@
 ## Imagined Self-Motion
+[@Tian:2012ui]
+Is mental imagery an internal predictive process?
+
+Mental imagery is further hypothesized to be a predictive process (for future perceptual states), in which the dynamics of perceptual experience can be retrieved/calculated and reconstructed internally (Moulton and Kosslyn, 2009).
+
+Mental imagery is oftern portrayed as being the result of top-down reconstruction of activity in perceptual circuits in the absence of physical stimulation of the sensors, and is opposed to bottom-up perception.
+However, in the case of the vestibular system, it becomes clear that this view is inadequate (the same is true for other modalities): the sensory that the vestibular sensors provide is noisy and ambiguous (examples), and in order to deal with these facts, the brain must use several strategies (which are well-known from control theory).
+It is assumed that the brain uses (at least approximatively) statistically optimal filtering algorithms.
+Initia research cast the type of processing in the framework of control theory (optimal observer models); however @Selva:2012gsa showed that these models are equivalent to statistically optimal models (Kalman filters), in the case of linear dynamical systems.
+
+That is, given an initial point, the series of future perceptual states can be internally simulated by following the regularity (temporal and causal constraints) stored in declarative memory (general knowl- edge).
+The mapping between internal simulation and the per- ception of external stimulation is thought not to be necessarily isomorphic (Goldman, 1989), as only the essential intermediate states are required to have a one-to-one mapping (Fisher, 2006).
 
@@
 
-The computations performed for SPT require a generative model. Further, we can identify two main topics:
+                                                                                                                        The computations performed for SPT require a generative model.
+Further, we can identify two main topics:
 
@@
 Exteranlly applied forces need to be compensated for, whereas this is often not the case for volitional movements.
+
+# Interferene
+We walk through several common conceptions of "interference" and explain these in terms of an agent performing mental simulations whilst simultaneously attempting to interact with its environment by performing sensory inference.
+
+We can define interference as the phenomenon that sensory inference (filtering) is affected by another process (simulation) using (parts of) the same generative model.
+The agent has two goals: firstly, it has to interact with the environment, using its sensors, and it has to analyze the data.
+Secondly, the agent wishes to plan or simulate, using the same generative model.
+It is currently unclear to what extent the brain is able to segregate the two processes.
 
@@
 # Model description
+
+# Particle Filtering
+Using difference equations, the stochastic state space model can be written as:
+    $$ x_{t} = f(x_{t-1}, u_{t-1}, \nu) $$
+    $$ y_{t} = g(x_{t}, \eta) $$
+
+    where ...
+
+The state space model consists of an unobservale state process, $x$, and an observable process $y$.
+The latent state is modelled as a first-order Markov process, which means that $x_t$ is conditionally independent of the past, given the previous value.
+Therefore, the latent process is completely described by its iniial conditions, $x_0$, and the transition distribution.
+The observation process depends only on the current state.
 

### 2017-01-22 manuscript.Rmd
added new content and bibliography

--- 2016-12-05 manuscript.Rmd
@@
 ---
-title             : "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
-shorttitle        : "A Computational Perspective"
+title             : "Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion"
+# title             : "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
+# title             : "Probabilistic Computations Underlying Simulated Self-Rotations"
+# title: Probabilistic Computations Underlying Vestibular Cognition
+shorttitle        : "Vestibular Contributions to SPT"
 
@@
   - name          : "Andrew W. Ellis"
-    affiliation   : "1"
+    # affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
@@
   - name          : "Fred W. Mast"
-    affiliation   : "1"
+    # affiliation   : "1"
 
 affiliation:
-  - id            : "1"
-    institution   : "Department of Psychology, University of Bern"
+  # - id            : "1"
+  - institution   : "Department of Psychology, University of Bern, Switzerland"
 
 author_note: >
-  Andrew W. Ellis, Fred W. Mast, Department of Psychology.
+  Andrew W. Ellis & Fred W. Mast, Department of Psychology.
 
@@
 
-  Raw data and analysis scripts are available at https://osf.io/u26mq.
+  Code for simulations is available at https://osf.io/u26mq.
 
 abstract: >
-    Recent research showed the involvement of vestibular information in cognitive processes.
-We provide a computational framework for vestibular cognition and thereby focus on perspective transformations, which require a simulated rotation of the self.
+    Recent research has indicated that the vestibular network is involved in cognitive processes.
+However, the nature of this involvement remains unclear.
+We provide a computational framework for vestibular cognition and discuss this in the context of perspective transformations, which require a simulated rotation of the self.
 We explore how the ability to model the effects of body rotation using vestibular input might enable the ability to imagine body rotations.
@@
 
-keywords          : Probabilistic graphical models, Particle filter, Embodied cognition, Vestibular cognition, Spatial perspective taking, Mental imagery
-wordcount         : "X"
+keywords: probabilistic graphical models, particle filter, embodied cognition, vestibular cognition, spatial perspective taking, mental imagery
+wordcount: "X"
 
-bibliography      : ["references.bib"]
+bibliography: "bibliography.bib"
 
 figsintext        : yes
-figurelist        : no
-tablelist         : no
+figurelist        : yes
+tablelist         : yes
 footnotelist      : no
 lineno            : yes
+citeproc          : yes
+csl               : apa.csl
 
@@
 # output            : papaja::apa6_word
-output            : papaja::apa6_pdf
+output:
+     papaja::apa6_pdf:
+        pandoc_args:
+        - --filter
+        - pandoc-fignos
+fignos-cleveref: On
+fignos-plus-name: Figure
+header-includes: \usepackage{caption}
 ---
 
-```{r message = FALSE, warning = FALSE}
-library("papaja")
+```{r preliminary, message = FALSE, warning = FALSE}
+library(papaja)
+library(pander)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(mvtnorm)
+# library(ggthemes)
+library(cowplot)
+
+source('~/thesis/papers/pgm-spt/R/utils.R')
+source('~/thesis/papers/pgm-spt/R/2d-particle-filter.R')
+source('~/thesis/papers/pgm-spt/simulations/rbiips-model.R')
+
+knitr::opts_chunk$set(cache = TRUE,
+                      fig.width = 12, fig.height = 8,
+                      dpi = 600, fig.path = "../generated-figures/",
+                      dev=c('pdf', 'png', 'cairo_ps'))
 ```
 
-# Results
+# Introduction
+ <!-- convey how the work constraints current theorizing over and above the current literature -->
 
-# Discussion
+<!-- The aim of this paper is to describe potential contributions of the vestibular system to spatial perspective transformations in the context. of probabilistic sensory inference. -->
+
+Adopting the spatial perspective of another person is generally considered an essential cognitive ability, and plays a vital role in our ability to determine what another person can see and to understand another person's action possibilities, in order to coordinate joint actions [@CreemRegehr:2013kx, @Pezzulo:2013iy].
+Many studies have also pointed out that is of fundamental importance for language comprehension and communication [@Beveridge:2013gx], and social cognition in general [@Deroualle:2014ho].
+
+A number of recent studies have demonstrated that the vestibular system, which deals with sensory signal relating to movements of the head, is involved in such spatial transformations [@Lenggenhager:2008et, @Deroualle:2015gk, @Falconer:2012cl, @Gardner:2016kd].
+However, to date, there has little attempt to understand the involvement of the vestibular system in higher-level cognitive abilities from a computational perspective.
+In this paper, we focus specifically on the role of the vestibular system for spatial perspective transformations, and we draw the link between probabilistic computations performed by the vestibular system in the service of real-time interaction with the world, and the brain's ability to run mental simulations for the purpose of adopting the perspective of another person.
+
+Consider the use of deictic pronouns, such as "this" or "that" and "left" or "right".
+These terms only assume a their meaning when placed within a certain reference frame.
+For example, in a dialogue between two speakers, the statement "I would like that one", or "I would like the one on the right" can only be understood under the assumption that the deictic terms "that" and "right" are used in an egocentric reference frame.
+In addition, both interlocutors must be able to access the other's reference frame with respect to an external reference frame, which is independent of both interlocutors.
+In other words: In order to give instructions, the speaker must be able to mentally assume the addressee's position.
+A further example that requires a similar spatial transformation is when two people want to perform a coordinated action, such as lifting a heavy object.
+This requires having knowledge of the other's potential actions in order to adjust one's own actions, and this again requires knowledge of the other's position with an external reference frame.
+Furthermore, it is considered likely that merely possessing a visual representation is not sufficient [citation].
+A judgement about what kind of actions the other person can perform, requires an embodied simulation using a representation of one's own body schema [@Pezzulo:2013iy].
+The information necessary to perform this simulation is thus the position of the addressee's limbs, in their egocentric body coordinates, the position of the addressee's reference frame realtive to the external reference frame, and the difference between the speaker and the addressee's egocentric reference frames.
+Perspective taking thus can be conceived of as a aligning a representation of one's body with that of another person in order to access spatial information from a viewpoint other than one's own egocentric viewpoint.
+In order to achieve this alignment, one must perform translations and rotations of one's own egocentric reference frame.
+Processing translations and rotations of the head and body is the core domain of the vestibular system, and thus is of particular interest for research into mental spatial transformations.
+Spatial transformations in 6 dimensions (translation along 3 axes, and rotations around 3 axes) are highly complex, and non-linear, and it is commonly assumed that such transformations of perspective are performed by running embodied simulations [@Pezzulo:2013iy].
+Thus, it seems that the brain adopts the strategy of simulating motion of the body through space in order achieve a desired target position.
+This effectively amounts to a running a simulation of dynamical system, consisting of limb movements and the resulting sensory consequences.
+Such a simulation has therefore been hypothesized to involve both motor and sensory components [citation].
+@fig:spt-schematic (a) shows a simple scenario in which a speaker wants to give an addressee, standing opposite, verbal instructions regarding the location of an external object.
+In order to give appropriate instructions, the speaker must determine the most suitable action, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred reference frame to quickly simulate the best action.
+In this example, alignment with the addressee's reference frame consists of a forwards translation and a rotation around the vertical axis.
+This can be described as an imagined movement [citation], in which the goal is achieve a specific position.
+In order to do this, the speaker must first determine the parameters related to the movement, such as the type of movement, and the direction, velocity and duration that lead to the desired outcome.
+This means that in order to know what to simulate, the speaker must work backwards from the desired outcome, and infer the actions that are most likely to lead to that outcome.
+This is a well-known problem in the field of reinforcement learning, and planning in humans has been described in similar terms, as reverse inference [@Botvinick:2012fe].
+@fig:spt-schematic (b) shows that if the goal is to simply simulate the perceptual experience of rotation, rather than project a representation of one's body onto a target, the problem is vastly simplified; however, the computational steps remain essential the same.
+In order to imagine a whole-body rotation around the vertical axis, the same parameters must be chosen.
+In this study, we focus on the simplified problem of imagined rotation around the body's vertical (yaw) axis.
+
+![SPT and imagined self-motion involve common spatial transformations of a representation of the self.
+(a) If a speaker's goal is to take the perspective of an addressee, he must perform a translation and rotation of his own egocentric reference frame.
+These operations can be viewed as imagined movements.(b) Focussing on just the rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.pdf){#fig:spt-schematic}
+
+Performing a simulation of the behaviour of a dynamical system requires an internal model of the system.
+Whereas earlier notions of internal models followed the control-theoretic approach taken in sensorimotor research [@Wolpert:1998cc, @Shadmehr:2012vu], this idea has recently been applied more generally due to the use of Bayesian models of perception and cognition [Friston, Griffiths, Tenenbaum].
+
+[@Pickering:2014bo] discuss the relationship between control-theoretic notions of forward models and and more genereal concepts of a generative model in a hierarchical Bayesian framework.
+
+In short, running a simulation of a system requires a generative model. Where does this knowledge come from? __Here we make the link to models of vestibular sensry processing__.
+
+For this reason, we claim that the vestibular system offers a very promising field of research. On the one hand,
+
+## Involvement of the vestibular system
+If the vestibular system is involved in spatial perspective taking, then this must be related to mental whole-body rotation, as the vestibular system is directly concerned with rotations and translations.
+In particular, the semi-circular canals (SCC) measure angular accelerations, and the upstream vestibular system provides the neural circuitry involved in processing these sensory measurements.
+The vestibular system faces a number of challenges in interpreting these sensory measurements.
+On the one hand, the signal provided by the SCC and otoliths are very noisy, and ambiguous [MacNeilage]; the otoliths respond to translations and tilt of the head relative to gravity, and in order to disambiguate between translation and tilt, the brain must perform sophisticated computations at the level of the brain stem.
+It has become clear that the nature of these computations is probabilistic, and that the brain makes use of strong prior beliefs in order to interpret these sensory signals.
+
+A further challenge faced by the brain is that it cannot determine whether the sensory signals provided by the SCC and otolith signals m
+
+## Focus on vestibular system
+- active motion, efference copy, etc
+- generative model
+
+- mental simulations involves forward models
+- interesting, because vestibular system is well-described in terms of control theory
+
+<!-- give a few short examples: -->
+<!-- - deictic pronouns: here, there, left, right, this or that -->
+<!-- - performing joint actions (shared action space): lifting a heavy object -->
+
+Imagined motion through space vs. rotating contents of visual buffer
+
+# Dynamic Bayesian models for sensory inference
+
+## Related Work: Previous models for sensory inference
+We begin by briefly reviewing previous probabilistic models of sensory inference:
+Whereas early studies largely adopted approached inspired by control theory, several recent studies have described vestibular sensory processing as dynamic Bayesian inference [@PaulRMacNeilage:2008bv, @Karmali:2012cv, @Laurens:2006be].
+In particular, @Karmali:2012cv and @Laurens:2006be described sensory inference in terms of particle filtering, which is a particular inference algorithm that can be used to perform inference in a non-linear state-space model [@Doucet:2000bh, @Doucet:2009us].
+
+State-space models can be seen as special cases of more general Bayesian hierarchcal models [@Bishop:2006ui; @Murphy:2012ua; @Koller:2009ty].
+
+## Summary
+
+## Reframe as PGM
+
+# Extend previous models with control input
+
+Here we extend the basic particle filtering (algorithm) model, in order to:
+
+- provide the capability to model the sensory consequences of self-initiated actions
+- provide the capability to run a Monte Carlo simulation using the same generative model as is used for sensory inference
+
+## State-Space Model
+
+## Inference Algorithm 
+Sequential Monte Carlo
+
+# Modelling questions
+Experimental results to be explained: @Wertheim:2001cb report a cognitive suppression of otolith responses, based on participants' prior beliefs about possible motions.
+
+# An Investigation of Various Priors
+Here, we identify the various forms of predictive processing in the context of real-time sensory inference using SMC
+## Related to Filtering Algorithm
+## Higher-level Priors
+
+# Various sources of interference
+Here, we identify various sources of intereference (2-way) between thought (higher-level cognition) and lower-level sensory processing
+
+## Model mis-specification
+## Re-use of same generative model
+This may lead to leakage, i.e. inadvertently incorporating sensory evidence
+
+## Decision-making
+The interference may be explained in terms of higher-level perceptual decision -making mechanisms
+## evidence accumulation
+## bias (Starting point)
+
+# Simulations
+
+# General Discussion
+
+# Summary and conclusions
+# Motion Data
+
+```{r generate-data-1, echo=FALSE, message=FALSE, warning=FALSE}
+dt <- 0.1
+motiondata <-  generate_data(T = 2, frequency = 0.5,
+                             amplitude = 20,
+                             sensor_sd = 2.0)
+# plot_trajectories(motiondata = motiondata)
+
+motiondata_hidden <- motiondata %>%
+    gather(type, value = value, -time) %>%
+    filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration")))
+
+p1 <- motiondata_hidden %>%
+        ggplot(aes(x = time, y = value)) +
+            geom_line(aes(linetype = type), size = 1.4) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "white",
+                       fill = "white", size = 6, stroke = 2) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "black",
+                       fill = "white", size = 4, stroke = 2) +
+            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+            xlab("Time (sec)") +
+            ylab("Angular position (deg)") +
+            theme_apa() +
+            theme(legend.title=element_blank())
+
+motiondata_2 <-  generate_data(T = 2, frequency = 0.5,
+                             amplitude = -20,
+                             sensor_sd = 1.8)
+# plot_trajectories(motiondata = motiondata)
+
+motiondata_hidden_2 <- motiondata_2 %>%
+    gather(type, value = value, -time) %>%
+    filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration")))
+
+p2 <- motiondata_hidden_2 %>%
+        ggplot(aes(x = time, y = value)) +
+            geom_line(aes(linetype = type), size = 1.4) +
+            geom_point(data = motiondata_2,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "white",
+                       fill = "white", size = 6, stroke = 2) +
+            geom_point(data = motiondata_2,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "black",
+                       fill = "white", size = 4, stroke = 2) +
+            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+            xlab("Time (sec)") +
+            ylab("Angular position (deg)") +
+            theme_apa() +
+            theme(legend.title=element_blank())
+
+cowplot::plot_grid(p1, p2, nrow = 2, labels = c("a", "b"))
+```
+
+# Test Simulations
+
+```{r passive, echo=FALSE, message=FALSE, warning=FALSE}
+params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
+                x_init = c(0, 0),
+                sdx_init = 0.5*eye(2),
+                A = 20, N = 1000,
+                dt = 0.1, freq = 0.5,
+                transfun = transfun)
+
+passive <- run_particle_filter(motiondata = motiondata, params = params,
+                            resample_particles = TRUE, rs_thresh = 0.5,
+                            selfmotion = "passive")
+
+plot(passive, panel_label = "", print_loglik = TRUE, predict = FALSE, imagine = FALSE)
+```
+
+```{r echo=FALSE}
+plot_pf_estimates(object = passive, predict = FALSE)
+```
+
+# Tables
+
+  Right     Left     Center     Default
+-------     ------ ----------   -------
+     12     12        12            12
+    123     123       123          123
+      1     1          1             1
+
+Table:  Demonstration of simple table syntax.
+
+-------------------------------------------------------------
+ Centered   Default           Right Left
+  Header    Aligned         Aligned Aligned
+----------- ------- --------------- -------------------------
+   First    row                12.0 Example of a row that
+                                    spans multiple lines.
+
+  Second    row                 5.0 Here's another one. Note
+                                    the blank line between
+                                    rows.
+-------------------------------------------------------------
+
+Table: Here's the caption. It, too, may span
+multiple lines.
 

### 2017-01-22 notes.Rmd
added new content and bibliography

# head direction cells

[@Taube:2007fk]

- mental simulation cannot just er-use the state used in sensory inference, because the velocity is fed to the head directino cells. The brain would get confused

- directinal signal is not just a simple integration of velocity signals from the vestibular system
- signals from motor efferecence copy or proprioceptive systems appear to be crucial for upating the preferred firing direction over time
- what about this? you don't know where you are, because you were passively moved. Can you update if I tell you (verbally instruct you)?
- vestibular input is eesntial for establishing the head direction signal? How can we reconcile this with the assumption that re-afference is cancelled, i.e. input due to active motion is cancelled. Maybe it's the control model...

# Corollary discharge
[@Crapse:2008jc]

# Predictive control 
[@Stahl:2011dl]

- movemements can induce sensory input
- sensory input due to own actions is indistinguishable from input due to external forces
- strategy: efference copy -> corollary discharge

it seems intuitively obvious that this mechanism has something to do with mental imagery.
This has been pointed out many times; mental imagery and mental simulation more generelly is somehow related to the capacity to compute corollay discharge, based on the efference copy and a forward model.

> Necessary components:  efference copy and a forward model

Sensory receptors don't care about their input. The brain has to deal with the problem further downstream

### 2017-01-22 fix_apa_citations.R

@@ -1,23 +1,0 @@
-library("stringr")
-
-fix_inline_citations <- function(text) {
-    re_inline_year <- "[(]\\d{4}[)]"
-    re_author <- "[[:alpha:]- ]+"
-    re_author_year <- paste(re_author, re_inline_year)
-    re_ampersand <- " & "
-    re_ampersand_author_year <- sprintf("%s(?=%s)", re_ampersand, re_author_year)
-    str_replace_all(text, re_ampersand_author_year, " and ")
-}
-
-do_fix <- c(
-    "Jones & Name (2005) found...",
-    "Jones & Hyphen-Name (2005) found...",
-    "Jones & Space Name (2005) found...",
-    "Marge, Maggie, & Lisa (2005) found...")
-
-
-do_not_fix <- c(
-    "...have been found (Jones & Name, 2005)",
-    "...have been found (Jones & Hyphen-Name, 2005)",
-    "...have been found (Jones & Space Name, 2005)",
-    "...have been found (Marge, Maggie, & Lisa, 2005)")

### 2017-01-22 manuscript.Rmd
before possible computer crash

@@
 
-Here we extend the basic particle filtering (algorithm) model, in order to:
+For sensory inference of a passive disturbance, the algorithm can be run with fixed parameters. Here we extend the basic particle filtering (algorithm) model, in order to:
 
-- provide the capability to model the sensory consequences of self-initiated actions
+- __provide the capability to model the sensory consequences of self-initiated actions__ (i.e. predict/infer the consequences of one's own actions)
 - provide the capability to run a Monte Carlo simulation using the same generative model as is used for sensory inference
 
-## State-Space Model
+## Generative Model
+The generative model is a state-space model, which itself is a dynamic model consisting of latent (state variables) and observed (measured) variables.
 
@@
 ## Higher-level Priors
+control input -> infer the consequences of one's own actions.
 
@@
 ## Model mis-specification
+If an observer uses a strong higher level prior, i.e. the wrong direction, this means that the data may be completely mis-interpreted. (call this: the effect of inappropriate priors [@Schwabe:2008ho])
+
 ## Re-use of same generative model
@@
 
+## Summary
+Interference may the result of either inappropriate priors, leading to a mis-specified filtering algorithm, or biased decision-making.
+We should be able to investigate which by applying cognitive process models that explain response times and choices.
+
 # Simulations
@@
 # General Discussion
+## A comparison with previous models
+Unfortunately, there aren't many, but see [@Hiatt:2004tv]
+
+## Support for our model: Nigmatullina
+
+## Neuronal implementation
+[@Lopez:2012ek; @Lopez:2011cc]
+[@Klingner:2016ia]
+[@zuEulenburg:2012bh]
 
 # Summary and conclusions
+
+# Acknowledgements
+
 # Motion Data

### 2017-01-23 manuscript.Rmd
up to particle filter

@@
 abstract: >
-    Recent research has indicated that the vestibular network is involved in cognitive processes.
+        Recent research has indicated that the vestibular network is involved in various cognitive tasks involving mental spatial transformations.
 However, the nature of this involvement remains unclear.
-We provide a computational framework for vestibular cognition and discuss this in the context of perspective transformations, which require a simulated rotation of the self.
+Here, we provide a computational framework for vestibular cognition and discuss this in the context of perspective transformations, which require a simulated rotation of the self.
 We explore how the ability to model the effects of body rotation using vestibular input might enable the ability to imagine body rotations.
-We approach this problem by considering a simple agent endowed with a complex statistical model of the hidden causes of its sensory input.
-We explain how this statistical model enables the agent to predict its sensory input and hidden sensory variables.
-Based on a probabilistic graphical model, we formalize qualitative concepts of mental simulation.
+We approach this problem by extending previous models of vestibular sensory processing to include the ability to model the sensory consequences of self-initiated movements, and discuss how this can be further extended to allow mentally simulated rotations of the self.
+We ...
 Predicting the sensory input is key to understanding imagined rotations.
@@
 footnotelist      : no
+number_sections   : FALSE
 lineno            : yes
-citeproc          : yes
-csl               : apa.csl
+linkcolor         : "blue"
+# citeproc          : yes
+# csl               : apa.csl
 
@@
         - pandoc-fignos
+        fig_caption: true
 fignos-cleveref: On
@@
 
-Adopting the spatial perspective of another person is generally considered an essential cognitive ability, and plays a vital role in our ability to determine what another person can see and to understand another person's action possibilities, in order to coordinate joint actions [@CreemRegehr:2013kx, @Pezzulo:2013iy].
-Many studies have also pointed out that is of fundamental importance for language comprehension and communication [@Beveridge:2013gx], and social cognition in general [@Deroualle:2014ho].
+<!-- The goal of this study is to discuss the role of the vestibular system in embodied mental transformations. -->
+
+Adopting the spatial perspective of another person is generally considered an essential cognitive ability, and plays a vital role in our ability to determine what another person can see and in order to coordinate joint actions [@CreemRegehr:2013kx, @Pezzulo:2013iy].
+Many studies have also pointed out that perspective taking is of fundamental importance for language comprehension and communication [@Beveridge:2013gx], and social cognition in general [@Deroualle:2014ho].
 
 A number of recent studies have demonstrated that the vestibular system, which deals with sensory signal relating to movements of the head, is involved in such spatial transformations [@Lenggenhager:2008et, @Deroualle:2015gk, @Falconer:2012cl, @Gardner:2016kd].
+This is demonstrated in the form of interference effects between mental transformations and concurrent sensory stimulation.
 However, to date, there has little attempt to understand the involvement of the vestibular system in higher-level cognitive abilities from a computational perspective.
@@
 Consider the use of deictic pronouns, such as "this" or "that" and "left" or "right".
-These terms only assume a their meaning when placed within a certain reference frame.
-For example, in a dialogue between two speakers, the statement "I would like that one", or "I would like the one on the right" can only be understood under the assumption that the deictic terms "that" and "right" are used in an egocentric reference frame.
-In addition, both interlocutors must be able to access the other's reference frame with respect to an external reference frame, which is independent of both interlocutors.
-In other words: In order to give instructions, the speaker must be able to mentally assume the addressee's position.
+These terms only assume a their meaning when placed within a certain frame of reference (FOR).
+For example, in a dialogue between two speakers, the statement "I would like that one", or "I would like the one on the right" can only be understood under the assumption that the deictic terms "that" and "right" are used in an egocentric FOR.
+In addition, both interlocutors must be able to access the other's FOR with respect to an external FOR.
+In other words: In order to give instructions, the speaker must be able to mentally assume the addressee's position in space.
 A further example that requires a similar spatial transformation is when two people want to perform a coordinated action, such as lifting a heavy object.
-This requires having knowledge of the other's potential actions in order to adjust one's own actions, and this again requires knowledge of the other's position with an external reference frame.
-Furthermore, it is considered likely that merely possessing a visual representation is not sufficient [citation].
-A judgement about what kind of actions the other person can perform, requires an embodied simulation using a representation of one's own body schema [@Pezzulo:2013iy].
-The information necessary to perform this simulation is thus the position of the addressee's limbs, in their egocentric body coordinates, the position of the addressee's reference frame realtive to the external reference frame, and the difference between the speaker and the addressee's egocentric reference frames.
+This requires having knowledge of the other's potential actions in order to adjust one's own actions, and this again requires knowledge of the other's position with an external FOR.
+A judgement about what kind of actions the other person can perform requires an embodied simulation using a representation of one's own body schema [@Pezzulo:2013iy].
+Giving an instruction to another person thus requires running an embodied simulation in order to infer what kind of motor commands the other person should perform __[Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod, 2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo, 2011a,b]__.
+This embodied simulation is performed using one's own body schema.
+Communicating this information to the other actor requires knowing the position of the addressee's FOR relative to the external reference frame, and the difference between the speaker and the addressee's egocentric reference frames.
 Perspective taking thus can be conceived of as a aligning a representation of one's body with that of another person in order to access spatial information from a viewpoint other than one's own egocentric viewpoint.
@@
 Processing translations and rotations of the head and body is the core domain of the vestibular system, and thus is of particular interest for research into mental spatial transformations.
-Spatial transformations in 6 dimensions (translation along 3 axes, and rotations around 3 axes) are highly complex, and non-linear, and it is commonly assumed that such transformations of perspective are performed by running embodied simulations [@Pezzulo:2013iy].
-Thus, it seems that the brain adopts the strategy of simulating motion of the body through space in order achieve a desired target position.
-This effectively amounts to a running a simulation of dynamical system, consisting of limb movements and the resulting sensory consequences.
-Such a simulation has therefore been hypothesized to involve both motor and sensory components [citation].
-@fig:spt-schematic (a) shows a simple scenario in which a speaker wants to give an addressee, standing opposite, verbal instructions regarding the location of an external object.
-In order to give appropriate instructions, the speaker must determine the most suitable action, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred reference frame to quickly simulate the best action.
+Spatial transformations in 6 dimensions (translation along 3 axes, and rotations around 3 axes) are highly complex, and non-linear, and we propose that the brain adopts the strategy of simulating motion of the body through space in order achieve a desired target position.
+This effectively amounts to a running a simulation of dynamical system, consisting of simulated limb movements and the resulting sensory consequences.
+\autoref{fig:spt-schematic} (a) shows a simple scenario in which a speaker wants to give an addressee, standing opposite, verbal instructions about an action to perform on an external object.
+The speaker must determine the most suitable action, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred reference frame to quickly simulate the best action.
 In this example, alignment with the addressee's reference frame consists of a forwards translation and a rotation around the vertical axis.
-This can be described as an imagined movement [citation], in which the goal is achieve a specific position.
-In order to do this, the speaker must first determine the parameters related to the movement, such as the type of movement, and the direction, velocity and duration that lead to the desired outcome.
-This means that in order to know what to simulate, the speaker must work backwards from the desired outcome, and infer the actions that are most likely to lead to that outcome.
-This is a well-known problem in the field of reinforcement learning, and planning in humans has been described in similar terms, as reverse inference [@Botvinick:2012fe].
-@fig:spt-schematic (b) shows that if the goal is to simply simulate the perceptual experience of rotation, rather than project a representation of one's body onto a target, the problem is vastly simplified; however, the computational steps remain essential the same.
-In order to imagine a whole-body rotation around the vertical axis, the same parameters must be chosen.
-In this study, we focus on the simplified problem of imagined rotation around the body's vertical (yaw) axis.
+This can be described as an imagined movement, in which the goal is achieve a specific position.
 
-![SPT and imagined self-motion involve common spatial transformations of a representation of the self.
-(a) If a speaker's goal is to take the perspective of an addressee, he must perform a translation and rotation of his own egocentric reference frame.
-These operations can be viewed as imagined movements.(b) Focussing on just the rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.pdf){#fig:spt-schematic}
+![Perspective taking and imagined self-motion involve spatial transformations of a representation of the self.
+(a) If a speaker's goal is to take the perspective of an addressee, he must first solve the inverse problem of determining the appropriate action, and then run a forwards simulation in order to perform a translation and rotation of his own egocentric reference frame.
+These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.pdf){#fig:spt-schematic width=50%}
 
-Performing a simulation of the behaviour of a dynamical system requires an internal model of the system.
-Whereas earlier notions of internal models followed the control-theoretic approach taken in sensorimotor research [@Wolpert:1998cc, @Shadmehr:2012vu], this idea has recently been applied more generally due to the use of Bayesian models of perception and cognition [Friston, Griffiths, Tenenbaum].
+In order to do this, the speaker must first determine what type of movement to perform, and the direction, velocity and duration that lead to the desired outcome.
+This means that in order to know what to simulate, the speaker must first work backwards from the desired outcome, and infer the actions that are most likely to lead to that outcome.
+This is known in the field of motor control as the inverse problem [@Wolpert:1998td].
+In order to achieve the desired position, the motor commands can then be simulated, and the expected sensory consequences can be estimated; this is known as the forward problem.
+@fig:spt-schematic (b) shows that if the goal is to simply simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action.
+As demonstrated by @Penny:2013iv, these computations can be performed in a common probabilistic model.
+Interestingly, this has been discussed in the context of reinforcement learning, and planning in humans has been described as reverse inference [@Botvinick:2012fe].
+Further, [@Pickering:2014bo] discuss the relationship between control-theoretic notions of inverse/forward models, known collectively as internal models, and and more general concepts of a generative model in a hierarchical Bayesian framework.
 
-[@Pickering:2014bo] discuss the relationship between control-theoretic notions of forward models and and more genereal concepts of a generative model in a hierarchical Bayesian framework.
+Performing a simulation of body movements and estimating the sensory consquences amounts to simulating the behaviour of a dynamical system; this requires a generative model of the system.
+The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations.
+The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Baysian inference.
+The vestibular system faces a number of challenges in interpreting the sensory measurements provided by the semi-circular canals (SCC) and otolith organs.
+On the one hand, these signals  are very noisy, and ambiguous [@PaulRMacNeilage:2008bv]; the otoliths respond both to translations and tilt of the head relative to gravity, and in order to disambiguate between translation and tilt, the brain must perform sophisticated computations at the level of the brain stem.
+It has become clear that the nature of these computations is probabilistic, and that the brain makes use of strong prior beliefs in order to interpret these sensory signals.
+A further challenge faced by the vestibular system is that it needs to distinguish between sensory signals that are the result of active motion, and those resulting from external forces [@Cullen:2009dc, @Cullen:2014gx].
+The ability to distinguishing re-afferent from ex-afferent sensory signals is an essential capability of all organisms that move actively [@Crapse:2008jc].
 
-In short, running a simulation of a system requires a generative model. Where does this knowledge come from? __Here we make the link to models of vestibular sensry processing__.
+Furthermore, the apparent connection between mental imagery and predictive processing has repeatedly been alluded to [@Moulton:2009ij, @Clark:2012vf, @Gambi:2015gv], and @Grush:2004wx claims that mental imagery is performed by emulator circuits implementing a forward model.
+These ideas can also be addressed from a computational point of view in the context of dynamic Bayesian inference, where there are multiple roles for sensory predictions, over different time scales.
+This was recently discussed by [@Tian:2012ui] in the context of imagined speech.
 
-For this reason, we claim that the vestibular system offers a very promising field of research. On the one hand,
+We proceed by introducing recent models of Bayesian inference for real-time, dynamic sensory processing of vestibular data, based on particle filters [@Laurens:2006be, @Karmali:2012cv], and we reframe these as probabilistic graphical models.
+This allows us to investigate the generative model separately from the algorithm used for inference, and we then extend the basic model for sensory inference in order to provide the ability to model the consequences of self-initiated, active movements.
+Our aims are to:
 
-## Involvement of the vestibular system
-If the vestibular system is involved in spatial perspective taking, then this must be related to mental whole-body rotation, as the vestibular system is directly concerned with rotations and translations.
-In particular, the semi-circular canals (SCC) measure angular accelerations, and the upstream vestibular system provides the neural circuitry involved in processing these sensory measurements.
-The vestibular system faces a number of challenges in interpreting these sensory measurements.
-On the one hand, the signal provided by the SCC and otoliths are very noisy, and ambiguous [MacNeilage]; the otoliths respond to translations and tilt of the head relative to gravity, and in order to disambiguate between translation and tilt, the brain must perform sophisticated computations at the level of the brain stem.
-It has become clear that the nature of these computations is probabilistic, and that the brain makes use of strong prior beliefs in order to interpret these sensory signals.
-
-A further challenge faced by the brain is that it cannot determine whether the sensory signals provided by the SCC and otolith signals m
-
-## Focus on vestibular system
-- active motion, efference copy, etc
-- generative model
-
-- mental simulations involves forward models
-- interesting, because vestibular system is well-described in terms of control theory
-
-<!-- give a few short examples: -->
-<!-- - deictic pronouns: here, there, left, right, this or that -->
-<!-- - performing joint actions (shared action space): lifting a heavy object -->
-
-Imagined motion through space vs. rotating contents of visual buffer
+1) identify various forms of predictive processing in the context of dynamic sensory processing.
+2) identify various sources of interference between higher-level cognition and lower-level sensory processing.
+3) provide suggestions for future research aiming to investigate the connections between cognitive and sensory processing.
 
@@
 
-## Related Work: Previous models for sensory inference
-We begin by briefly reviewing previous probabilistic models of sensory inference:
-Whereas early studies largely adopted approached inspired by control theory, several recent studies have described vestibular sensory processing as dynamic Bayesian inference [@PaulRMacNeilage:2008bv, @Karmali:2012cv, @Laurens:2006be].
-In particular, @Karmali:2012cv and @Laurens:2006be described sensory inference in terms of particle filtering, which is a particular inference algorithm that can be used to perform inference in a non-linear state-space model [@Doucet:2000bh, @Doucet:2009us].
+## Previous work
+Whereas early models of vestibular sensory processing were inspired by control theory [@Borah:1988fd, @Merfeld:1999cg], several recent studies have described vestibular sensory processing as dynamic Bayesian inference [@Laurens:2006be, @PaulRMacNeilage:2008bv, @Karmali:2012cv].
+@Selva:2012gsa
+provide a recent review of the relationship between optimal observer models and models which employ Kalman filtering, and show that these are largely equivalent.
+Both @Karmali:2012cv and @Laurens:2006be described vestibular sensory inference in terms of particle filtering, which is a particular inference algorithm that can be used to perform inference in state-space models.
+[@Doucet:2000bh, @Doucet:2009us].
+Using their model, @Laurens:2006be were able to simulate perceptual responses to centrifugation and off-vertical axis rotations.
 
-State-space models can be seen as special cases of more general Bayesian hierarchcal models [@Bishop:2006ui; @Murphy:2012ua; @Koller:2009ty].
+A further well-known phenomenon of vestibular perception is velocity storage, which is defined as the slower decay of perceptual and oculomotor responses, relative to the SCC afferent responses [@Raphan:1979hn], __[Robinson 1977]__.
+This is generally seen as evidence for the existence of an internal model in the processing of vestibular signals.
+@Karmali:2012cv focussed on one-dimensional angular velocity during yaw rotations; their model was able to explain the phenomenon of velocity storage as a consequence of particle filtering, with the time constant of velocity storage depending on the variability of the sensory afferent signals.
+
+## Generative model for sensory inference
+We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can perform inference in state-space models.
+While Kalman filters are restricted to linear Gaussian State-space models, particle filtering can be applied to arbitrary nonlinear models.
+State-space models can be seen as special cases of more general Bayesian hierarchical models [@Bishop:2006ui; @Murphy:2012ua].
+Following the same approach as @Penny:2013iv, we first describe the generative model used for sensory inference as a dynamic probabilistic graphical model.
+
+A state-space is special case of a dynamic latent variable model which is suitable for modelling dynamical system in terms of latent, unobservable processes and noisy observations.
+As an example, consider the case where we want to model the velocity and position during a leftward head turn.
+While the SCC measures angular acceleration, the afferent signals are thought to be proportional to the angular velocity of the head with a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv].
+The measurements are observations of the true, unobservable angular velocity; however, they are contaminted by Gaussian noise, which is determined by characteristics of the sensory organs.
+\autoref{fig:head-turn-data} shows a sequence of noisy measurements during a two-second turn of the head.
+The acceleration generating the movement is sinusoidal, given for each time step $t$ by $Asin(2\pi t/T)$, where $A$ is the amplitude of the movement, $T = 1/f$ is the duration of the movement, and $f$ is the frequency.
+This acceleration results in a peak velocity $\omega_{peak} = AT/\pi$ and a final angular position $\theta_{final} = AT^2/2\pi$.
+
+```{r head-turn-data, echo=FALSE, fig.cap= "Sensory measurements obtained during a short head turn.
+The goal is to infer the latent angular velocity and position, based on the noisy measurements.
+\\label{fig:head-turn}", message=FALSE, warning=FALSE, dpi=100}
+dt <- 0.1
+motiondata <-  generate_data(T = 2, frequency = 0.5,
+                             amplitude = 20,
+                             sensor_sd = 2.0)
+# plot_trajectories(motiondata = motiondata)
+
+motiondata_hidden <- motiondata %>%
+    gather(type, value = value, -time) %>%
+    filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration")))
+
+p1 <- motiondata_hidden %>%
+        ggplot(aes(x = time, y = value)) +
+            geom_line(aes(linetype = type), size = 1.4) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "white",
+                       fill = "white", size = 6, stroke = 2) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "black",
+                       fill = "white", size = 4, stroke = 2) +
+            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+            xlab("Time (sec)") +
+            ylab("Angular position (deg)") +
+            theme_apa() +
+            theme(legend.title=element_blank())
+# p1
+plot_trajectories(motiondata = motiondata)
+```
+
+The key idea is that we can write model the dynamics of a system by the evolution of the latent state variables in discrete time steps as a Markov process, where the state at each time point $t$ depends only on the previous state at time $t-1$.
+This is referred to as the process model.
+Estimating the velocity $\omega$ and position $\theta$ in discrete time therefore requires writing the kinematic equations as a set of first order difference equations.
+
+$$\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t$$
+$$\omega_{t} = \omega_{t-1}$$
+
+Note that we have omitted the acceleration from the process model, for the sake of simplicity; this amounts to assuming a constant velocity model.
+These kinematic equations form a linear model, and could optionally be written in matrix notation.
+For further simplicity, we retain the current notation.
+Next, we consider the measurements at each time step $t$ as depending only on the current state.
+This simply implements the idea that the sensory signal depends only on the current head velocity.
+The above equations represent the expected values of the state and observation variables.
+In order to complete the generative model, we require a description of the stochastic nature dependancies.
+We model both process and measurement noise a being normally distributed, leading to the following linear Gaussian state-space model:
+
+$$x_t = f(x_{t-1}, \epsilon)$$
+$$y_{t} = g(x_t, \delta)$$
+$$\epsilon \sim N(0, \sigma^2_\epsilon) $$
+$$\delta \sim N(0, \sigma^2_\delta) $$
+
+where the state variable $x_t$ consists of both the velocity $\omega_t$ and position $\theta_t$, and $y_{t}$ denotes the sensory measurement.[^sensormodel] $\epsilon$ and $\delta$ are zero-mean Gaussian noise terms with variances $\sigma^2_\epsilon$ and $\sigma^2_\delta$.
+These refer to the process and measurement noise, respectively.
+The process noise
+
+[^sensormodel]: Note that we although both @Laurens:2006be, @Karmali:2012cv consider a realistic sensor model, for simplicity, we implement the sensor as being a noisy realisation of the velocity.
+Our model is not intended to be a realistic model of the SCC; rather, we focus on the probabilistic computations.
+
+<!-- \autoref{fig:graphical-model-inference} -->
+\autoref{fig:graphical-model-inference} shows a representation of state-space model as a graphical model, unrolled over multiple time steps.
+Starting at time 0, the state variables evolve according to the process model, with the measurements at each time step being generated according to the measurement model.
+The 2-dimensiontal latent state variable is unshaded, to indicate that it cannot be observed, wheres the observed sensory measurements are shaded.
+Whereas previous models were discussed in terms of the particular filtering algorithm used, the graphical notation highlights the fact that a state-space model can be viewed as a dynamic latent factor model [@Blei:2014cp].
+This opens up the possibility of extending the graphical model to include higher-level variables.
+After discussing particle filtering, an inference algorithm often used to infer the values of the latent state variables, we will introduce our extension to the model that will allow a prediction of the consequences of active movements.
+
+<!-- ![Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.
+](diagrams/dynamic-belief-network-inference-1.pdf){#fig:graphical-model-inference width=100%} -->
+
+```{r graphical-model-inference, fig.cap="Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.", echo=FALSE, out.width='100%'}
+knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.pdf')
+```
+
+## Particle filtering
+We give a brief overview of particle filtering, concentrating on those aspects most relevant for this study.
+For a more thorough treatment, see @Doucet:2000bh and @Doucet:2009us.
+@Speekenbrink:2016kc provides a recent review with applications in psychological research.
+Particle filtering is an algorithm for performing approximate Bayesian inference in a graphical model.
+It proceeds by recursively computing the distribution of the current state $x_t$ as a a function of its parent nodes, consisting in our example of the state at the previous time step.
+
+The general procedure for recursive Bayesian inference is to perform alternating prediction and updating steps.
+
+1)	Predict step: At time t, the current state is predicted according to the process model. This prediction is maintained in the form of probability distributions over all the values.
+
+2)	Update step: a measurement is obtained, and this measurement is used in order to update the state estimate.
+
+By recursively estimating the current state, the observations can be processed continuously; thus, the agent does not need to store the entire history of observations – only the previous belief state estimate and the current observation are needed in order to estimate the current state.
+At time step t, the agent observes the value of , and this observation is used to update the previous belief:
+
+\autoref{fig:particle-filter-explained} shows a diagram ...
+
+![Particle filter computations performed at every time step.
+The cloud of particles (grey circles) represents the current belief state.
+Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model.
+This is equivalent to sampling from the prior distribution.
+Then, during the updating step, an observation  is obtained (shown as a black diamond).
+Each particle is assigned an importance weight, according to how well it explains the observation.
+Each particle's diameter reflects its importance weight.
+Subsequently, a new set of particles is formed by weighted multinomial resampling, with each particle's probability of being chosen proportional to its importance weight.
+The particle cloud's mean and standard deviation are shown as light grey circles and lines.
+This resulting cloud of particles represents the posterior distribution.
+This posterior subsequently forms the prior distribution for the next time step.](diagrams/figure-particle-filter-explained.pdf){#fig:particle-filter-explained width=100%}
+
+\autoref{fig:passive} shows the estimated velocity ...
+
+```{r passive, fig.cap="Particle filter applied...", echo=FALSE, message=FALSE, warning=FALSE}
+params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
+                x_init = c(0, 0),
+                sdx_init = 0.5*eye(2),
+                A = 20, N = 1000,
+                dt = 0.1, freq = 0.5,
+                transfun = transfun)
+
+passive <- run_particle_filter(motiondata = motiondata, params = params,
+                            resample_particles = TRUE, rs_thresh = 0.5,
+                            selfmotion = "passive")
+
+plot(passive, panel_label = "", print_loglik = FALSE, predict = FALSE, imagine = FALSE)
+```
+
+```{r eval=FALSE, include=FALSE}
+plot_pf_estimates(object = passive, predict = FALSE)
+```
 
 ## Summary
-
-## Reframe as PGM
 
@@
 # General Discussion
+
+The main result of this study is 
+- place the emerging field of vestibular cognition with a framewotk for dynamic probabilistic inference and draw links to established models of sensory processing
+- interpret result from vestibular cognition studies showing interference between thought and perception as being the result of biased sensory processing
+
 ## A comparison with previous models
-Unfortunately, there aren't many, but see [@Hiatt:2004tv]
+Unfortunately, there aren't many, but see [@Hiatt:2004tv]: imagined motion through space vs. rotating contents of visual buffer
 
@@
 
+### Sensory gating
+[@Gale:2016cx]
+
+### Head direction cells
+Why dynamic simulation? Maybe in order to provide vestibular in put to head direction cells [@Taube:2007fk]
+
 # Summary and conclusions
@@
 
-```{r generate-data-1, echo=FALSE, message=FALSE, warning=FALSE}
+```{r generate-data-1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 dt <- 0.1
@@
 
-# Test Simulations
-
-```{r passive, echo=FALSE, message=FALSE, warning=FALSE}
-params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
-                x_init = c(0, 0),
-                sdx_init = 0.5*eye(2),
-                A = 20, N = 1000,
-                dt = 0.1, freq = 0.5,
-                transfun = transfun)
-
-passive <- run_particle_filter(motiondata = motiondata, params = params,
-                            resample_particles = TRUE, rs_thresh = 0.5,
-                            selfmotion = "passive")
-
-plot(passive, panel_label = "", print_loglik = TRUE, predict = FALSE, imagine = FALSE)
-```
-
-```{r echo=FALSE}
-plot_pf_estimates(object = passive, predict = FALSE)
-```
-
 # Tables

### 2017-01-24 manuscript.tex
fixed figure captions

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Vestibular Contributions to Spatial Perspective Transformations: From
Sensory Inference To Imagined Self-Motion}

  \shorttitle{Vestibular Contributions to SPT}

  \author{Andrew W. Ellis~\& Fred W. Mast}

  \def\affdep{{"", ""}}%
  \def\affcity{{"", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{} Department of Psychology, University of Bern, Switzerland  }

  \authornote{
    \newcounter{author}
    Andrew W. Ellis \& Fred W. Mast, Department of Psychology. AE and FM
    planned and conducted the research; AE and FM analyzed the data and
    wrote the manuscript. Code for simulations is available at
    \url{https://osf.io/u26mq}.

                      Correspondence concerning this article should be addressed to Andrew W. Ellis, Fabrikstrasse 8, 3012 Bern, Switzerland. E-mail: \href{mailto:andrew.ellis@psy.unibe.ch}{\nolinkurl{andrew.ellis@psy.unibe.ch}}
                          }

  \abstract{Recent research has indicated that the vestibular network is involved in
various cognitive tasks involving mental spatial transformations.
However, the nature of this involvement remains unclear. Here, we
provide a computational framework for vestibular cognition and discuss
this in the context of perspective transformations, which require a
simulated rotation of the self. We explore how the ability to model the
effects of body rotation using vestibular input might enable the ability
to imagine body rotations. We approach this problem by extending
previous models of vestibular sensory processing to include the ability
to model the sensory consequences of self-initiated movements, and
discuss how this can be further extended to allow mentally simulated
rotations of the self. We \ldots{} Predicting the sensory input is key
to understanding imagined rotations. Mental imagery goes beyond
prediction, rather, it involves the use of counterfactual queries in a
probabilistic model.}
  \keywords{probabilistic graphical models, particle filter, embodied cognition,
vestibular cognition, spatial perspective taking, mental imagery \\

    \indent Word count: X
  }

\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}

\cftpagenumbersoff{table}
\renewcommand{\cfttabpresnum}{\itshape\tablename\enspace}
\renewcommand{\cfttabaftersnum}{.\space}
\setlength{\cfttabindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cfttabnumwidth}{Table 10.\qquad}

  \usepackage{caption}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

% pandoc-xnos: macro to create a caption without a prefix
\makeatletter
\long\def\@makenoprefixcaption#1#2{
  \vskip\abovecaptionskip
  \sbox\@tempboxa{#2}
  \ifdim \wd\@tempboxa >\hsize
    #2\par
  \else
    \global \@minipagefalse
    \hb@xt@\hsize{\hfil\box\@tempboxa\hfil}
  \fi
  \vskip\belowcaptionskip}
\makeatother

% pandoc-fignos: save original macros
\makeatletter
\let\@oldmakecaption=\@makecaption
\let\oldthefigure=\thefigure
\let\oldtheHfigure=\theHfigure
\makeatother

% pandoc-fignos: environment disables figure caption prefixes
\makeatletter
\newcounter{figno}
\newenvironment{no-prefix-figure-caption}{
  \let\@makecaption=\@makenoprefixcaption
  \renewcommand\thefigure{x.\thefigno}
  \renewcommand\theHfigure{x.\thefigno}
  \stepcounter{figno}
}{
  \let\thefigure=\oldthefigure
  \let\theHfigure=\oldtheHfigure
  \let\@makecaption=\@oldmakecaption
  \addtocounter{figure}{-1}
}
\makeatother

% pandoc-xnos: cleveref fakery
\newcommand{\plusnamesingular}{}
\newcommand{\starnamesingular}{}
\newcommand{\xrefname}[1]{\protect\renewcommand{\plusnamesingular}{#1}}
\newcommand{\Xrefname}[1]{\protect\renewcommand{\starnamesingular}{#1}}
\providecommand{\cref}{\plusnamesingular~\ref}
\providecommand{\Cref}{\starnamesingular~\ref}
\providecommand{\crefformat}[2]{}
\providecommand{\Crefformat}[2]{}

% pandoc-xnos: cleveref formatting
\crefformat{figure}{Figure~#2#1#3}
\Crefformat{figure}{Figure~#2#1#3}

\section{Introduction}\label{introduction}

Adopting the spatial perspective of another person is generally
considered an essential cognitive ability, and plays a vital role in our
ability to determine what another person can see and in order to
coordinate joint actions (Creem-Regehr, Gagnon, Geuss, \& Stefanucci,
2013, Pezzulo, Iodice, Ferraina, \& Kessler (2013)). Many studies have
also pointed out that perspective taking is of fundamental importance
for language comprehension and communication (Beveridge \& Pickering,
2013), and social cognition in general (Deroualle \& Lopez, 2014).

A number of recent studies have demonstrated that the vestibular system,
which deals with sensory signal relating to movements of the head, is
involved in such spatial transformations (Lenggenhager, Lopez, \&
Blanke, 2008, Deroualle, Borel, Devèze, \& Lopez (2015), Falconer \&
Mast (2012), Gardner, Stent, Mohr, \& Golding (2016)). This is
demonstrated in the form of interference effects between mental
transformations and concurrent sensory stimulation. However, to date,
there has little attempt to understand the involvement of the vestibular
system in higher-level cognitive abilities from a computational
perspective. In this paper, we focus specifically on the role of the
vestibular system for spatial perspective transformations, and we draw
the link between probabilistic computations performed by the vestibular
system in the service of real-time interaction with the world, and the
brain's ability to run mental simulations for the purpose of adopting
the perspective of another person.

Consider the use of deictic pronouns, such as \enquote{this} or
\enquote{that} and \enquote{left} or \enquote{right}. These terms only
assume a their meaning when placed within a certain frame of reference
(FOR). For example, in a dialogue between two speakers, the statement
\enquote{I would like that one}, or \enquote{I would like the one on the
right} can only be understood under the assumption that the deictic
terms \enquote{that} and \enquote{right} are used in an egocentric FOR.
In addition, both interlocutors must be able to access the other's FOR
with respect to an external FOR. In other words: In order to give
instructions, the speaker must be able to mentally assume the
addressee's position in space. A further example that requires a similar
spatial transformation is when two people want to perform a coordinated
action, such as lifting a heavy object. This requires having knowledge
of the other's potential actions in order to adjust one's own actions,
and this again requires knowledge of the other's position with an
external FOR. A judgement about what kind of actions the other person
can perform requires an embodied simulation using a representation of
one's own body schema (Pezzulo et al., 2013). Giving an instruction to
another person thus requires running an embodied simulation in order to
infer what kind of motor commands the other person should perform
\textbf{{[}Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod,
2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo,
2011a,b{]}}. This embodied simulation is performed using one's own body
schema. Communicating this information to the other actor requires
knowing the position of the addressee's FOR relative to the external
reference frame, and the difference between the speaker and the
addressee's egocentric reference frames. Perspective taking thus can be
conceived of as a aligning a representation of one's body with that of
another person in order to access spatial information from a viewpoint
other than one's own egocentric viewpoint. In order to achieve this
alignment, one must perform translations and rotations of one's own
egocentric reference frame. Processing translations and rotations of the
head and body is the core domain of the vestibular system, and thus is
of particular interest for research into mental spatial transformations.
Spatial transformations in 6 dimensions (translation along 3 axes, and
rotations around 3 axes) are highly complex, and non-linear, and we
propose that the brain adopts the strategy of simulating motion of the
body through space in order achieve a desired target position. This
effectively amounts to a running a simulation of dynamical system,
consisting of simulated limb movements and the resulting sensory
consequences. \autoref{fig:spt-schematic} (a) shows a simple scenario in
which a speaker wants to give an addressee, standing opposite, verbal
instructions about an action to perform on an external object. The
speaker must determine the most suitable action, given the addressee's
position relative to the external object, and in order to achieve this
goal, the speaker must align himself with the addressee, so that he can
use his own body-centred reference frame to quickly simulate the best
action. In this example, alignment with the addressee's reference frame
consists of a forwards translation and a rotation around the vertical
axis. This can be described as an imagined movement, in which the goal
is achieve a specific position.

\begin{figure}
\centering
\includegraphics[width=0.50000\textwidth]{diagrams/fig-1-common-process.pdf}
\caption{Perspective taking and imagined self-motion involve spatial
transformations of a representation of the self. (a) If a speaker's goal
is to take the perspective of an addressee, he must first solve the
inverse problem of determining the appropriate action, and then run a
forwards simulation in order to perform a translation and rotation of
his own egocentric reference frame. These operations can be viewed as
imagined movements.(b) Focussing on imagined rotation around the
vertical axis allows us to simplify the problem, and highlights the fact
that perspective taking and mental imagery involve common
computations.\label{fig:spt-schematic}}
\end{figure}

In order to do this, the speaker must first determine what type of
movement to perform, and the direction, velocity and duration that lead
to the desired outcome. This means that in order to know what to
simulate, the speaker must first work backwards from the desired
outcome, and infer the actions that are most likely to lead to that
outcome. This is known in the field of motor control as the inverse
problem (Wolpert \& Kawato, 1998). In order to achieve the desired
position, the motor commands can then be simulated, and the expected
sensory consequences can be estimated; this is known as the forward
problem. \xrefname{Figure}\cref{fig:spt-schematic} (b) shows that if the
goal is to simply simulate the perceptual experience of rotation around
the body's vertical (yaw) axis, rather than project a representation of
one's body onto a target, the problem is simplified; the actor can
perform forward inference without the need to determine the optimal
motor action. As demonstrated by Penny, Zeidman, \& Burgess (2013),
these computations can be performed in a common probabilistic model.
Interestingly, this has been discussed in the context of reinforcement
learning, and planning in humans has been described as reverse inference
(Botvinick \& Toussaint, 2012). Further, (Pickering \& Clark, 2014)
discuss the relationship between control-theoretic notions of
inverse/forward models, known collectively as internal models, and and
more general concepts of a generative model in a hierarchical Bayesian
framework.

Performing a simulation of body movements and estimating the sensory
consquences amounts to simulating the behaviour of a dynamical system;
this requires a generative model of the system. The aim of this paper is
to specifically investigate the generative model that may be used for
mental spatial transformations. The ability to run embodied simulations
is often claimed to be grounded in the relevant sensorimotor circuits,
and we claim that the vestibular system is ideally suited to investigate
this connection, because the generative models used by the vestibular
system for dynamic sensory inference have been extensively investigated
using approaches from optimal control theory, and more recently, dynamic
Baysian inference. The vestibular system faces a number of challenges in
interpreting the sensory measurements provided by the semi-circular
canals (SCC) and otolith organs. On the one hand, these signals are very
noisy, and ambiguous (MacNeilage, Ganesan, \& Angelaki, 2008); the
otoliths respond both to translations and tilt of the head relative to
gravity, and in order to disambiguate between translation and tilt, the
brain must perform sophisticated computations at the level of the brain
stem. It has become clear that the nature of these computations is
probabilistic, and that the brain makes use of strong prior beliefs in
order to interpret these sensory signals. A further challenge faced by
the vestibular system is that it needs to distinguish between sensory
signals that are the result of active motion, and those resulting from
external forces (Cullen, Brooks, \& Sadeghi, 2009, Cullen (2014)). The
ability to distinguishing re-afferent from ex-afferent sensory signals
is an essential capability of all organisms that move actively (Crapse
\& Sommer, 2008).

Furthermore, the apparent connection between mental imagery and
predictive processing has repeatedly been alluded to (Moulton \&
Kosslyn, 2009, Clark (2012), Gambi \& Pickering (2015)), and Grush
(2004) claims that mental imagery is performed by emulator circuits
implementing a forward model. These ideas can also be addressed from a
computational point of view in the context of dynamic Bayesian
inference, where there are multiple roles for sensory predictions, over
different time scales. This was recently discussed by (Tian \& Poeppel,
2012) in the context of imagined speech.

We proceed by introducing recent models of Bayesian inference for
real-time, dynamic sensory processing of vestibular data, based on
particle filters (Laurens \& Droulez, 2007, Karmali \& Merfeld (2012)),
and we reframe these as probabilistic graphical models. This allows us
to investigate the generative model separately from the algorithm used
for inference, and we then extend the basic model for sensory inference
in order to provide the ability to model the consequences of
self-initiated, active movements. Our aims are to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  identify various forms of predictive processing in the context of
  dynamic sensory processing.
\item
  identify various sources of interference between higher-level
  cognition and lower-level sensory processing.
\item
  provide suggestions for future research aiming to investigate the
  connections between cognitive and sensory processing.
\end{enumerate}

\section{Dynamic Bayesian models for sensory
inference}\label{dynamic-bayesian-models-for-sensory-inference}

\subsection{Previous work}\label{previous-work}

Whereas early models of vestibular sensory processing were inspired by
control theory (Borah, Young, \& Curry, 1988, Merfeld, Zupan, \& Peterka
(1999)), several recent studies have described vestibular sensory
processing as dynamic Bayesian inference (Laurens \& Droulez, 2007,
MacNeilage et al. (2008), Karmali \& Merfeld (2012)). Selva \& Oman
(2012) provide a recent review of the relationship between optimal
observer models and models which employ Kalman filtering, and show that
these are largely equivalent. Both Karmali \& Merfeld (2012) and Laurens
\& Droulez (2007) described vestibular sensory inference in terms of
particle filtering, which is a particular inference algorithm that can
be used to perform inference in state-space models. (Doucet, Godsill, \&
Andrieu, 2000, Doucet \& Johansen (2009)). Using their model, Laurens \&
Droulez (2007) were able to simulate perceptual responses to
centrifugation and off-vertical axis rotations.

A further well-known phenomenon of vestibular perception is velocity
storage, which is defined as the slower decay of perceptual and
oculomotor responses, relative to the SCC afferent responses (Raphan,
Matsuo, \& Cohen, 1979), \textbf{{[}Robinson 1977{]}}. This is generally
seen as evidence for the existence of an internal model in the
processing of vestibular signals. Karmali \& Merfeld (2012) focussed on
one-dimensional angular velocity during yaw rotations; their model was
able to explain the phenomenon of velocity storage as a consequence of
particle filtering, with the time constant of velocity storage depending
on the variability of the sensory afferent signals.

\subsection{Generative model for sensory
inference}\label{generative-model-for-sensory-inference}

We proceed by providing a brief introduction to the concept of
state-space models, and particle filtering, which is a particular
algorithm that can perform inference in state-space models. While Kalman
filters are restricted to linear Gaussian State-space models, particle
filtering can be applied to arbitrary nonlinear models. State-space
models can be seen as special cases of more general Bayesian
hierarchical models (Bishop, 2006; Murphy, 2012). Following the same
approach as Penny et al. (2013), we first describe the generative model
used for sensory inference as a dynamic probabilistic graphical model.

A state-space is special case of a dynamic latent variable model which
is suitable for modelling dynamical system in terms of latent,
unobservable processes and noisy observations. As an example, consider
the case where we want to model the velocity and position during a
leftward head turn. While the SCC measures angular acceleration, the
afferent signals are thought to be proportional to the angular velocity
of the head with a frequency band of approximately 0.1 -- 10 Hz, which
constitutes the range of natural head movements (Carriot, Jamali,
Chacron, \& Cullen, 2014; Grabherr, Nicoucar, Mast, \& Merfeld, 2008).
The measurements are observations of the true, unobservable angular
velocity; however, they are contaminted by Gaussian noise, which is
determined by characteristics of the sensory organs.
\autoref{fig:head-turn-data} shows a sequence of noisy measurements
during a two-second turn of the head. The acceleration generating the
movement is sinusoidal, given for each time step \(t\) by
\(Asin(2\pi t/T)\), where \(A\) is the amplitude of the movement,
\(T = 1/f\) is the duration of the movement, and \(f\) is the frequency.
This acceleration results in a peak velocity \(\omega_{peak} = AT/\pi\)
and a final angular position \(\theta_{final} = AT^2/2\pi\).

\begin{no-prefix-figure-caption}

\begin{figure}
\centering
\includegraphics{../generated-figures/head-turn-data-1.pdf}
\caption{\label{fig:head-turn-data}Sensory measurements obtained during a
short head turn. The goal is to infer the latent angular velocity and
position, based on the noisy measurements. \label{fig:head-turn}}
\end{figure}

\end{no-prefix-figure-caption}

The key idea is that we can write model the dynamics of a system by the
evolution of the latent state variables in discrete time steps as a
Markov process, where the state at each time point \(t\) depends only on
the previous state at time \(t-1\). This is referred to as the process
model. Estimating the velocity \(\omega\) and position \(\theta\) in
discrete time therefore requires writing the kinematic equations as a
set of first order difference equations.

\[\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t\]
\[\omega_{t} = \omega_{t-1}\]

Note that we have omitted the acceleration from the process model, for
the sake of simplicity; this amounts to assuming a constant velocity
model. These kinematic equations form a linear model, and could
optionally be written in matrix notation. For further simplicity, we
retain the current notation. Next, we consider the measurements at each
time step \(t\) as depending only on the current state. This simply
implements the idea that the sensory signal depends only on the current
head velocity. The above equations represent the expected values of the
state and observation variables. In order to complete the generative
model, we require a description of the stochastic nature dependancies.
We model both process and measurement noise a being normally
distributed, leading to the following linear Gaussian state-space model:

\[x_t = f(x_{t-1}, \epsilon)\] \[y_{t} = g(x_t, \delta)\]
\[\epsilon \sim N(0, \sigma^2_\epsilon) \]
\[\delta \sim N(0, \sigma^2_\delta) \]

where the state variable \(x_t\) consists of both the velocity
\(\omega_t\) and position \(\theta_t\), and \(y_{t}\) denotes the
sensory measurement.\footnote{Note that we although both Laurens \&
  Droulez (2007), Karmali \& Merfeld (2012) consider a realistic sensor
  model, for simplicity, we implement the sensor as being a noisy
  realisation of the velocity. Our model is not intended to be a
  realistic model of the SCC; rather, we focus on the probabilistic
  computations.} \(\epsilon\) and \(\delta\) are zero-mean Gaussian
noise terms with variances \(\sigma^2_\epsilon\) and
\(\sigma^2_\delta\). These refer to the process and measurement noise,
respectively. The process noise

\autoref{fig:graphical-model-inference} shows a representation of
state-space model as a graphical model, unrolled over multiple time
steps. Starting at time 0, the state variables evolve according to the
process model, with the measurements at each time step being generated
according to the measurement model. The 2-dimensiontal latent state
variable is unshaded, to indicate that it cannot be observed, wheres the
observed sensory measurements are shaded. Whereas previous models were
discussed in terms of the particular filtering algorithm used, the
graphical notation highlights the fact that a state-space model can be
viewed as a dynamic latent factor model (Blei, 2014). This opens up the
possibility of extending the graphical model to include higher-level
variables. After discussing particle filtering, an inference algorithm
often used to infer the values of the latent state variables, we will
introduce our extension to the model that will allow a prediction of the
consequences of active movements.

\begin{figure}
\includegraphics[width=1\linewidth]{diagrams/dynamic-belief-network-inference-1} \caption{Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.}\label{fig:graphical-model-inference}
\end{figure}

\subsection{Particle filtering}\label{particle-filtering}

We give a brief overview of particle filtering, concentrating on those
aspects most relevant for this study. For a more thorough treatment, see
Doucet et al. (2000) and Doucet \& Johansen (2009). Speekenbrink (2016)
provides a recent review with applications in psychological research.
Particle filtering is an algorithm for performing approximate Bayesian
inference in a graphical model. It proceeds by recursively computing the
distribution of the current state \(x_t\) as a a function of its parent
nodes, consisting in our example of the state at the previous time step.

The general procedure for recursive Bayesian inference is to perform
alternating prediction and updating steps.

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Predict step: At time t, the current state is predicted according to
  the process model. This prediction is maintained in the form of
  probability distributions over all the values.
\item
  Update step: a measurement is obtained, and this measurement is used
  in order to update the state estimate.
\end{enumerate}

By recursively estimating the current state, the observations can be
processed continuously; thus, the agent does not need to store the
entire history of observations -- only the previous belief state
estimate and the current observation are needed in order to estimate the
current state. At time step t, the agent observes the value of , and
this observation is used to update the previous belief:

\autoref{fig:particle-filter-explained} shows a diagram \ldots{}

\begin{figure}
\includegraphics[width=1\linewidth]{diagrams/figure-particle-filter-explained} \caption{Particle filter computations performed at every time step. The cloud of particles (grey circles) represents the current belief state. Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model. This is equivalent to sampling from the prior distribution. Then, during the updating step, an observation  is obtained (shown as a black diamond). Each particle is assigned an importance weight, according to how well it explains the observation. Each particle's diameter reflects its importance weight. Subsequently, a new set of particles is formed by weighted multinomial resampling, with each particle's probability of being chosen proportional to its importance weight. The particle cloud's mean and standard deviation are shown as light grey circles and lines. This resulting cloud of particles represents the posterior distribution. This posterior subsequently forms the prior distribution for the next time step.}\label{fig:particle-filter-explained}
\end{figure}

\autoref{fig:passive} shows the estimated velocity \ldots{}

\begin{no-prefix-figure-caption}

\begin{figure}
\centering
\includegraphics{../generated-figures/passive-1.pdf}
\caption{\label{fig:passive}Particle filter applied\ldots{}}
\end{figure}

\end{no-prefix-figure-caption}

\subsection{Summary}\label{summary}

\section{Extend previous models with control
input}\label{extend-previous-models-with-control-input}

For sensory inference of a passive disturbance, the algorithm can be run
with fixed parameters. Here we extend the basic particle filtering
(algorithm) model, in order to:

\begin{itemize}
\tightlist
\item
  \textbf{provide the capability to model the sensory consequences of
  self-initiated actions} (i.e.~predict/infer the consequences of one's
  own actions)
\item
  provide the capability to run a Monte Carlo simulation using the same
  generative model as is used for sensory inference
\end{itemize}

\subsection{Generative Model}\label{generative-model}

The generative model is a state-space model, which itself is a dynamic
model consisting of latent (state variables) and observed (measured)
variables.

\subsection{Inference Algorithm}\label{inference-algorithm}

Sequential Monte Carlo

\section{Modelling questions}\label{modelling-questions}

Experimental results to be explained: Wertheim, Mesland, \& Bles (2001)
report a cognitive suppression of otolith responses, based on
participants' prior beliefs about possible motions.

\section{An Investigation of Various
Priors}\label{an-investigation-of-various-priors}

Here, we identify the various forms of predictive processing in the
context of real-time sensory inference using SMC \#\# Related to
Filtering Algorithm \#\# Higher-level Priors control input
-\textgreater{} infer the consequences of one's own actions.

\section{Various sources of
interference}\label{various-sources-of-interference}

Here, we identify various sources of intereference (2-way) between
thought (higher-level cognition) and lower-level sensory processing

\subsection{Model mis-specification}\label{model-mis-specification}

If an observer uses a strong higher level prior, i.e.~the wrong
direction, this means that the data may be completely mis-interpreted.
(call this: the effect of inappropriate priors (Schwabe \& Blanke,
2008))

\subsection{Re-use of same generative
model}\label{re-use-of-same-generative-model}

This may lead to leakage, i.e.~inadvertently incorporating sensory
evidence

\subsection{Decision-making}\label{decision-making}

The interference may be explained in terms of higher-level perceptual
decision -making mechanisms \#\# evidence accumulation \#\# bias
(Starting point)

\subsection{Summary}\label{summary-1}

Interference may the result of either inappropriate priors, leading to a
mis-specified filtering algorithm, or biased decision-making. We should
be able to investigate which by applying cognitive process models that
explain response times and choices.

\section{Simulations}\label{simulations}

\section{General Discussion}\label{general-discussion}

The main result of this study is - place the emerging field of
vestibular cognition with a framewotk for dynamic probabilistic
inference and draw links to established models of sensory processing -
interpret result from vestibular cognition studies showing interference
between thought and perception as being the result of biased sensory
processing

\subsection{A comparison with previous
models}\label{a-comparison-with-previous-models}

Unfortunately, there aren't many, but see (Hiatt, Trafton, Harrison, \&
Schultz, 2004): imagined motion through space vs.~rotating contents of
visual buffer

\subsection{Support for our model:
Nigmatullina}\label{support-for-our-model-nigmatullina}

\subsection{Neuronal implementation}\label{neuronal-implementation}

(Lopez \& Blanke, 2011; Lopez, Blanke, \& Mast, 2012) (Klingner, Axer,
Brodoehl, \& Witte, 2016) (Eulenburg, Müller-Forell, \& Dieterich, 2012)

\subsubsection{Sensory gating}\label{sensory-gating}

(Gale et al., 2016)

\subsubsection{Head direction cells}\label{head-direction-cells}

Why dynamic simulation? Maybe in order to provide vestibular in put to
head direction cells (Taube, 2007)

\section{Summary and conclusions}\label{summary-and-conclusions}

\section{Acknowledgements}\label{acknowledgements}

\section{Motion Data}\label{motion-data}

\section{Tables}\label{tables}

\begin{longtable}[]{@{}rlcl@{}}
\caption{Demonstration of simple table syntax.}\tabularnewline
\toprule
Right & Left & Center & Default\tabularnewline
\midrule
\endfirsthead
\toprule
Right & Left & Center & Default\tabularnewline
\midrule
\endhead
12 & 12 & 12 & 12\tabularnewline
123 & 123 & 123 & 123\tabularnewline
1 & 1 & 1 & 1\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}clrl@{}}
\caption{Here's the caption. It, too, may span multiple
lines.}\tabularnewline
\toprule
\begin{minipage}[b]{0.15\columnwidth}\centering\strut
Centered Header\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Default Aligned\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\raggedleft\strut
Right Aligned\strut
\end{minipage} & \begin{minipage}[b]{0.31\columnwidth}\raggedright\strut
Left Aligned\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.15\columnwidth}\centering\strut
Centered Header\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Default Aligned\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\raggedleft\strut
Right Aligned\strut
\end{minipage} & \begin{minipage}[b]{0.31\columnwidth}\raggedright\strut
Left Aligned\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering\strut
First\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
row\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedleft\strut
12.0\strut
\end{minipage} & \begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
Example of a row that spans multiple lines.\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.15\columnwidth}\centering\strut
Second\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
row\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedleft\strut
5.0\strut
\end{minipage} & \begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
Here's another one. Note the blank line between rows.\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\newpage

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-Beveridge:2013gx}{}
Beveridge, M. E. L., \& Pickering, M. J. (2013). Perspective taking in
language: integrating the spatial and action domains. \emph{Frontiers in
Human Neuroscience}, \emph{7}.

\hypertarget{ref-Bishop:2006ui}{}
Bishop, C. M. (2006). \emph{Pattern Recognition and Machine Learning}.
Springer Verlag.

\hypertarget{ref-Blei:2014cp}{}
Blei, D. M. (2014). Build, Compute, Critique, Repeat: Data Analysis with
Latent Variable Models. \emph{Annual Review of Statistics and Its
Application}, \emph{1}(1), 203--232.

\hypertarget{ref-Borah:1988fd}{}
Borah, J., Young, L. R., \& Curry, R. E. (1988). Optimal Estimator Model
for Human Spatial Orientation. \emph{Annals of the New York Academy of
Sciences}, \emph{545}(1), 51--73.

\hypertarget{ref-Botvinick:2012fe}{}
Botvinick, M. M., \& Toussaint, M. (2012). Planning as inference.
\emph{Trends In Cognitive Sciences}, \emph{16}(10), 485--488.

\hypertarget{ref-Carriot:2014fv}{}
Carriot, J., Jamali, M., Chacron, M. J., \& Cullen, K. E. (2014).
Statistics of the vestibular input experienced during natural
self-motion: implications for neural processing. \emph{The Journal of
Neuroscience : The Official Journal of the Society for Neuroscience},
\emph{34}(24), 8347--8357.

\hypertarget{ref-Clark:2012vf}{}
Clark, A. (2012). Whatever Next? Predictive Brains, Situated Agents, and
the Future of Cognitive Science. \emph{Behavioral and Brain Sciences}.

\hypertarget{ref-Crapse:2008jc}{}
Crapse, T. B., \& Sommer, M. A. (2008). Corollary discharge across the
animal kingdom. \emph{Nature Reviews Neuroscience}, \emph{9}(8),
587--600.

\hypertarget{ref-CreemRegehr:2013kx}{}
Creem-Regehr, S. H., Gagnon, K. T., Geuss, M. N., \& Stefanucci, J. K.
(2013). Relating spatial perspective taking to the perception of other's
affordances: providing a foundation for predicting the future behavior
of others. \emph{Frontiers in Human Neuroscience}, \emph{7}, 596.

\hypertarget{ref-Cullen:2014gx}{}
Cullen, K. E. (2014). The neural encoding of self-generated and
externally applied movement: implications for the perception of
self-motion and spatial memory. \emph{Frontiers in Integrative
Neuroscience}, \emph{7}, 108.

\hypertarget{ref-Cullen:2009dc}{}
Cullen, K. E., Brooks, J. X., \& Sadeghi, S. G. (2009). How actions
alter sensory processing: reafference in the vestibular system.
\emph{Annals of the New York Academy of Sciences}, \emph{1164}, 29--36.

\hypertarget{ref-Deroualle:2014ho}{}
Deroualle, D., \& Lopez, C. (2014). Toward a vestibular contribution to
social cognition. \emph{Frontiers in Integrative Neuroscience},
\emph{8}.

\hypertarget{ref-Deroualle:2015gk}{}
Deroualle, D., Borel, L., Devèze, A., \& Lopez, C. (2015). Changing
perspective: The role of vestibular signals. \emph{Neuropsychologia},
\emph{79}, 175--185.

\hypertarget{ref-Doucet:2009us}{}
Doucet, A., \& Johansen, A. M. (2009). A tutorial on particle filtering
and smoothing: Fifteen years later. \emph{Handbook of Nonlinear
Filtering}.

\hypertarget{ref-Doucet:2000bh}{}
Doucet, A., Godsill, S., \& Andrieu, C. (2000). On sequential Monte
Carlo sampling methods for Bayesian filtering. \emph{Statistics and
Computing}, \emph{10}(3), 197--208.

\hypertarget{ref-zuEulenburg:2012bh}{}
Eulenburg, P. zu, Müller-Forell, W., \& Dieterich, M. (2012). On the
recall of vestibular sensations. \emph{Brain Structure and Function},
\emph{218}(1), 255--267.

\hypertarget{ref-Falconer:2012cl}{}
Falconer, C. J., \& Mast, F. W. (2012). Balancing the mind: vestibular
induced facilitation of egocentric mental transformations.
\emph{Experimental Psychology}, \emph{59}(6), 332--339.

\hypertarget{ref-Gale:2016cx}{}
Gale, S., Prsa, M., Schurger, A., Gay, A., Paillard, A., Herbelin, B.,
\ldots{} Blanke, O. (2016). Oscillatory neural responses evoked by
natural vestibular stimuli in humans. \emph{Journal of Neurophysiology},
\emph{115}(3), 1228--1242.

\hypertarget{ref-Gambi:2015gv}{}
Gambi, C., \& Pickering, M. J. (2015). Predicting and imagining
language. \emph{Language}.

\hypertarget{ref-Gardner:2016kd}{}
Gardner, M. R., Stent, C., Mohr, C., \& Golding, J. F. (2016). Embodied
perspective-taking indicated by selective disruption from aberrant self
motion. \emph{Psychological Research}, 1--10.

\hypertarget{ref-Grabherr:2008kk}{}
Grabherr, L., Nicoucar, K., Mast, F. W., \& Merfeld, D. (2008).
Vestibular thresholds for yaw rotation about an earth-vertical axis as a
function of frequency. \emph{Experimental Brain Research. Experimentelle
Hirnforschung. Expérimentation cérébrale}, \emph{186}(4), 677--681.

\hypertarget{ref-Grush:2004wx}{}
Grush, R. (2004). The emulation theory of representation: motor control,
imagery, and perception. \emph{The Behavioral and Brain Sciences},
\emph{27}(3), 377--96--discussion396--442.

\hypertarget{ref-Hiatt:2004tv}{}
Hiatt, L. M., Trafton, J. G., Harrison, A. M., \& Schultz, A. C. (2004).
A Cognitive Model for Spatial Perspective taking. In \emph{Proceedings
of the sixth international conference on cognitive modeling} (pp.
354--355).

\hypertarget{ref-Karmali:2012cv}{}
Karmali, F., \& Merfeld, D. (2012). A distributed, dynamic, parallel
computational model: the role of noise in velocity storage.
\emph{Journal of Neurophysiology}, \emph{108}(2), 390--405.

\hypertarget{ref-Klingner:2016ia}{}
Klingner, C. M., Axer, H., Brodoehl, S., \& Witte, O. W. (2016). Vertigo
and the processing of vestibular information: A review in the context of
predictive coding. \emph{Neuroscience and Biobehavioral Reviews},
\emph{71}, 379--387.

\hypertarget{ref-Laurens:2006be}{}
Laurens, J., \& Droulez, J. (2007). Bayesian processing of vestibular
information. \emph{Biological Cybernetics}, \emph{96}(4), 389--404.

\hypertarget{ref-Lenggenhager:2008et}{}
Lenggenhager, B., Lopez, C., \& Blanke, O. (2008). Influence of galvanic
vestibular stimulation on egocentric and object-based mental
transformations. \emph{Experimental Brain Research}, \emph{184}(2),
211--221.

\hypertarget{ref-Lopez:2011cc}{}
Lopez, C., \& Blanke, O. (2011). The thalamocortical vestibular system
in animals and humans. \emph{Brain Research Reviews}, \emph{67}(1-2),
119--146.

\hypertarget{ref-Lopez:2012ek}{}
Lopez, C., Blanke, O., \& Mast, F. W. (2012). The human vestibular
cortex revealed by coordinate-based activation likelihood estimation
meta-analysis. \emph{Neuroscience}, \emph{212}, 159--179.

\hypertarget{ref-PaulRMacNeilage:2008bv}{}
MacNeilage, P. R., Ganesan, N., \& Angelaki, D. E. (2008). Computational
approaches to spatial orientation: from transfer functions to dynamic
Bayesian inference. \emph{Journal of Neurophysiology}, \emph{100}(6),
2981--2996.

\hypertarget{ref-Merfeld:1999cg}{}
Merfeld, D., Zupan, L. H., \& Peterka, R. J. (1999). Humans use internal
models to estimate gravity and linear acceleration. \emph{Nature},
\emph{398}(6728), 615--618.

\hypertarget{ref-Moulton:2009ij}{}
Moulton, S. T., \& Kosslyn, S. M. (2009). Imagining predictions: mental
imagery as mental emulation. \emph{Philosophical Transactions of the
Royal Society of London Series B, Biological Sciences},
\emph{364}(1521), 1273--1280.

\hypertarget{ref-Murphy:2012ua}{}
Murphy, K. P. (2012). \emph{Machine Learning}. MIT Press.

\hypertarget{ref-Penny:2013iv}{}
Penny, W. D., Zeidman, P., \& Burgess, N. (2013). Forward and backward
inference in spatial cognition. \emph{PLoS Computational Biology},
\emph{9}(12), e1003383.

\hypertarget{ref-Pezzulo:2013iy}{}
Pezzulo, G., Iodice, P., Ferraina, S., \& Kessler, K. (2013). Shared
action spaces: a basis function framework for social re-calibration of
sensorimotor representations supporting joint action. \emph{Frontiers in
Human Neuroscience}, \emph{7}, 800.

\hypertarget{ref-Pickering:2014bo}{}
Pickering, M. J., \& Clark, A. (2014). Getting ahead: forward models and
their place in cognitive architecture. \emph{Trends In Cognitive
Sciences}, \emph{18}(9), 451--456.

\hypertarget{ref-Raphan:1979hn}{}
Raphan, T., Matsuo, V., \& Cohen, B. (1979). Velocity storage in the
vestibulo-ocular reflex arc (VOR). \emph{Experimental Brain Research}.

\hypertarget{ref-Schwabe:2008ho}{}
Schwabe, L., \& Blanke, O. (2008). The vestibular component in
out-of-body experiences: a computational approach. \emph{Frontiers in
Human Neuroscience}, \emph{2}.

\hypertarget{ref-Selva:2012gsa}{}
Selva, P., \& Oman, C. M. (2012). Relationships between observer and
Kalman Filter models for human dynamic spatial orientation.
\emph{Journal of Vestibular Research : Equilibrium \& Orientation},
\emph{22}(2), 69--80.

\hypertarget{ref-Speekenbrink:2016kc}{}
Speekenbrink, M. (2016). A tutorial on particle filters. \emph{Journal
of Mathematical Psychology}.

\hypertarget{ref-Taube:2007fk}{}
Taube, J. S. (2007). The head direction signal: origins and
sensory-motor integration. \emph{Annual Review of Neuroscience},
\emph{30}, 181--207.

\hypertarget{ref-Tian:2012ui}{}
Tian, X., \& Poeppel, D. (2012). Mental imagery of speech: linking motor
and perceptual systems through internal simulation and estimation.
\emph{Frontiers in Human Neuroscience}.

\hypertarget{ref-Wertheim:2001cb}{}
Wertheim, A. H., Mesland, B. S., \& Bles, W. (2001). Cognitive
suppression of tilt sensations during linear horizontal self-motion in
the dark. \emph{Perception}, \emph{30}(6), 733--741.

\hypertarget{ref-Wolpert:1998td}{}
Wolpert, D. M., \& Kawato, M. (1998). Multiple paired forward and
inverse models for motor control. \emph{Neural Networks : The Official
Journal of the International Neural Network Society}, \emph{11}(7-8),
1317--1329.

\clearpage
\renewcommand{\listtablename}{Table captions}
\listoftables

\clearpage
\renewcommand{\listfigurename}{Figure captions}
\listoffigures

\end{document}

### 2017-01-25 manuscript.Rmd
new content, new diagrams

@@
 ---
+# title             : "Modeling Cognitive Effects Vestibular Sensory Inference: From Sensory Inference To Imagined Self-Motion"
+
 title             : "Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion"
+
 # title             : "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
@@
 
-  AE and FM planned and conducted the research; AE and FM analyzed the data and wrote the manuscript.
+  AE and FM planned and conducted the research; AE and FM performed the simulations and wrote the manuscript.
 
@@
 footnotelist      : no
-number_sections   : FALSE
+number_sections   : no
 lineno            : yes
@@
 output:
+     # papaja::apa6_word
      papaja::apa6_pdf:
         pandoc_args:
-        - --filter
-        - pandoc-fignos
         fig_caption: true
-fignos-cleveref: On
-fignos-plus-name: Figure
-header-includes: \usepackage{caption}
 ---
 
-```{r preliminary, message = FALSE, warning = FALSE}
+```{r preliminary, cache=FALSE, message = FALSE, warning = FALSE}
 library(papaja)
@@
 (a) If a speaker's goal is to take the perspective of an addressee, he must first solve the inverse problem of determining the appropriate action, and then run a forwards simulation in order to perform a translation and rotation of his own egocentric reference frame.
-These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.pdf){#fig:spt-schematic width=50%}
+These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.pdf){#fig:spt-schematic width=75%}
 
@@
 In order to achieve the desired position, the motor commands can then be simulated, and the expected sensory consequences can be estimated; this is known as the forward problem.
-@fig:spt-schematic (b) shows that if the goal is to simply simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action.
+\autoref{fig:spt-schematic} (b) shows that if the goal is to simply simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action.
 As demonstrated by @Penny:2013iv, these computations can be performed in a common probabilistic model.
@@
 
+The putative link between predictive processing and higher cognitive processing has a firm basis in predictive processing accounts of sensory inference [Friston].
+The predictive processing framework has been applied to almost all sensory modalities [CITATIONS: pain, proprioception, auditory , visual, etc.], and the picture that has emerged in recent years is that the brain employs sophisticated inference algorithms in order to "interpret" the stream of incoming sensory data.
+The brain seems to be actively engaged in formulating hypotheses, and testing these on sensory data; this has been extended to the concept of active inference [Friston], in which an agent attempts to either change it's internal model of world, or to interven on the world in order to fulfil sensory predictions.
+We claim that the vestibular modality is particularly well-suited to investigating predictive processing at multiple hierarchical scales, from the lowest levels of sensory inference to higher levels, where goals, intentions and beliefs of the organism become relevant.
+A case in point is the study by @Wertheim:2001cb, who provided evidence that participants' perceptions of tilt induced by vestibular stimulation are shaped to a large extent by the beliefs about possible motions they formed prior to taking part in the experiement.
+
+Interference effects between higher-level cognitive processes, such as SPT and mental imagery, and sensory inference, are thought to occur as a result of shared circuitry [@Deroualle:2014ho, @Nigmatullina:2015gh].
+@Ellis:2017fg recently claimed that these types of interference effects might arise as a result of cognitive processing exerting an influence on circuits performing sensory inference via shared higher-level priors.
+In this paper, we pursue this claim, and we attempt to the grounding of cognitive abilities in the computations performed by low-level sensory circuits in the service of dynamic real-time interactions with the world.
+
 We proceed by introducing recent models of Bayesian inference for real-time, dynamic sensory processing of vestibular data, based on particle filters [@Laurens:2006be, @Karmali:2012cv], and we reframe these as probabilistic graphical models.
-This allows us to investigate the generative model separately from the algorithm used for inference, and we then extend the basic model for sensory inference in order to provide the ability to model the consequences of self-initiated, active movements.
+This allows us to investigate the generative model separately from the algorithm used for inference, and we then extend the basic model for sensory inference in order to focus specifically on the ability to model the consequences of self-initiated, active movements.
+This is an implementation of the idea that information about motor commands are projected to sensory areas by means of an efference copy, based on which a forward model then computes the sensory consequences.
 Our aims are to:
@@
 Using their model, @Laurens:2006be were able to simulate perceptual responses to centrifugation and off-vertical axis rotations.
-
-A further well-known phenomenon of vestibular perception is velocity storage, which is defined as the slower decay of perceptual and oculomotor responses, relative to the SCC afferent responses [@Raphan:1979hn], __[Robinson 1977]__.
+A well-known phenomenon of vestibular perception is velocity storage, which is defined as the slower decay of perceptual and oculomotor responses, relative to the SCC afferent responses [@Raphan:1979hn], __[Robinson 1977]__.
 This is generally seen as evidence for the existence of an internal model in the processing of vestibular signals.
@@
 ## Generative model for sensory inference
-We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can perform inference in state-space models.
-While Kalman filters are restricted to linear Gaussian State-space models, particle filtering can be applied to arbitrary nonlinear models.
+We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can be used to perform inference in state-space models.
+While Kalman filters provide the optimal solution when the restrictions of linear Gaussian State-space models are met, they are not easily extensible to non-linear models, which is relevant when considering multi-dimensional motion.
+Particle filtering can be applied to arbitrary nonlinear models.
 State-space models can be seen as special cases of more general Bayesian hierarchical models [@Bishop:2006ui; @Murphy:2012ua].
-Following the same approach as @Penny:2013iv, we first describe the generative model used for sensory inference as a dynamic probabilistic graphical model.
-
-A state-space is special case of a dynamic latent variable model which is suitable for modelling dynamical system in terms of latent, unobservable processes and noisy observations.
-As an example, consider the case where we want to model the velocity and position during a leftward head turn.
-While the SCC measures angular acceleration, the afferent signals are thought to be proportional to the angular velocity of the head with a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv].
-The measurements are observations of the true, unobservable angular velocity; however, they are contaminted by Gaussian noise, which is determined by characteristics of the sensory organs.
+Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model
+ <!-- from a statistical perspective, rather an engineering perspective -->
+A state-space model is special case of a dynamic latent variable model which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations.
+As an example, consider trying to estimate the velocity and position during a leftward head turn.
+The afferent SCC signals are considered to be proportional to the angular velocity of the head with a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv].
+The measurements are therefore observations of the true, unobservable angular velocity; however, they are contaminted by Gaussian noise, which is determined by characteristics of the sensory organs.
 \autoref{fig:head-turn-data} shows a sequence of noisy measurements during a two-second turn of the head.
@@
 This acceleration results in a peak velocity $\omega_{peak} = AT/\pi$ and a final angular position $\theta_{final} = AT^2/2\pi$.
+Due to the inherent noisiness of the measurements, and the dynamic nature of the stimulus, the measurements need to be interpreted according to the context in which they occur.
 
-```{r head-turn-data, echo=FALSE, fig.cap= "Sensory measurements obtained during a short head turn.
+```{r head-turn-data, echo=FALSE, fig.cap= "Sensory measurements obtained during a short leftward head turn.
 The goal is to infer the latent angular velocity and position, based on the noisy measurements.
@@
 
-The key idea is that we can write model the dynamics of a system by the evolution of the latent state variables in discrete time steps as a Markov process, where the state at each time point $t$ depends only on the previous state at time $t-1$.
+The key idea is that we can model the dynamics of a system by the evolution of the latent state variables, representing the velocity and position, as a first-order Markov process, where the state at each time point $t$ depends only on the previous state at time $t-1$.
 This is referred to as the process model.
-Estimating the velocity $\omega$ and position $\theta$ in discrete time therefore requires writing the kinematic equations as a set of first order difference equations.
+Estimating the velocity $\omega$ and position $\theta$ in discrete time requires writing the kinematic equations as a set of first order difference equations.
 
@@
 
-Note that we have omitted the acceleration from the process model, for the sake of simplicity; this amounts to assuming a constant velocity model.
+Note that we have omitted the acceleration from the process model, for the sake of simplicity; this amounts to assuming a constant velocity model, which just means that we expect the current velocity to be same as the previous velocity.
 These kinematic equations form a linear model, and could optionally be written in matrix notation.
@@
 These refer to the process and measurement noise, respectively.
-The process noise
+The process noise represents the uncertainty in the dependance of the current state on the previous state; under the constant velocity assumption, the process noise must be large enough if the model is to allow for change in velocity due to unmodelled acceleration.
+Note that the measurement noise is a model of the noise characteristics of the sensor, and not a physical property of the sensor itself.
+The function $f$ and $g$ represent the process transition function and the measurement function, respectively.
+In the case of a linear dynamical system, the transition function can be written as:
+$$x_t \sim N(Ax_{t-1}), \epsilon^2)$$,
+where the matrix $A$ replaces the transition function $f$.
 
 [^sensormodel]: Note that we although both @Laurens:2006be, @Karmali:2012cv consider a realistic sensor model, for simplicity, we implement the sensor as being a noisy realisation of the velocity.
-Our model is not intended to be a realistic model of the SCC; rather, we focus on the probabilistic computations.
+Our model is not intended to be a realistic model of the SCC; rather, we focus on the probabilistic computations relevant for higher-level cognition.
 
 <!-- \autoref{fig:graphical-model-inference} -->
-\autoref{fig:graphical-model-inference} shows a representation of state-space model as a graphical model, unrolled over multiple time steps.
-Starting at time 0, the state variables evolve according to the process model, with the measurements at each time step being generated according to the measurement model.
+\autoref{fig:graphical-model-inference} shows a representation of a state-space model as a graphical model, unrolled over multiple time steps.
+Starting at time 0, the state evolves according to the process model, with the measurements at each time step being generated according to the measurement model.
 The 2-dimensiontal latent state variable is unshaded, to indicate that it cannot be observed, wheres the observed sensory measurements are shaded.
@@
 This opens up the possibility of extending the graphical model to include higher-level variables.
-After discussing particle filtering, an inference algorithm often used to infer the values of the latent state variables, we will introduce our extension to the model that will allow a prediction of the consequences of active movements.
+Given a generative model, we can apply various sampling algorithms, either in order to run a Monte Carlo simulation by sampling from the prior without incorporating any observed variables, or to perform inference, by conditioning on observed variables.
+After a brief discussion particle filtering, an inference algorithm often used to infer the values of the latent state variables, we will introduce our extension to the model that will allow a prediction of the consequences of active movements.
 
@@
 
-```{r graphical-model-inference, fig.cap="Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.", echo=FALSE, out.width='100%'}
+```{r graphical-model-inference, fig.cap="Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.", echo=FALSE}
 knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.pdf')
@@
 
-## Particle filtering
-We give a brief overview of particle filtering, concentrating on those aspects most relevant for this study.
+## Bayesian inference using particle filtering
+
+We give a brief overview of particle filtering, an sequential Monte Carlo algorithm for performing approximate Bayesian inference in a graphical model.
+We descibe a simple bootstrap particle filter, concentrating on those aspects most relevant for this study.
 For a more thorough treatment, see @Doucet:2000bh and @Doucet:2009us.
 @Speekenbrink:2016kc provides a recent review with applications in psychological research.
-Particle filtering is an algorithm for performing approximate Bayesian inference in a graphical model.
-It proceeds by recursively computing the distribution of the current state $x_t$ as a a function of its parent nodes, consisting in our example of the state at the previous time step.
+Particle filtering is .
+It proceeds by recursively estimating the distribution of the current state, $P(x_t | y_{1:t})$, as a function of its parent nodes, conditional upon the sequence of measurements $y_{1:t}$.
+The distributions are represented by a collection of discrete weighted samples, known as particles.
 
-The general procedure for recursive Bayesian inference is to perform alternating prediction and updating steps.
+The general procedure for recursive Bayesian inference is to start from an initial prior distribution, and the to perform alternating prediction and updating steps.
+In other words: first, a hypothesis about the expected state is generated, and then this hypothesis is tested by evaluating how well it predicts the data.
+The algorithm starts with a prior distribution of the particles, and at each subsequent time step, proceeds by first sampling from the so-called proposal distribution $P(x_t | y_{1:t-1})$, which is based on the specific choice of transition function, and therefore reflects the probabilistic causal model of the sensory measurements.
+Next, during the updating step, the current measurement is incorporated by evaluating the likelihood, $P(y_t | x_t)$, which gives the probability of observing the measurement for each individual particle.
+The particle are then assigned an importance weight, based on how well they predicted the measurement.
+A common problem in  sequential Monte Carlo methods is that the particle weights tend to become very small over time; this is known as the degeneracy problem.
+In order to counteract this, multinomial resampling is performed when effective sample size falls under a threshold.
+A new set of particles is then chosen by weighted sampling with replacement, which each particle’s probability of being selected given by its importance weight.
+Resampling ensures that only particles with a relatively large probability of ‘explaining’ the observations are maintained.
+This introduces flexibility in the algorithm’s method for incorporating evidence into its beliefs about the latent states.
 
-1)	Predict step: At time t, the current state is predicted according to the process model. This prediction is maintained in the form of probability distributions over all the values.
+\autoref{fig:particle-filter-explained} shows a visualization of the computations performed by a particle filter at each time step, in order to track the angular velocity.
+Initially, the particles are propagated forwards in time according to the proposal distribution.
+They are equally weighted, indicated by their diameters.
+During the update step, the particle are reweighted, with each particle’s weights reflecting how well it explains the measurement.
+This is visualized by the change in diameter; those particles lying close to the observation (sensory input from the horizontal canal) have a larger diameter.
+After resampling, the new set of particles is pulled towards the measurement, and the weights are reset.
+This adjustment toward the observation depends on both the process noise, and the modelled sensor noise.
+The uncertainty of the belief, as indicated by the particle cloud's standard deviation (light grey lines), increases at every prediction step, and subsequently decreases after updating.
+This cycle of prediction, updating and resampling ensures that the collection of particles comes to represent the 'true' filtering distribution, by eliminating those particles with low weights.
+The particle filter represents all beliefs in the form of probability distributions; the mean and variance of the belief distribution can be computed at each time step, and the logarithm of the weights $log(\frac{1}{N}\sum^N_{i=1}w_i)$, averaged over all $N$ particles, may be summed over all time steps in order to compute the log-likelihood $\sum^T_{t=1}\Big[log(\sum^N_{i=1}w_i) - log(N)\Big]$, which can be used in order to evaluate how well the model was able to fit the sensory evidence.
+The summed log-likelihood can be used for parameter estimation in extensions of the basic particle filtering algorithm [@Andrieu:2010gc].
 
-2)	Update step: a measurement is obtained, and this measurement is used in order to update the state estimate.
+<!-- By recursively estimating the current state, the observations can be processed continuously; thus, the agent does not need to store the entire history of observations – only the previous belief state estimate and the current observation are needed in order to estimate the current state.
+At time step t, the agent observes the value of , and this observation is used to update the previous belief: -->
 
-By recursively estimating the current state, the observations can be processed continuously; thus, the agent does not need to store the entire history of observations – only the previous belief state estimate and the current observation are needed in order to estimate the current state.
-At time step t, the agent observes the value of , and this observation is used to update the previous belief:
-
-\autoref{fig:particle-filter-explained} shows a diagram ...
-
-![Particle filter computations performed at every time step.
+<!-- ![Particle filter computations performed at every time step.
 The cloud of particles (grey circles) represents the current belief state.
@@
 This resulting cloud of particles represents the posterior distribution.
-This posterior subsequently forms the prior distribution for the next time step.](diagrams/figure-particle-filter-explained.pdf){#fig:particle-filter-explained width=100%}
+This posterior subsequently forms the prior distribution for the next time step.](diagrams/figure-particle-filter-explained.pdf){#fig:particle-filter-explained width=100%} -->
 
-\autoref{fig:passive} shows the estimated velocity ...
+```{r particle-filter-explained, fig.cap="Particle filter computations performed at every time step.
+The cloud of particles (grey circles) represents the current belief state.
+Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model.
+This is equivalent to sampling from the prior distribution.
+Then, during the updating step, an observation  is obtained (shown as a black diamond).
+Each particle is assigned an importance weight, according to how well it explains the observation.
+Each particle's diameter reflects its importance weight.
+Subsequently, a new set of particles is formed by weighted multinomial resampling, with each particle's probability of being chosen proportional to its importance weight.
+The particle cloud's mean and standard deviation are shown as light grey circles and lines.
+This resulting cloud of particles represents the posterior distribution.
+This posterior subsequently forms the prior distribution for the next time step.", echo=FALSE}
+knitr::include_graphics('diagrams/figure-particle-filter-explained.pdf')
+```
 
-```{r passive, fig.cap="Particle filter applied...", echo=FALSE, message=FALSE, warning=FALSE}
+\autoref{fig:passive} shows the result of applying the particle filtering algorithm to the graphical model depicted in \autoref{fig:graphical-model-inference}, in order to infer the velocity and position during a two-second leftward head turn.
+
+```{r passive, cache=TRUE, fig.cap="Particle filter for inferring velocity and position during a two-second head turn.", echo=FALSE, message=FALSE, warning=FALSE}
 params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
@@
 
+The measurement are shown as squares, and the true velocity that generated the data is shown as a dotted line.
+This velocity is the result of an external force being applied.
+The solid lines depict the estimated velocity (left) and position (right), and the shaded areas show the $95\%$ credible intervals.
+The model is able to estimate the latent state despite making the assumption that the velocity is constant.
+It does this by allowing for random changes in velocity by letting the process noise be sufficiently large.
+Note that both velcoity and position estimates lag the true values; this is due the fact that the acceleration is not modelled explicitely, and thus the system is slow to react to changes in velocity.
+
+This graphical model assumes that the velocity is constant, and does not  the explicitly estimate acceleration; however, it allows for random changes in its velocity by in terms of the process noise.
+The estimation can be improved by increasing the process noise.
+\autoref{fig:passive-2} shows the resulting estimates.
+Whilst the lag is somewhat reduced, a consequence of increasing the process noise is that the measurement are given more weight, resulting in a tendency to overfit the noise in the measurements.
+The constant velocity (random acceleration) model is unable to capture the underlying smooth trajectory of the velocity and position, instead showing random jumps in the velocity estimate.
+
+```{r passive-2, cache=TRUE, fig.cap="Particle filter for inferring velocity and position during a two-second head turn with increased process noise.", echo=FALSE, message=FALSE, warning=FALSE}
+params <- list(sdx = c(1.8, 1.8) * eye(2), sdy = 2.0,
+                x_init = c(0, 0),
+                sdx_init = 0.5*eye(2),
+                A = 20, N = 1000,
+                dt = 0.1, freq = 0.5,
+                transfun = transfun)
+
+passive_2 <- run_particle_filter(motiondata = motiondata, params = params,
+                            resample_particles = TRUE, rs_thresh = 0.5,
+                            selfmotion = "passive")
+
+plot(passive_2, panel_label = "", print_loglik = FALSE, predict = FALSE, imagine = FALSE)
+```
+
 ```{r eval=FALSE, include=FALSE}
@@
 
-## Summary
+In order to imporove the model, the acceleration can be included in the latent state, resulting in a 3-dimensional latent state variable:
 
-# Extend previous models with control input
+$$\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2}\Delta t^2$$
 
-For sensory inference of a passive disturbance, the algorithm can be run with fixed parameters. Here we extend the basic particle filtering (algorithm) model, in order to:
+$$\omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t$$
+The effect that this has in terms of the inference algorithm is that the proposal distribution $P(x_t | y_{1:t})$, from which samples are drawn from during the 'predict' step, is improved.
+By including the acceleration in the latent state, the the changes in velocity are modelled.
+Of course, the acceleration itself is not modelled, and the changes in acceleration (jerk) are thus treated as random, resuling in a random jerk model.
 
-- __provide the capability to model the sensory consequences of self-initiated actions__ (i.e. predict/infer the consequences of one's own actions)
-- provide the capability to run a Monte Carlo simulation using the same generative model as is used for sensory inference
+> QUESTION: shall we show a model that includes acceleration in its state?
 
-## Generative Model
-The generative model is a state-space model, which itself is a dynamic model consisting of latent (state variables) and observed (measured) variables.
+<!-- ## Performance for various types of natural head movements -->
 
-## Inference Algorithm 
-Sequential Monte Carlo
+```{r shaking-head, eval=FALSE, fig.cap="Particle filter applied to...", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
+dt <- 0.1
+head_shaking_data <- generate_data(T = 2, frequency = 2.0,
+                             amplitude = 20,
+                             sensor_sd = 2.0)
+
+# plot_trajectories(motiondata = abrupt_stop_data)
+params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
+                x_init = c(0, 0),
+                sdx_init = 0.5*eye(2),
+                A = 20, N = 1000,
+                dt = 0.1, freq = 0.5,
+                transfun = transfun)
+
+head_shaking <- run_particle_filter(motiondata = head_shaking_data, 
+                                    params = params,
+                                    selfmotion = "passive")
+
+plot(head_shaking, panel_label = "", print_loglik = FALSE, predict = FALSE, imagine = FALSE)
+```
+
+```{r two-external-forces, eval=FALSE, fig.cap="Particle filter applied to...", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
+dt <- 0.1
+ext_data_1 <- generate_data(T = 4, frequency = 0.7,
+                             amplitude = 20,
+                             sensor_sd = 2.0)
+ext_data_2 <- generate_data(T = 4, frequency = 0.5,
+                             amplitude = -5,
+                             sensor_sd = 2.0)
+superposition_data <- ext_data_1 + ext_data_2
+
+# plot_trajectories(motiondata = abrupt_stop_data)
+params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
+                x_init = c(0, 0),
+                sdx_init = 0.5*eye(2),
+                A = 20, N = 1000,
+                dt = 0.1, freq = 0.5,
+                transfun = transfun)
+
+superposition <- run_particle_filter(motiondata = superposition_data, 
+                                    params = params,
+                                    selfmotion = "passive")
+
+plot(superposition, panel_label = "", print_loglik = FALSE, predict = FALSE, imagine = FALSE)
+```
+
+# Extension allowing the prediction of active movements
+While the model described above can be used in order to track the behaviour of a dynamic system by including the acceleration as a latent state, it does not include any means of incorporating an active prediction of the kinematics.
+During sensory inference the measurements are incorporated, and and a posterior estimate of the latent state is obtained by updating the prior with the likelihood.
+The underlying kinematic model thus relies on the sensory data to inform its estimates.
+While this model is perfectly suitable for sensory inference, for the task of predicting the sensory consequences of self-initiated movemenents, the model is inadequate, as it does not allow including further knowledge about kinematics, which can be used to improve the proposal distribution.
+This knowledge is obtained by means of an efference copy of executed or planned motor commands, which is sent to a forward model.
+The forward model must then translate the information from the efference copy in a format that is suitable for the sensory system.
+
+In terms of state-space models, this is implemented as a control input $u$, which extends the graphical model shown in \autoref{fig:graphical-model-inference}.
+In this model, the estimated acceleration is replaced by an known acceleration signal.
+This capability to predict the consequences of self-initiated movements is essential in order to distinguish re-afference from ex-afference - the system has knowledge about how it moves, and can use this to predict the sensory consequences.
+A further benefit of including a control input is that the model can now be used in order to run counterfactual simulations of the dynamic behaviour of the system for various types of movements.
+
+#TODO: model implementation
+
+\autoref{fig:graphical-model-control} shows the resulting graphical model.
+
+```{r graphical-model-control, fig.cap="Graphical model extended with hierarchical model for self-initiated movements.", echo=FALSE}
+knitr::include_graphics('diagrams/dynamic-belief-network-inference-2.pdf')
+```
 
@@
 Here, we identify the various forms of predictive processing in the context of real-time sensory inference using SMC
-## Related to Filtering Algorithm
-## Higher-level Priors
+
+- Related to Filtering Algorithm
+- Higher-level Priors: The process model determines the kind of motion that is expected. Call this _hypothesis-driven_ sensory inference. It implements an expectation and is a strong prior.
+
+THe brain implements multiple priors:
+
+- ability to select type of motion
+- parameters of tht motion
+
 control input -> infer the consequences of one's own actions.
@@
 # Various sources of interference
-Here, we identify various sources of intereference (2-way) between thought (higher-level cognition) and lower-level sensory processing
+Here, we identify various sources of intereference (2-way) between thought (higher-level cognition) and lower-level sensory processing.
+This interference can be either due to low-level influences in sensory inference, due to priors on higher-level parameters, choice of process model (false expectations of certain type of motion).
+This is shown by model simulations.
 
-## Model mis-specification
-If an observer uses a strong higher level prior, i.e. the wrong direction, this means that the data may be completely mis-interpreted. (call this: the effect of inappropriate priors [@Schwabe:2008ho])
+- Model mis-specification: If an observer uses a strong higher level prior, i.e.
+the wrong direction, this means that the data may be completely mis-interpreted.
+(call this: the effect of inappropriate priors [@Schwabe:2008ho])
 
-## Re-use of same generative model
-This may lead to leakage, i.e. inadvertently incorporating sensory evidence
+- Re-use of same generative model: This may lead to leakage, i.e. inadvertently incorporating sensory evidence
 
-## Decision-making
-The interference may be explained in terms of higher-level perceptual decision -making mechanisms
-## evidence accumulation
-## bias (Starting point)
+- Decision-making: The interference may be explained in terms of higher-level perceptual decision-making mechanisms
+    - evidence accumulation
+    - bias (Starting point)
 
@@
 
-The main result of this study is 
-- place the emerging field of vestibular cognition with a framewotk for dynamic probabilistic inference and draw links to established models of sensory processing
-- interpret result from vestibular cognition studies showing interference between thought and perception as being the result of biased sensory processing
+The main results of this study are:
+
+- place the emerging field of vestibular cognition with a framewokk for dynamic probabilistic inference and draw links to established models of sensory processing
+
+- interpret result from vestibular cognition studies showing interference between thought and perception as being the result of biased sensory processing.
+This interference can be either due to low-level influences in sensory inference, due to priors on higher-level parameters, choice of process model (false expectations of certain type of motion).
+This is shown by model simulations.
+
+- A further possibility for interference is a bias introduced in the decision-making process (discuss potential model of decision-making based on sequential Bayes factor and contrast with [@Merfeld:2016il].)
 
@@
 
-## Support for our model: Nigmatullina
+## What is the difference between predicting and imagining?
+
+Imagining can be described as counterfactual inference [@Krishnan:2015wr].
+Imagined motion (for planning, decision-making, perspective transformations) is performed under counterfactual assumptions.
+Predictions/anticipation is performed under factual assumptions.
+
+## Support for our model:
+
+- Nigmatullina
+- Wertheim
 
@@
 
-# Motion Data
-
-```{r generate-data-1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-dt <- 0.1
-motiondata <-  generate_data(T = 2, frequency = 0.5,
-                             amplitude = 20,
-                             sensor_sd = 2.0)
-# plot_trajectories(motiondata = motiondata)
-
-motiondata_hidden <- motiondata %>%
-    gather(type, value = value, -time) %>%
-    filter(type != "observations") %>%
-    mutate(type = factor(type,
-                         levels = c("position",
-                                    "velocity",
-                                    "acceleration")))
-
-p1 <- motiondata_hidden %>%
-        ggplot(aes(x = time, y = value)) +
-            geom_line(aes(linetype = type), size = 1.4) +
-            geom_point(data = motiondata,
-                       aes(x = time, y = observations),
-                       shape = 21, colour = "white",
-                       fill = "white", size = 6, stroke = 2) +
-            geom_point(data = motiondata,
-                       aes(x = time, y = observations),
-                       shape = 21, colour = "black",
-                       fill = "white", size = 4, stroke = 2) +
-            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
-            xlab("Time (sec)") +
-            ylab("Angular position (deg)") +
-            theme_apa() +
-            theme(legend.title=element_blank())
-
-motiondata_2 <-  generate_data(T = 2, frequency = 0.5,
-                             amplitude = -20,
-                             sensor_sd = 1.8)
-# plot_trajectories(motiondata = motiondata)
-
-motiondata_hidden_2 <- motiondata_2 %>%
-    gather(type, value = value, -time) %>%
-    filter(type != "observations") %>%
-    mutate(type = factor(type,
-                         levels = c("position",
-                                    "velocity",
-                                    "acceleration")))
-
-p2 <- motiondata_hidden_2 %>%
-        ggplot(aes(x = time, y = value)) +
-            geom_line(aes(linetype = type), size = 1.4) +
-            geom_point(data = motiondata_2,
-                       aes(x = time, y = observations),
-                       shape = 21, colour = "white",
-                       fill = "white", size = 6, stroke = 2) +
-            geom_point(data = motiondata_2,
-                       aes(x = time, y = observations),
-                       shape = 21, colour = "black",
-                       fill = "white", size = 4, stroke = 2) +
-            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
-            xlab("Time (sec)") +
-            ylab("Angular position (deg)") +
-            theme_apa() +
-            theme(legend.title=element_blank())
-
-cowplot::plot_grid(p1, p2, nrow = 2, labels = c("a", "b"))
-```
-
-# Tables
-
-  Right     Left     Center     Default
--------     ------ ----------   -------
-     12     12        12            12
-    123     123       123          123
-      1     1          1             1
-
-Table:  Demonstration of simple table syntax.
-
--------------------------------------------------------------
- Centered   Default           Right Left
-  Header    Aligned         Aligned Aligned
------------ ------- --------------- -------------------------
-   First    row                12.0 Example of a row that
-                                    spans multiple lines.
-
-  Second    row                 5.0 Here's another one. Note
-                                    the blank line between
-                                    rows.
--------------------------------------------------------------
-
-Table: Here's the caption. It, too, may span
-multiple lines.
-
 \newpage

### 2017-01-25 notes.Rmd
new content, new diagrams

@@
 Sensory receptors don't care about their input. The brain has to deal with the problem further downstream
+
+```{r generate-data-1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+dt <- 0.1
+motiondata <-  generate_data(T = 2, frequency = 0.5,
+                             amplitude = 20,
+                             sensor_sd = 2.0)
+# plot_trajectories(motiondata = motiondata)
+
+motiondata_hidden <- motiondata %>%
+    gather(type, value = value, -time) %>%
+    filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration")))
+
+p1 <- motiondata_hidden %>%
+        ggplot(aes(x = time, y = value)) +
+            geom_line(aes(linetype = type), size = 1.4) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "white",
+                       fill = "white", size = 6, stroke = 2) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "black",
+                       fill = "white", size = 4, stroke = 2) +
+            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+            xlab("Time (sec)") +
+            ylab("Angular position (deg)") +
+            theme_apa() +
+            theme(legend.title=element_blank())
+
+motiondata_2 <-  generate_data(T = 2, frequency = 0.5,
+                             amplitude = -20,
+                             sensor_sd = 1.8)
+# plot_trajectories(motiondata = motiondata)
+
+motiondata_hidden_2 <- motiondata_2 %>%
+    gather(type, value = value, -time) %>%
+    filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration")))
+
+p2 <- motiondata_hidden_2 %>%
+        ggplot(aes(x = time, y = value)) +
+            geom_line(aes(linetype = type), size = 1.4) +
+            geom_point(data = motiondata_2,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "white",
+                       fill = "white", size = 6, stroke = 2) +
+            geom_point(data = motiondata_2,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "black",
+                       fill = "white", size = 4, stroke = 2) +
+            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+            xlab("Time (sec)") +
+            ylab("Angular position (deg)") +
+            theme_apa() +
+            theme(legend.title=element_blank())
+
+cowplot::plot_grid(p1, p2, nrow = 2, labels = c("a", "b"))
+```

### 2017-01-25 manuscript.Rmd
fixed word conversion problem

@@
 class             : "man"
-# output            : papaja::apa6_word
 output:
-     # papaja::apa6_word
+     # papaja::apa6_word:
      papaja::apa6_pdf:
@@
                       fig.width = 12, fig.height = 8,
-                      dpi = 600, fig.path = "../generated-figures/",
-                      dev=c('pdf', 'png', 'cairo_ps'))
+                      dpi = 600, fig.path = "generated-figures/",
+                      dev=c('pdf', 'png', 'tiff', 'cairo_ps'))
 ```
@@
 
-![Perspective taking and imagined self-motion involve spatial transformations of a representation of the self.
+<!-- ![Perspective taking and imagined self-motion involve spatial transformations of a representation of the self.
 (a) If a speaker's goal is to take the perspective of an addressee, he must first solve the inverse problem of determining the appropriate action, and then run a forwards simulation in order to perform a translation and rotation of his own egocentric reference frame.
-These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.pdf){#fig:spt-schematic width=75%}
+These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.tiff){#fig:spt-schematic width=75%} -->
+
+```{r spt-schematic, fig.cap="Perspective taking and imagined self-motion involve spatial transformations of a representation of the self.
+(a) If a speaker's goal is to take the perspective of an addressee, he must first solve the inverse problem of determining the appropriate action, and then run a forwards simulation in order to perform a translation and rotation of his own egocentric reference frame.
+These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.", echo=FALSE}
+knitr::include_graphics('diagrams/fig-1-common-process.tiff')
+```
 
@@
                              sensor_sd = 2.0)
-# plot_trajectories(motiondata = motiondata)
+plot_trajectories(motiondata = motiondata)
 
-motiondata_hidden <- motiondata %>%
-    gather(type, value = value, -time) %>%
-    filter(type != "observations") %>%
-    mutate(type = factor(type,
-                         levels = c("position",
-                                    "velocity",
-                                    "acceleration")))
-
-p1 <- motiondata_hidden %>%
-        ggplot(aes(x = time, y = value)) +
-            geom_line(aes(linetype = type), size = 1.4) +
-            geom_point(data = motiondata,
-                       aes(x = time, y = observations),
-                       shape = 21, colour = "white",
-                       fill = "white", size = 6, stroke = 2) +
-            geom_point(data = motiondata,
-                       aes(x = time, y = observations),
-                       shape = 21, colour = "black",
-                       fill = "white", size = 4, stroke = 2) +
-            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
-            xlab("Time (sec)") +
-            ylab("Angular position (deg)") +
-            theme_apa() +
-            theme(legend.title=element_blank())
+# motiondata_hidden <- motiondata %>%
+#     gather(type, value = value, -time) %>%
+#     filter(type != "observations") %>%
+#     mutate(type = factor(type,
+#                          levels = c("position",
+#                                     "velocity",
+#                                     "acceleration")))
+# 
+# p1 <- motiondata_hidden %>%
+#         ggplot(aes(x = time, y = value)) +
+#             geom_line(aes(linetype = type), size = 1.4) +
+#             geom_point(data = motiondata,
+#                        aes(x = time, y = observations),
+#                        shape = 21, colour = "white",
+#                        fill = "white", size = 6, stroke = 2) +
+#             geom_point(data = motiondata,
+#                        aes(x = time, y = observations),
+#                        shape = 21, colour = "black",
+#                        fill = "white", size = 4, stroke = 2) +
+#             guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+#             xlab("Time (sec)") +
+#             ylab("Angular position (deg)") +
+#             theme_apa() +
+#             theme(legend.title=element_blank())
 # p1
-plot_trajectories(motiondata = motiondata)
 ```
@@
 ```{r graphical-model-inference, fig.cap="Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.", echo=FALSE}
-knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.pdf')
+knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.tiff')
 ```
@@
 This posterior subsequently forms the prior distribution for the next time step.", echo=FALSE}
-knitr::include_graphics('diagrams/figure-particle-filter-explained.pdf')
+knitr::include_graphics('diagrams/figure-particle-filter-explained.tiff')
 ```
@@
 ```{r eval=FALSE, include=FALSE}
-plot_pf_estimates(object = passive, predict = FALSE)
+# plot_pf_estimates(object = passive, predict = FALSE)
 ```
@@
 
-$$\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2}\Delta t^2$$
+$$ \theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2}\Delta t^2 $$
 
-$$\omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t$$
+$$ \omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t $$
+
 The effect that this has in terms of the inference algorithm is that the proposal distribution $P(x_t | y_{1:t})$, from which samples are drawn from during the 'predict' step, is improved.
@@
 
-```{r shaking-head, eval=FALSE, fig.cap="Particle filter applied to...", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
+```{r shaking-head, eval=FALSE, fig.cap="Particle filter applied to ", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
 dt <- 0.1
@@
 
-```{r two-external-forces, eval=FALSE, fig.cap="Particle filter applied to...", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
+```{r two-external-forces, eval=FALSE, fig.cap="Particle filter applied to ", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
 dt <- 0.1
@@
 ```{r graphical-model-control, fig.cap="Graphical model extended with hierarchical model for self-initiated movements.", echo=FALSE}
-knitr::include_graphics('diagrams/dynamic-belief-network-inference-2.pdf')
+knitr::include_graphics('diagrams/dynamic-belief-network-inference-2.tiff')
 ```
@@
 
-THe brain implements multiple priors:
+The brain implements multiple priors:
 
@@
 
-control input -> infer the consequences of one's own actions.
+control input allows inferring the consequences of one's own actions.
 
@@
 
-- place the emerging field of vestibular cognition with a framewokk for dynamic probabilistic inference and draw links to established models of sensory processing
+- place the emerging field of vestibular cognition with a framework for dynamic probabilistic inference and draw links to established models of sensory processing
 

### 2017-01-25 manuscript.Rmd
fixed ref commands and added png versions of diagrams

@@
 This effectively amounts to a running a simulation of dynamical system, consisting of simulated limb movements and the resulting sensory consequences.
-\autoref{fig:spt-schematic} (a) shows a simple scenario in which a speaker wants to give an addressee, standing opposite, verbal instructions about an action to perform on an external object.
+Figure \@ref(fig:spt-schematic) (a) shows a simple scenario in which a speaker wants to give an addressee, standing opposite, verbal instructions about an action to perform on an external object.
 The speaker must determine the most suitable action, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred reference frame to quickly simulate the best action.
@@
 These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.", echo=FALSE}
-knitr::include_graphics('diagrams/fig-1-common-process.tiff')
+knitr::include_graphics('diagrams/fig-1-common-process.png')
 ```
@@
 In order to achieve the desired position, the motor commands can then be simulated, and the expected sensory consequences can be estimated; this is known as the forward problem.
-\autoref{fig:spt-schematic} (b) shows that if the goal is to simply simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action.
+Figure \@ref(fig:spt-schematic) (b) shows that if the goal is to simply simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action.
 As demonstrated by @Penny:2013iv, these computations can be performed in a common probabilistic model.
@@
 The measurements are therefore observations of the true, unobservable angular velocity; however, they are contaminted by Gaussian noise, which is determined by characteristics of the sensory organs.
-\autoref{fig:head-turn-data} shows a sequence of noisy measurements during a two-second turn of the head.
+Figure \@ref(fig:head-turn-data) shows a sequence of noisy measurements during a two-second turn of the head.
 The acceleration generating the movement is sinusoidal, given for each time step $t$ by $Asin(2\pi t/T)$, where $A$ is the amplitude of the movement, $T = 1/f$ is the duration of the movement, and $f$ is the frequency.
@@
 <!-- \autoref{fig:graphical-model-inference} -->
-\autoref{fig:graphical-model-inference} shows a representation of a state-space model as a graphical model, unrolled over multiple time steps.
+\@ref(fig:graphical-model-inference) shows a representation of a state-space model as a graphical model, unrolled over multiple time steps.
 Starting at time 0, the state evolves according to the process model, with the measurements at each time step being generated according to the measurement model.
@@
 <!-- ![Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.
-](diagrams/dynamic-belief-network-inference-1.pdf){#fig:graphical-model-inference width=100%} -->
+](diagrams/dynamic-belief-network-inference-1.png){#fig:graphical-model-inference width=100%} -->
 
 ```{r graphical-model-inference, fig.cap="Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.", echo=FALSE}
-knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.tiff')
+knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.png')
 ```
@@
 
-\autoref{fig:particle-filter-explained} shows a visualization of the computations performed by a particle filter at each time step, in order to track the angular velocity.
+Figure \@ref(fig:particle-filter-explained) shows a visualization of the computations performed by a particle filter at each time step, in order to track the angular velocity.
 Initially, the particles are propagated forwards in time according to the proposal distribution.
@@
 This posterior subsequently forms the prior distribution for the next time step.", echo=FALSE}
-knitr::include_graphics('diagrams/figure-particle-filter-explained.tiff')
+knitr::include_graphics('diagrams/figure-particle-filter-explained.png')
 ```
 
-\autoref{fig:passive} shows the result of applying the particle filtering algorithm to the graphical model depicted in \autoref{fig:graphical-model-inference}, in order to infer the velocity and position during a two-second leftward head turn.
+Figure \@ref(fig:passive) shows the result of applying the particle filtering algorithm to the graphical model depicted in Figure \@ref(fig:graphical-model-inference), in order to infer the velocity and position during a two-second leftward head turn.
 
@@
 The estimation can be improved by increasing the process noise.
-\autoref{fig:passive-2} shows the resulting estimates.
+Figure \@ref(fig:passive-2) shows the resulting estimates.
 Whilst the lag is somewhat reduced, a consequence of increasing the process noise is that the measurement are given more weight, resulting in a tendency to overfit the noise in the measurements.
@@
 
-In terms of state-space models, this is implemented as a control input $u$, which extends the graphical model shown in \autoref{fig:graphical-model-inference}.
+In terms of state-space models, this is implemented as a control input $u$, which extends the graphical model shown in Figure \@ref(fig:graphical-model-inference).
 In this model, the estimated acceleration is replaced by an known acceleration signal.
@@
 
-\autoref{fig:graphical-model-control} shows the resulting graphical model.
+Figure \@ref(fig:graphical-model-control) shows the resulting graphical model.
 
 ```{r graphical-model-control, fig.cap="Graphical model extended with hierarchical model for self-initiated movements.", echo=FALSE}
-knitr::include_graphics('diagrams/dynamic-belief-network-inference-2.tiff')
+knitr::include_graphics('diagrams/dynamic-belief-network-inference-2.png')
 ```

### 2017-01-27 draft-manuscript.Rmd
added Notes, model-fits and new content

@@
 
-Mental imagery is oftern portrayed as being the result of top-down reconstruction of activity in perceptual circuits in the absence of physical stimulation of the sensors, and is opposed to bottom-up perception.
+> Mental imagery is oftern portrayed as being the result of top-down reconstruction of activity in perceptual circuits in the absence of physical stimulation of the sensors, and is opposed to bottom-up perception.
 However, in the case of the vestibular system, it becomes clear that this view is inadequate (the same is true for other modalities): the sensory that the vestibular sensors provide is noisy and ambiguous (examples), and in order to deal with these facts, the brain must use several strategies (which are well-known from control theory).
@@
 
-Head direction cells are thought to integrate signals of vestibular origin to maintain a signal of cumulative rotation (reviewed in @Taube:2007fk and are found in many brain areas, including the postsubiculum, retrosplenial cortex, thalamus, lateral mammillary nucleus, dorsal tegmental nucleus, striatum and entorhinal cortex [@Cullen:2014gx].
+~Head direction cells are thought to integrate signals of vestibular origin to maintain a signal of cumulative rotation (reviewed in @Taube:2007fk and are found in many brain areas, including the postsubiculum, retrosplenial cortex, thalamus, lateral mammillary nucleus, dorsal tegmental nucleus, striatum and entorhinal cortex [@Cullen:2014gx].
 This may be necessary in order to mentally simulate being in another position, i.e.
@@
 Why is this?
-Because vestibular connections make a major contribution to spatial orientation; information about the position in space obtained through the vestibular sensory system is computed by integrating velocity over time.
+Because vestibular connections make a major contribution to spatial orientation; information about the position in space obtained through the vestibular sensory system is computed by integrating velocity over time.~
 

### 2017-01-27 manuscript.Rmd
added Notes, model-fits and new content

@@
 ---
-# title             : "Modeling Cognitive Effects Vestibular Sensory Inference: From Sensory Inference To Imagined Self-Motion"
+# title: "Modeling Cognitive Effects Vestibular Sensory Inference: From Sensory Inference To Imagined Self-Motion"
 
-title             : "Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion"
+title: "Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion"
 
-# title             : "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
-# title             : "Probabilistic Computations Underlying Simulated Self-Rotations"
+# title: "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
+# title: "Probabilistic Computations Underlying Simulated Self-Rotations"
 # title: Probabilistic Computations Underlying Vestibular Cognition
@@
 
-keywords: probabilistic graphical models, particle filter, embodied cognition, vestibular cognition, spatial perspective taking, mental imagery
+keywords: probabilistic graphical models, particle filter, embodied cognition, vestibular cognition, spatial perspective taking, mental imagery, mental simulation
+
 wordcount: "X"
@@
 linkcolor         : "blue"
+fig_caption       : true
+
 # citeproc          : yes
@@
      # papaja::apa6_word:
-     papaja::apa6_pdf:
-        pandoc_args:
-        fig_caption: true
+     papaja::apa6_pdf
+        # pandoc_args:
+        # - --filter
+        #   pandoc-fignos
+
 ---
@@
 library(papaja)
-library(pander)
+# library(pander)
 library(ggplot2)
@@
 
-Adopting the spatial perspective of another person is generally considered an essential cognitive ability, and plays a vital role in our ability to determine what another person can see and in order to coordinate joint actions [@CreemRegehr:2013kx, @Pezzulo:2013iy].
-Many studies have also pointed out that perspective taking is of fundamental importance for language comprehension and communication [@Beveridge:2013gx], and social cognition in general [@Deroualle:2014ho].
+Adopting the spatial perspective of another person is an essential cognitive ability, and plays a vital role in our ability to determine possible actions of another person, and in order to coordinate joint actions [@CreemRegehr:2013kx, @Pezzulo:2013iy].
+Perspective taking is also considered to be of fundamental importance for language comprehension and communication [@Beveridge:2013gx], and social cognition in general [@Deroualle:2014ho].
 
@@
 
-Performing a simulation of body movements and estimating the sensory consquences amounts to simulating the behaviour of a dynamical system; this requires a generative model of the system.
-The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations.
+Performing a simulation of body movements and estimating the sensory consquences amounts to simulating the behaviour of a dynamical system using a forward model; this requires a generative model of the system.The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations.
 The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Baysian inference.
@@
 In this paper, we pursue this claim, and we attempt to the grounding of cognitive abilities in the computations performed by low-level sensory circuits in the service of dynamic real-time interactions with the world.
+A problem that needs to addressed is whether models that are applicable to lower-level sensory inference are scalable, in the sense that they are able to represent hierarchically structured knowledge, and can be applied to tasks requiring compositionality (?) [@Kwisthout:2016fx, @Griffiths:2012ii].
+In the field of cognitive science, sophisticated model of structured representations using Bayesian networks [@Pearl:1988wz] have been developed [@Tenenbaum:2011hd, @Lucas:2010tu], with applications to concept learning [CITATION], inductive biases in reasoning, in cognitive models of development [@Perfors:2011ex], language [@Goodman:2016go] and social cognition [@Baker:2009kx].
+@Kwisthout:2016fx make the point that predictive processing accounts are situated at low levels in the processing hierarchy, usually at the level of sensory inference or at the level describing kinematics.
+In this paper, we address this issue by starting from a model developed for low-level sensory processing, incorporating basic abilities to predict sensory consequences of movement based on knowledge about the kinematics, and extending this to account for interactions between thought processes, operationalized as mental simulations of sensorimotor experience, and sensory processing.
+
+> TODO: CITE [@Battaglia:2013fm]
+
+To our knowledge, few studies have addressed the idea that higher level cognitive processes are grounded in the circuits performing sensory inference.
+We claim that if cognition is grounded in sensory inference, and that this is to be experimentally studied by analyzing interactions between thought and perception, then it is vital that this be carried within the constraints of a theoretical (computational) framework.
+A suitable formalism for this is the framework of probabilistic graphical models, and within this framework, interactions between thought and perception can be modelled as effects of higher level priors and hyperpriors on inference at a computational level.
+Investigating inference at the level of algorithmic implementations might provide further insight.
 
@@
 
+We investigate one specific type of head movement, a rotation of the head about the yaw axis, and discuss this in the context of the probabilistic computations that are performed when the goal is to:
+
+1) infer motion caused by external events
+2) infer the consequences of one's own actions.
+
 # Dynamic Bayesian models for sensory inference
@@
 @Selva:2012gsa
-provide a recent review of the relationship between optimal observer models and models which employ Kalman filtering, and show that these are largely equivalent.
-Both @Karmali:2012cv and @Laurens:2006be described vestibular sensory inference in terms of particle filtering, which is a particular inference algorithm that can be used to perform inference in state-space models.
+provide a recent review of the relationship between optimal observer models and models that employ Kalman filtering, and show that these are largely equivalent.
+Both @Karmali:2012cv and @Laurens:2006be described vestibular sensory inference in terms of particle filtering, which is a particular algorithm that can be used to perform inference in state-space models.
 [@Doucet:2000bh, @Doucet:2009us].
@@
 We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can be used to perform inference in state-space models.
-While Kalman filters provide the optimal solution when the restrictions of linear Gaussian State-space models are met, they are not easily extensible to non-linear models, which is relevant when considering multi-dimensional motion.
+While Kalman filters provide the optimal solution when the restrictions of linear Gaussian state-space models are met, they are not easily extensible to non-linear models, which is relevant when considering multi-dimensional motion.
 Particle filtering can be applied to arbitrary nonlinear models.
 State-space models can be seen as special cases of more general Bayesian hierarchical models [@Bishop:2006ui; @Murphy:2012ua].
-Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model
+Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model.
  <!-- from a statistical perspective, rather an engineering perspective -->
-A state-space model is special case of a dynamic latent variable model which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations.
-As an example, consider trying to estimate the velocity and position during a leftward head turn.
+A state-space model is special case of a dynamic latent variable model, which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations.
+As such, a state-space model can be viewed as a statistical model of the problem at hand.
+
+For example, consider trying to estimate the velocity and position during a leftward head turn.
 The afferent SCC signals are considered to be proportional to the angular velocity of the head with a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv].
@@
 <!-- \autoref{fig:graphical-model-inference} -->
-\@ref(fig:graphical-model-inference) shows a representation of a state-space model as a graphical model, unrolled over multiple time steps.
+Figure \@ref(fig:graphical-model-inference) shows a representation of a state-space model as a graphical model, unrolled over multiple time steps.
 Starting at time 0, the state evolves according to the process model, with the measurements at each time step being generated according to the measurement model.
@@
 $$ \omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t $$
+$$ \alpha_t = \alpha_{t-1} $$
 
@@
 > QUESTION: shall we show a model that includes acceleration in its state?
+
+# TODO
+How do we explain higher-level anticipation/expectations in such a model?
+This form of anticipation is not the same as the 1-step ahead predictions made during inference, but rather implement higher level prior beliefs about the direction of movement, for example.
+
+An obvious way of including this is the mechanism that allows one to predict the sensory consequences of self-initiated movement via an efference copy.
 
@@
 
+__A further issue is that the sensory inference model cannot incorporate any top-down expectations or biases.
+For example, imagine seeing that someone is going to push you.
+You can use this information, derived from the visual modality, in order prepare yourself in order to compensate for this.
+However, you do not want to act too early, but rather wait until you are actually moving.__
+
 In terms of state-space models, this is implemented as a control input $u$, which extends the graphical model shown in Figure \@ref(fig:graphical-model-inference).
@@
 
-#TODO: model implementation
+We implement the control in the following manner: during an active head movement, the model generates a sequence of control signals $u_{1:T}$.
+For example, during a leftward head turn, using the sinusoidal acceleration profile shown in Figure \@ref(fig:head-turn-data), the acceleration is given by $Asin(2\pi t/T)$.
+Assuming the model has full knowledge about the onset of movement, and the type of movement (sinusoidal head turn), the model requires information about the frequency, $f = 1/T$, amplitude, $A$, and direction, which is given by the sign of the amplitude.
+These are
+
+<!-- Equation XXX [refer to acceleration] -->
 
@@
 
+Show a table demonstrating the difference between:
+
+- prediction: 1-step ahead during inference
+- expectation: higher level priors
+
 The brain implements multiple priors:
@@
 - ability to select type of motion
-- parameters of tht motion
+- parameters of the motion
 
@@
 # Various sources of interference
-Here, we identify various sources of intereference (2-way) between thought (higher-level cognition) and lower-level sensory processing.
+
+Interference is rather surprising; if the goal is to perform simulation without incorporating the data, then we should run a Monte Carlo simulation.
+Why should this interfere with sensory inference perform by a particle filter?
+We propose that the answer is in the shared generative model.
+
+Here, we identify various sources of interference (2-way) between thought (higher-level cognition) and lower-level sensory processing.
 This interference can be either due to low-level influences in sensory inference, due to priors on higher-level parameters, choice of process model (false expectations of certain type of motion).
@@
 ## Summary
-Interference may the result of either inappropriate priors, leading to a mis-specified filtering algorithm, or biased decision-making.
+Interference may the result of either:
+
+1) mis-specified proposal distribution: inappropriate priors, or wrong type of motion or
+2) biased decision-making
+
 We should be able to investigate which by applying cognitive process models that explain response times and choices.
@@
 # Simulations
+
+## Model specifications
+
+## 3-d state space model
+
+Figure \@ref(fig:3d-statespace) show a 3-d state space model
+
+```{r 3d-statespace, echo=FALSE, fig.cap="3-d state space model.", message=FALSE, warning=FALSE}
+
+load("simulations/model-fits/m1-3d-statespace.Rda")
+
+df_w <- with(m1_3d_statespace$summ_smc, {
+    data_frame(t = 1:length(w$f$mean),
+               mean = w$f$mean,
+               lower = w$f$quant$`0.025`,
+               upper = w$f$quant$`0.975`)
+})
+df_w$param <- "Velocity"
+df_w$true <- m1_3d_statespace$model$data()$w_true
+df_w$observations <- m1_3d_statespace$model$data()$y
+
+df_p <- with(m1_3d_statespace$summ_smc, {
+    data_frame(t = 1:length(p$f$mean),
+               mean = p$f$mean,
+               lower = p$f$quant$`0.025`,
+               upper = p$f$quant$`0.975`)
+})
+df_p$param <- "Position"
+df_p$true <- m1_3d_statespace$model$data()$p_true
+df_p$observations <- m1_3d_statespace$model$data()$y
+
+df <- bind_rows(df_p, df_w)
+
+p1 <- df %>%
+    filter(param == "Position") %>%
+    ggplot(data = ., aes(x = t)) +
+    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
+
+    geom_line(aes(y = true), alpha = 0.9,
+              linetype = 4, size = 1.2) +
+
+    geom_line(aes(y = mean), size = 1.6) +
+    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
+    scale_x_continuous(limits = c(1, 20),
+                       breaks = c(0, 5, 10, 15, 20),
+                       labels = c("0", "0.5", "1", "1.5", "2")) +
+    # xlab("Time (sec)") + 
+    ylab("Angular position (deg)") +
+    papaja::theme_apa(base_size = 18) +
+    theme(axis.title.x=element_blank()) +
+    theme(panel.border = element_blank(), 
+          axis.line = element_line(colour="black", size= 0.75, lineend="square"))
+
+p2 <- df %>%
+    filter(param == "Velocity") %>%
+    ggplot(data = ., aes(x = t)) +
+    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
+
+    geom_line(aes(y = true), alpha = 0.9,
+              linetype = 4, size = 1.2) +
+
+    geom_line(aes(y = mean), size = 1.6) +
+    geom_point(aes(y = observations),
+               alpha = 1, fill = "white",
+               colour = "white",
+               shape = 21, size = 6) +
+    geom_point(aes(y = observations), alpha = 1,
+               fill = "black",
+               colour = "grey40", shape = 21, size = 4) +
+    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
+    scale_x_continuous(limits = c(1, 20),
+                       breaks = c(0, 5, 10, 15, 20),
+                       labels = c("0", "0.5", "1", "1.5", "2")) +
+    xlab("Time (sec)") + 
+    ylab("Angular Velocity (deg/s)") +
+    papaja::theme_apa(base_size = 18) +
+    theme(panel.border = element_blank(), 
+          axis.line = element_line(colour="black", size = 0.75, lineend="square"))
+
+cowplot::plot_grid(p1, p2, ncol = 1) 
+                   # labels = c("Position", "Velocity"),
+                   # label_size = 12)
+
+```
+
+## Monte Carlo simulation
+Show dynamics when simulating using various proposal distributions implementing various natural head motions:
+
+- turing left
+- shaking head quickly
+- shaking head slowly (communicative gesture)
+
+## Expectations based on higher level priors
+In our model, there is no difference between modelling active self-motion, and an anticpation of passive self-motion due to predictable external forces. This is a testable hypothesis.
+
+## Model mis-specification
+### Inappropriate priors
+
+### Wrong type of motion model
+For example, expect shaking instead of head turn
+
+```{r active-1, cache=TRUE, fig.cap="Mis-specified proposal distribution.", echo=FALSE, message=FALSE, warning=FALSE}
+
+move <- function(A, t, Time) {
+    rnorm(1, A * sin(2 * pi * (t - 1)/Time), 0.1)
+}
+
+transfun <- function(x, t, Time, A, sigma, N, dt = 0.1, selfmotion) {
+    mu <- matrix(rep(0, 2*N), nrow = N)
+
+    if (selfmotion == "active") {
+        mu[, 1] = x[, t-1, 1] + dt * x[, t-1, 2] + dt^2/2 * move(A, t, Time)
+        mu[, 2] = x[, t-1, 2] + dt * move(A, t, Time)
+    } else if (selfmotion == "passive") {
+            mu[, 1] = x[, t-1, 1] +  dt * x[, t-1, 2]
+            mu[, 2] = x[, t-1, 2]
+        }
+        else {
+            mu[, 1] = 0
+            mu[, 2] = 0
+        }
+    # TODO: more elegant way of doing this?
+    t(apply(mu, 1, function(x) {rmvnorm(1, mean = x, sigma = sigma)}))
+}
+
+params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
+                x_init = c(0, 0),
+                sdx_init = 0.5*eye(2),
+                A = 10, N = 1000,
+                dt = 0.1, freq = 0.5,
+                transfun = transfun)
+
+m1 <- run_particle_filter(motiondata = motiondata, params = params,
+                            resample_particles = TRUE, rs_thresh = 0.5,
+                            selfmotion = "active")
+
+plot(m1, panel_label = "", print_loglik = FALSE, predict = FALSE, imagine = FALSE)
+```
 
@@
 - Wertheim
+[@Grabherr:2011ej] showed that patients with bilateral or unilateral vestibular loss show impaired ability to mentally transform images of bodies and body parts.
 
@@
 ### Head direction cells
-Why dynamic simulation? Maybe in order to provide vestibular in put to head direction cells [@Taube:2007fk]
+Head direction cells are thought to integrate signals of vestibular origin to maintain a signal of cumulative rotation (reviewed in @Taube:2007fk and are found in many brain areas, including the postsubiculum, retrosplenial cortex, thalamus, lateral mammillary nucleus, dorsal tegmental nucleus, striatum and entorhinal cortex [@Cullen:2014gx].
+This may be necessary in order to mentally simulate being in another position, i.e.
+the brain apparently needs to simulate the n-step dynamics of its motion through space.
+Why is this?
+Because vestibular connections make a major contribution to spatial orientation; information about the position in space obtained through the vestibular sensory system is computed by integrating velocity over time.
+
+- mental simulation cannot just re-use the state used in sensory inference, because the velocity is fed to the head direction cells. The brain would get confused
+- directional signal is not just a simple integration of velocity signals from the vestibular system
+- signals from motor efferecence copy or proprioceptive systems appear to be crucial for upating the preferred firing direction over time
+
+## Testable hypotheses
+Show a table of testable hypotheses
 

### 2017-01-27 notes.Rmd
added Notes, model-fits and new content

--- 2017-01-22 notes.Rmd
@@
-# head direction cells
+---
+output: 
+  html_document: 
+    fig_caption: yes
+    number_sections: yes
+    theme: spacelab
+---
 
-[@Taube:2007fk]
+# one
+that the brain is able to imagine, plan, perform counterfactual simulations because it coopts the mechanism evolved in order to predict the sensory consequences of its actions.
 
-- mental simulation cannot just er-use the state used in sensory inference, because the velocity is fed to the head directino cells. The brain would get confused
+interference arises when he brain, due to task demands, uses the control mechanism in a manner which is not entirely counterfactual
 
-- directinal signal is not just a simple integration of velocity signals from the vestibular system
-- signals from motor efferecence copy or proprioceptive systems appear to be crucial for upating the preferred firing direction over time
-- what about this? you don't know where you are, because you were passively moved. Can you update if I tell you (verbally instruct you)?
-- vestibular input is eesntial for establishing the head direction signal?
-How can we reconcile this with the assumption that re-afference is cancelled, i.e.
-input due to active motion is cancelled.
-Maybe it's the control model...
+# two
+interference is surprising, because a mental simulation should be counterfactual.
+if it isn't this might be a problem.
+why should thinking about rotating left have an effect on your ability to sense leftward motion?
 
-# Corollary discharge
-[@Crapse:2008jc]
+we show this type of interactions needs to be investigate pdf in a probabilistic framework
 
-# Predictive control 
-[@Stahl:2011dl]
+# three
+we identify 2 main problem that have not been addressed: in SPT how do people decide which way and for how long to rotate?
 
-- movemements can induce sensory input
-- sensory input due to own actions is indistinguishable from input due to external forces
-- strategy: efference copy -> corollary discharge
+and how does imagery interfere?
 
-it seems intuitively obvious that this mechanism has something to do with mental imagery.
-This has been pointed out many times; mental imagery and mental simulation more generelly is somehow related to the capacity to compute corollay discharge, based on the efference copy and a forward model.
+the first issue speaks to the fact that we have to solve the inverse problem before we can simulate in order to transform our FOR.
+the is equivalent to solving a planning problem, I.e.
+choosing the best action in order to reach a goal.
+this is actually counterfactual inference.
 
-> Necessary components:  efference copy and a forward model
+after introducing the graphical model, we show the types of inference, I.e. which nodes are clamped.
 
-Sensory receptors don't care about their input. The brain has to deal with the problem further downstream
+then we explain interference as biased perceptual decision making, concentrating on the Nigmatullina effect. how do biases arise? the basic model is not equipped to deal with bias.
+
+# four
+but how does interference through CVS on mental spatial transformation work?
+We hypothesize that CVS is a strange stimulus and needs interpretation by the brain.
+Passive actual movement puts the brain in an alert mode, because it might have to counteract the motion, I.e.
+the control circuits are engaged, leading to less capacity for planning.
+
+in any case, we need a closer look at the connection between menral simulation and planning and the ability to predict self initiated movements.
+
+# five
+finally, how can we explain the bias in the Nigmatullina experiment?
+we need to investigate whether it is due to a bias in the starting point, in terms of a traditional method of analyzing such data, i.e.
+DDM.
+this is implemented by a prior on the choice alternatives, leading to starting point or altered drift rate.
+
+or is it a model misspecification?
+
+The alternative model by Merfeld would say it's just the starting point.
+
+# six
+perform simulations to show a theoretical analysis of what may cause the Nigmatullina effect.
+
+model misspecification might be unlikely, but certainly worth a look
+
+# seven
+we tell the story in two parts.
+
+1) describe how we can simulate using a parsimonious model that refuses components of the sensory inference model and components involved in distinguishing active from passive movements.
+
+2) what happens when we simulate.
+switch off inputs, sensory gating corresponding to not resampling.
+show tables of correspondence mechanisms.
+show that interest,S of probabilistic computations the ability to model sensory consequences is a question of creating a proposal distribution for sampling from the prior.
+show table comparing Monterey carlo simulation with particle filtering.
+first show tables explaining components of filtering.
+
+# eight
+of course we first elucidate how the ability to simulate is related to sensory inference in real time.
+
+vestibular sensory inference is a very good example, because it se ms to be well described as sequential monte carlo
+
+# nine
+interference is either:
+
+1) mis-specified proposal distribution
+2) biased decision-making
+
+# 10
+we focus explicitly know:
+
+- forward simulation, explaining the requirements from a statistical / probabilistic point of view
+- interference from thought to inference, and explore how this might occur. we focus on this direction because it is easier to understand than the other direction
+
+show a sequence diagram showing a typical trial in an experiment
+
+# eleven
+we implement forward sampling by sampling from the proposal distribution without updating. this corresponds to monte carlo simulation
+
+In order to set up a particular simulation, the brain must create a proposal distribution such that it represents the kinematics of the system to be modelled.

### 2017-01-27 simulations/01-3d-statespace.R

@@ -1,0 +1,321 @@
+# Rbiips model: 3-d state  ----
+
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(Rbiips)
+
+# library(viridis)
+# color_palette <- viridis::plasma(n = 9)
+
+sd2precision <- function(sd) {
+    prec <- 1/(sd^2)
+    prec
+}
+
+neff <- function(weights) {
+    1/sum(weights^2)
+}
+
+Mode <- function(v) {
+    uniqv <- unique(v)
+    uniqv[which.max(tabulate(match(v, uniqv)))]
+}
+
+#
+# trdata <- vestcog::generate_data(T = 2, amplitude = 20, sensor_sd = 1.7,
+#                                  as_df = TRUE, seed = TRUE)
+
+dt <- 0.1
+t_max <- 2 # how long is the movement (in seconds)? #length(trdata$observations)
+freq <- .5 # what is the frequency (in Hz)?
+
+mean_a_init <- 0
+prec_a_init <- sd2precision(1)
+mean_w_init <- 0
+prec_w_init <- sd2precision(1)
+mean_p_init <- 0
+prec_p_init <- sd2precision(1)
+
+prec_a <- sd2precision(0.5)
+prec_w <- sd2precision(0.5)
+prec_p <- sd2precision(1.0)
+prec_y <- sd2precision(1.7)
+
+# f_eval <- function(x, k) { .5 * x + 25*x/(1+x^2) + 8*cos(1.2*k) }
+# f_dim <- function(x_dim, k_dim) { 1 }
+#
+# f_eval_2 <- function(A, freq, t) {
+#     A/2 * sin(two_pi*(t-1) * freq) + A * sin(two_pi*(t-1) * 4 * freq)
+# }
+
+data = list(t_max = t_max/dt,
+            dt = dt,
+            freq = dt * freq,
+            A = 20,
+            # y = trdata$observations,
+            #a = trajectory_data$acceleration/20,
+            mean_a_init = mean_a_init,
+            prec_a_init = prec_a_init,
+            mean_w_init = mean_w_init,
+            prec_w_init = prec_w_init,
+            mean_p_init = mean_p_init,
+            prec_p_init = prec_p_init,
+            prec_a = prec_a,
+            prec_w = prec_w,
+            prec_p = prec_p,
+            prec_y = prec_y,
+            two_pi = 2*pi)
+
+model_string <- "
[…]

### 2017-01-27 manuscript.Rmd
draft introduction

@@
 abstract: >
-        Recent research has indicated that the vestibular network is involved in various cognitive tasks involving mental spatial transformations.
+        Mental spatial transformations are considered to be an essential cognitive ability.
+Recent research has indicated that the vestibular network is involved in such mental spatial transformations.
 However, the nature of this involvement remains unclear.
-Here, we provide a computational framework for vestibular cognition and discuss this in the context of perspective transformations, which require a simulated rotation of the self.
-We explore how the ability to model the effects of body rotation using vestibular input might enable the ability to imagine body rotations.
-We approach this problem by extending previous models of vestibular sensory processing to include the ability to model the sensory consequences of self-initiated movements, and discuss how this can be further extended to allow mentally simulated rotations of the self.
-We ...
-Predicting the sensory input is key to understanding imagined rotations.
-Mental imagery goes beyond prediction, rather, it involves the use of counterfactual queries in a probabilistic model.
+We propose a computational framework for the involvement of the vestibular system in cognition and discuss this in the context of perspective transformations, which require a simulated rotation of the self.
+We explore how the ability to anticipate the sensory consequences of self-initiated actions might underlie the ability to imagine body rotations.
+We approach this problem by extending previous models of vestibular sensory inference to include the ability to model the sensory consequences of self-initiated movements, and discuss how this can be further extended to allow mentally simulated rotations of the self.
+Our model is able to explain various source of interference between simulations of self-rotation and sensory inference.
 
@@
 Perspective taking is also considered to be of fundamental importance for language comprehension and communication [@Beveridge:2013gx], and social cognition in general [@Deroualle:2014ho].
-
-A number of recent studies have demonstrated that the vestibular system, which deals with sensory signal relating to movements of the head, is involved in such spatial transformations [@Lenggenhager:2008et, @Deroualle:2015gk, @Falconer:2012cl, @Gardner:2016kd].
+<!-- [ok] -->
+A number of recent studies have demonstrated that the vestibular system, which processes sensory signal relating to angular and linear movements of the head, is involved in such spatial transformations [@Lenggenhager:2008et, @Deroualle:2015gk, @Falconer:2012cl, @Gardner:2016kd].
 This is demonstrated in the form of interference effects between mental transformations and concurrent sensory stimulation.
-However, to date, there has little attempt to understand the involvement of the vestibular system in higher-level cognitive abilities from a computational perspective.
-In this paper, we focus specifically on the role of the vestibular system for spatial perspective transformations, and we draw the link between probabilistic computations performed by the vestibular system in the service of real-time interaction with the world, and the brain's ability to run mental simulations for the purpose of adopting the perspective of another person.
-
+To date, however, there has little attempt to understand the involvement of the vestibular system in higher-level cognitive abilities in terms of the computational mechanisms.
+<!-- [ok] -->
 Consider the use of deictic pronouns, such as "this" or "that" and "left" or "right".
-These terms only assume a their meaning when placed within a certain frame of reference (FOR).
-For example, in a dialogue between two speakers, the statement "I would like that one", or "I would like the one on the right" can only be understood under the assumption that the deictic terms "that" and "right" are used in an egocentric FOR.
-In addition, both interlocutors must be able to access the other's FOR with respect to an external FOR.
-In other words: In order to give instructions, the speaker must be able to mentally assume the addressee's position in space.
+These terms only assume a meaning relative to a certain frame of reference (FOR).
+For example, in a dialogue between two speakers, the statement "I would like that one", or "I would like the one on the right" can only be understood if both interlocutors are able to access the other's FOR with respect to an external FOR.
 A further example that requires a similar spatial transformation is when two people want to perform a coordinated action, such as lifting a heavy object.
-This requires having knowledge of the other's potential actions in order to adjust one's own actions, and this again requires knowledge of the other's position with an external FOR.
-A judgement about what kind of actions the other person can perform requires an embodied simulation using a representation of one's own body schema [@Pezzulo:2013iy].
-Giving an instruction to another person thus requires running an embodied simulation in order to infer what kind of motor commands the other person should perform __[Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod, 2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo, 2011a,b]__.
-This embodied simulation is performed using one's own body schema.
-Communicating this information to the other actor requires knowing the position of the addressee's FOR relative to the external reference frame, and the difference between the speaker and the addressee's egocentric reference frames.
-Perspective taking thus can be conceived of as a aligning a representation of one's body with that of another person in order to access spatial information from a viewpoint other than one's own egocentric viewpoint.
+This requires having knowledge of the other's potential actions in order to adjust one's own actions, and this again requires knowledge of the other's position within an external FOR.
+A judgement about what kind of actions the other person can perform requires an embodied simulation using a representation of one's own body schema [@Pezzulo:2013iy] and giving an instruction to another person thus requires running an embodied simulation in order to infer what kind of motor commands the other person should perform __[Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod, 2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo, 2011a,b]__.
+But this again requires information about the addressee's FOR relative to the external reference frame, and the difference between the speaker and the addressee's egocentric reference frames.
+Perspective taking can be conceived of as a aligning a representation of one's body with that of another person in order to access spatial information from a viewpoint other than one's own egocentric viewpoint.
 In order to achieve this alignment, one must perform translations and rotations of one's own egocentric reference frame.
 Processing translations and rotations of the head and body is the core domain of the vestibular system, and thus is of particular interest for research into mental spatial transformations.
-Spatial transformations in 6 dimensions (translation along 3 axes, and rotations around 3 axes) are highly complex, and non-linear, and we propose that the brain adopts the strategy of simulating motion of the body through space in order achieve a desired target position.
+<!-- [ok] -->
+Spatial transformations in 6 dimensions (translation along 3 axes, and rotations around 3 axes) are highly complex, and non-linear, and it has been proposed that the brain adopts the strategy of simulating motion of the body through space in order achieve a desired target position [CITATION].
 This effectively amounts to a running a simulation of dynamical system, consisting of simulated limb movements and the resulting sensory consequences.
-Figure \@ref(fig:spt-schematic) (a) shows a simple scenario in which a speaker wants to give an addressee, standing opposite, verbal instructions about an action to perform on an external object.
-The speaker must determine the most suitable action, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred reference frame to quickly simulate the best action.
+Figure \@ref(fig:spt-schematic) (a) shows a simple scenario in which a speaker wants to tell an addressee, standing opposite, to pick something up from the floor.
+The speaker must determine the most suitable action by embodied simulation, given the addressee's position relative to the external object, and in order to achieve this goal, the speaker must align himself with the addressee, so that he can use his own body-centred FOR.
 In this example, alignment with the addressee's reference frame consists of a forwards translation and a rotation around the vertical axis.
@@
 
-In order to do this, the speaker must first determine what type of movement to perform, and the direction, velocity and duration that lead to the desired outcome.
-This means that in order to know what to simulate, the speaker must first work backwards from the desired outcome, and infer the actions that are most likely to lead to that outcome.
-This is known in the field of motor control as the inverse problem [@Wolpert:1998td].
+In order to do this, the speaker must first determine what type of movement to perform.
+This means that in order to know what to simulate, the speaker must first work backwards from the desired outcome, and infer the actions that are most likely to lead to that outcome [@Hamrick:2014wq].
+In the field of motor control, this is known as the inverse problem [@Wolpert:1998td].
 In order to achieve the desired position, the motor commands can then be simulated, and the expected sensory consequences can be estimated; this is known as the forward problem.
-Figure \@ref(fig:spt-schematic) (b) shows that if the goal is to simply simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action.
+Figure \@ref(fig:spt-schematic) (b) shows that if the goal is to deliberatively simulate the perceptual experience of rotation around the body's vertical (yaw) axis, rather than project a representation of one's body onto a target, the problem is simplified; the actor can perform forward inference without the need to determine the optimal motor action.
 As demonstrated by @Penny:2013iv, these computations can be performed in a common probabilistic model.
@@
 Further, [@Pickering:2014bo] discuss the relationship between control-theoretic notions of inverse/forward models, known collectively as internal models, and and more general concepts of a generative model in a hierarchical Bayesian framework.
+<!-- [ok] -->
+Simulating the behaviour of a dynamical system using a forward model requires a generative model of the system.
 
-Performing a simulation of body movements and estimating the sensory consquences amounts to simulating the behaviour of a dynamical system using a forward model; this requires a generative model of the system.The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations.
-The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Baysian inference.
+The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations.
+The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Baysian inference [Oman, CITATIONS, @PaulRMacNeilage:2008bv].
+
 The vestibular system faces a number of challenges in interpreting the sensory measurements provided by the semi-circular canals (SCC) and otolith organs.
-On the one hand, these signals  are very noisy, and ambiguous [@PaulRMacNeilage:2008bv]; the otoliths respond both to translations and tilt of the head relative to gravity, and in order to disambiguate between translation and tilt, the brain must perform sophisticated computations at the level of the brain stem.
-It has become clear that the nature of these computations is probabilistic, and that the brain makes use of strong prior beliefs in order to interpret these sensory signals.
+On the one hand, these signals  are very noisy, and ambiguous; the otoliths respond both to translations and tilt of the head relative to gravity, and in order to disambiguate between translation and tilt, the brain must perform sophisticated computations.
+The emerging picture is that the nature of these computations is probabilistic, and that the brain makes use of strong prior beliefs in order to interpret these sensory signals [Mittelstaedt, etc].
 A further challenge faced by the vestibular system is that it needs to distinguish between sensory signals that are the result of active motion, and those resulting from external forces [@Cullen:2009dc, @Cullen:2014gx].
@@
 This was recently discussed by [@Tian:2012ui] in the context of imagined speech.
-
+<!-- [ok] -->
 The putative link between predictive processing and higher cognitive processing has a firm basis in predictive processing accounts of sensory inference [Friston].
@@
 The brain seems to be actively engaged in formulating hypotheses, and testing these on sensory data; this has been extended to the concept of active inference [Friston], in which an agent attempts to either change it's internal model of world, or to interven on the world in order to fulfil sensory predictions.
+
 We claim that the vestibular modality is particularly well-suited to investigating predictive processing at multiple hierarchical scales, from the lowest levels of sensory inference to higher levels, where goals, intentions and beliefs of the organism become relevant.
 A case in point is the study by @Wertheim:2001cb, who provided evidence that participants' perceptions of tilt induced by vestibular stimulation are shaped to a large extent by the beliefs about possible motions they formed prior to taking part in the experiement.
+<!-- [ok] -->
 
 Interference effects between higher-level cognitive processes, such as SPT and mental imagery, and sensory inference, are thought to occur as a result of shared circuitry [@Deroualle:2014ho, @Nigmatullina:2015gh].
-@Ellis:2017fg recently claimed that these types of interference effects might arise as a result of cognitive processing exerting an influence on circuits performing sensory inference via shared higher-level priors.
-In this paper, we pursue this claim, and we attempt to the grounding of cognitive abilities in the computations performed by low-level sensory circuits in the service of dynamic real-time interactions with the world.
-A problem that needs to addressed is whether models that are applicable to lower-level sensory inference are scalable, in the sense that they are able to represent hierarchically structured knowledge, and can be applied to tasks requiring compositionality (?) [@Kwisthout:2016fx, @Griffiths:2012ii].
-In the field of cognitive science, sophisticated model of structured representations using Bayesian networks [@Pearl:1988wz] have been developed [@Tenenbaum:2011hd, @Lucas:2010tu], with applications to concept learning [CITATION], inductive biases in reasoning, in cognitive models of development [@Perfors:2011ex], language [@Goodman:2016go] and social cognition [@Baker:2009kx].
+However, the precise nature of such interactions, and what exaclty might be shared, is left unspecified.
+At first glance, it appear unlikely that such interference should be possible, as low levels of vestibular processing appear to be cognitively impenetrable [CITATION].
+In particular, @Nigmatullina:2015gh found that the onset of the vestibulo-ocular reflex (VOR) was earlier when the direction of an imagined self-rotation was congruent with the direction of a subsequently experienced rotation.
+The VOR, however, must be susceptible to the influence of intentions and goals; for example, if a person intends to fixate a target during a movement, the VOR should be cancelled.
+Thus, seemingly reflexive, and impenetrable, motor actions require a large amount of flexibility based on top-down information.
+Even accepting the fact that low level sensorimotor processing may be more cognitively penetrable than previously thought, the fact that imagining a rotation of the self should have an effect on a person's ability to detect rotations, or to discriminiate between directions of rotation, still requires an explanation.
+
+> NOTE: we need to include all studies demonstrating interference, especially in the direction cognition -> inference (Nigmatullina effect)
+
+@Ellis:2017fg recently claimed that the types of interference effects might arise as a result of cognitive processing exerting an influence on circuits performing sensory inference via shared higher-level priors.
+In particular, the interference described by @Nigmatullina:2015gh looks like an effect of a bias in the sensory discrimination task.
+Although vestibular perceptual thresholds have been extensively investigate for yaw rotations [@Merfeld:2011fs; @Grabherr:2008kk], the methods used for estimating thresholds require unbiased responding by the participants.
+There have no investigations of biasing effects on yaw rotations.
+
+In this paper, we pursue the claim that interference should be investigated as a bias, and we attempt to explain the grounding of higher level cognitive abilities in the computations performed by low-level sensory circuits in the service of dynamic real-time interactions with the world.
+<!-- [ok] -->
+
 @Kwisthout:2016fx make the point that predictive processing accounts are situated at low levels in the processing hierarchy, usually at the level of sensory inference or at the level describing kinematics.
-In this paper, we address this issue by starting from a model developed for low-level sensory processing, incorporating basic abilities to predict sensory consequences of movement based on knowledge about the kinematics, and extending this to account for interactions between thought processes, operationalized as mental simulations of sensorimotor experience, and sensory processing.
 
-> TODO: CITE [@Battaglia:2013fm]
+In this paper, we address this issue by starting from a model developed for low-level sensory inference, incorporating basic abilities to predict sensory consequences of movement based on knowledge about the kinematics, and extending this to account for interactions between thought processes, operationalized as mental simulations of sensorimotor experience, and sensory processing.
 
-To our knowledge, few studies have addressed the idea that higher level cognitive processes are grounded in the circuits performing sensory inference.
-We claim that if cognition is grounded in sensory inference, and that this is to be experimentally studied by analyzing interactions between thought and perception, then it is vital that this be carried within the constraints of a theoretical (computational) framework.
-A suitable formalism for this is the framework of probabilistic graphical models, and within this framework, interactions between thought and perception can be modelled as effects of higher level priors and hyperpriors on inference at a computational level.
+@Battaglia:2013fm addressed the idea that the higher level cognitive ability to perform mental rotation are grounded in the circuits performing sensory inference.
+We claim that if vestibular cognition is grounded in sensory inference, and that this is to be experimentally studied by analyzing interactions between thought and sensory inference, then it is vital that this be carried within the constraints of a theoretical (computational) framework.
+A suitable formalism for this is the framework of probabilistic graphical models [@Pearl:1988wz, @Koller:2009ty], and within this framework, interactions between thought and perception can be modelled as effects of higher level priors and hyperpriors on inference at a computational level.
 Investigating inference at the level of algorithmic implementations might provide further insight.
+
+Probabilistic graphical models are able to represent hierarchically structured knowledge [@Kwisthout:2016fx, @Griffiths:2012ii], which is required for Bayesian models of cognition [@Tenenbaum:2011hd, @Lucas:2010tu].
+This approach has been used to model concept learning [@Kemp:2007eo], inductive biases in reasoning, in cognitive models of development [@Perfors:2011ex], language [@Goodman:2016go] and social cognition [@Baker:2009kx].
+
+In this paper, we focus specifically on the role of the vestibular system for spatial perspective transformations, and we draw the link between probabilistic computations performed by the vestibular system, and the brain's ability to run mental simulations for the purpose of adopting the perspective of another person.
 
@@
 
-We investigate one specific type of head movement, a rotation of the head about the yaw axis, and discuss this in the context of the probabilistic computations that are performed when the goal is to:
+Following @Karmali:2012cv, we focus on one specific type of head movement, a rotation of the head about the yaw axis, and discuss this in the context of the probabilistic computations that are performed when the goal is to:
 
@@
 2) infer the consequences of one's own actions.
+
+<!-- [ok] -->
 
@@
 This is generally seen as evidence for the existence of an internal model in the processing of vestibular signals.
-@Karmali:2012cv focussed on one-dimensional angular velocity during yaw rotations; their model was able to explain the phenomenon of velocity storage as a consequence of particle filtering, with the time constant of velocity storage depending on the variability of the sensory afferent signals.
+@Karmali:2012cv focussed on one-dimensional angular velocity during yaw rotations; their model was able to explain the phenomenon of velocity storage as a consequence of particle filtering, with the time constant of velocity storage depending on the noise of the sensory afferent signals.
 
@@
 We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can be used to perform inference in state-space models.
-While Kalman filters provide the optimal solution when the restrictions of linear Gaussian state-space models are met, they are not easily extensible to non-linear models, which is relevant when considering multi-dimensional motion.
-Particle filtering can be applied to arbitrary nonlinear models.
 State-space models can be seen as special cases of more general Bayesian hierarchical models [@Bishop:2006ui; @Murphy:2012ua].
-Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model.
- <!-- from a statistical perspective, rather an engineering perspective -->
-A state-space model is special case of a dynamic latent variable model, which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations.
-As such, a state-space model can be viewed as a statistical model of the problem at hand.
+Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model, which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations.
 
@@
 This acceleration results in a peak velocity $\omega_{peak} = AT/\pi$ and a final angular position $\theta_{final} = AT^2/2\pi$.
-Due to the inherent noisiness of the measurements, and the dynamic nature of the stimulus, the measurements need to be interpreted according to the context in which they occur.
+Particle filtering provides a means of filtering out the noise inherent in the measurements, and allows estimation of evolution of the unobservable, latent state variables.
 
@@
 
-# motiondata_hidden <- motiondata %>%
-#     gather(type, value = value, -time) %>%
-#     filter(type != "observations") %>%
-#     mutate(type = factor(type,
-#                          levels = c("position",
-#                                     "velocity",
-#                                     "acceleration")))
-# 
-# p1 <- motiondata_hidden %>%
-#         ggplot(aes(x = time, y = value)) +
-#             geom_line(aes(linetype = type), size = 1.4) +
-#             geom_point(data = motiondata,
-#                        aes(x = time, y = observations),
-#                        shape = 21, colour = "white",
-#                        fill = "white", size = 6, stroke = 2) +
-#             geom_point(data = motiondata,
-#                        aes(x = time, y = observations),
-#                        shape = 21, colour = "black",
-#                        fill = "white", size = 4, stroke = 2) +
-#             guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
-#             xlab("Time (sec)") +
-#             ylab("Angular position (deg)") +
-#             theme_apa() +
-#             theme(legend.title=element_blank())
-# p1
+motiondata_hidden <- motiondata %>%
+    gather(type, value = value, -time) %>%
+    filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration")))
+
+p1 <- motiondata_hidden %>%
+        ggplot(aes(x = time, y = value)) +
+            geom_line(aes(linetype = type), size = 1.4) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "white",
+                       fill = "white", size = 6, stroke = 2) +
+            geom_point(data = motiondata,
+                       aes(x = time, y = observations),
+                       shape = 21, colour = "black",
+                       fill = "white", size = 4, stroke = 2) +
+            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+            xlab("Time (sec)") +
+            ylab("Angular position (deg)") +
+            theme_apa() +
+            theme(legend.title=element_blank())
+p1
 ```
@@
 
-# TODO
 How do we explain higher-level anticipation/expectations in such a model?

### 2017-01-27 notes.Rmd
draft introduction

@@
 
-interference arises when he brain, due to task demands, uses the control mechanism in a manner which is not entirely counterfactual
+__interference arises when he brain, due to task demands, uses the control mechanism in a manner which is not entirely counterfactual__
 
 # two
-interference is surprising, because a mental simulation should be counterfactual.
+~~Interference is surprising, because a mental simulation should be counterfactual.
 if it isn't this might be a problem.
-why should thinking about rotating left have an effect on your ability to sense leftward motion?
+why should thinking about rotating left have an effect on your ability to sense leftward motion?~~
 
-we show this type of interactions needs to be investigate pdf in a probabilistic framework
+~~we show this type of interactions needs to be investigate pdf in a probabilistic framework~~
 
@@
 
-in any case, we need a closer look at the connection between menral simulation and planning and the ability to predict self initiated movements.
+in any case, we need a closer look at the connection between mental simulation and planning and the ability to predict self initiated movements.
 

### 2017-01-27 simulations/01-3d-statespace.R

@@ -39,7 +39,7 @@ prec_p_init <- sd2precision(1)
 
 prec_a <- sd2precision(0.5)
 prec_w <- sd2precision(0.5)
-prec_p <- sd2precision(1.0)
+prec_p <- sd2precision(0.3)
 prec_y <- sd2precision(1.7)
 
 # f_eval <- function(x, k) { .5 * x + 25*x/(1+x^2) + 8*cos(1.2*k) }

### 2017-01-27 manuscript.Rmd
Word export is broken. need to fix.

@@
 output:
-     # papaja::apa6_word:
+     # papaja::apa6_word
      papaja::apa6_pdf
@@
 The aim of this paper is to specifically investigate the generative model that may be used for mental spatial transformations.
-The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Baysian inference [Oman, CITATIONS, @PaulRMacNeilage:2008bv].
+The ability to run embodied simulations is often claimed to be grounded in the relevant sensorimotor circuits, and we claim that the vestibular system is ideally suited to investigate this connection, because the generative models used by the vestibular system for dynamic sensory inference have been extensively investigated using approaches from optimal control theory, and more recently, dynamic Bayesian inference [Oman, CITATIONS, @PaulRMacNeilage:2008bv].
 
@@
 The predictive processing framework has been applied to almost all sensory modalities [CITATIONS: pain, proprioception, auditory , visual, etc.], and the picture that has emerged in recent years is that the brain employs sophisticated inference algorithms in order to "interpret" the stream of incoming sensory data.
-The brain seems to be actively engaged in formulating hypotheses, and testing these on sensory data; this has been extended to the concept of active inference [Friston], in which an agent attempts to either change it's internal model of world, or to interven on the world in order to fulfil sensory predictions.
+The brain seems to be actively engaged in formulating hypotheses, and testing these on sensory data; this has been extended to the concept of active inference [Friston], in which an agent attempts to either change it's internal model of world, or to intervene on the world in order to fulfill sensory predictions.
 
 We claim that the vestibular modality is particularly well-suited to investigating predictive processing at multiple hierarchical scales, from the lowest levels of sensory inference to higher levels, where goals, intentions and beliefs of the organism become relevant.
-A case in point is the study by @Wertheim:2001cb, who provided evidence that participants' perceptions of tilt induced by vestibular stimulation are shaped to a large extent by the beliefs about possible motions they formed prior to taking part in the experiement.
+A case in point is the study by @Wertheim:2001cb, who provided evidence that participants' perceptions of tilt induced by vestibular stimulation are shaped to a large extent by the beliefs about possible motions they formed prior to taking part in the experiment.
 <!-- [ok] -->
@@
 Interference effects between higher-level cognitive processes, such as SPT and mental imagery, and sensory inference, are thought to occur as a result of shared circuitry [@Deroualle:2014ho, @Nigmatullina:2015gh].
-However, the precise nature of such interactions, and what exaclty might be shared, is left unspecified.
+However, the precise nature of such interactions, and what exactly might be shared, is left unspecified.
 At first glance, it appear unlikely that such interference should be possible, as low levels of vestibular processing appear to be cognitively impenetrable [CITATION].
@@
 Thus, seemingly reflexive, and impenetrable, motor actions require a large amount of flexibility based on top-down information.
-Even accepting the fact that low level sensorimotor processing may be more cognitively penetrable than previously thought, the fact that imagining a rotation of the self should have an effect on a person's ability to detect rotations, or to discriminiate between directions of rotation, still requires an explanation.
+Even accepting the fact that low level sensorimotor processing may be more cognitively penetrable than previously thought, the fact that imagining a rotation of the self should have an effect on a person's ability to detect rotations, or to discriminate between directions of rotation, still requires an explanation.
 
@@
 
-We proceed by introducing recent models of Bayesian inference for real-time, dynamic sensory processing of vestibular data, based on particle filters [@Laurens:2006be, @Karmali:2012cv], and we reframe these as probabilistic graphical models.
+We proceed by introducing recent models of Bayesian inference for real-time, dynamic sensory processing of vestibular data, based on particle filters [@Laurens:2006be, @Karmali:2012cv], and we re-frame these as probabilistic graphical models.
 This allows us to investigate the generative model separately from the algorithm used for inference, and we then extend the basic model for sensory inference in order to focus specifically on the ability to model the consequences of self-initiated, active movements.
@@
 This is generally seen as evidence for the existence of an internal model in the processing of vestibular signals.
-@Karmali:2012cv focussed on one-dimensional angular velocity during yaw rotations; their model was able to explain the phenomenon of velocity storage as a consequence of particle filtering, with the time constant of velocity storage depending on the noise of the sensory afferent signals.
+@Karmali:2012cv focused on one-dimensional angular velocity during yaw rotations; their model was able to explain the phenomenon of velocity storage as a consequence of particle filtering, with the time constant of velocity storage depending on the noise of the sensory afferent signals.
 
@@
 We proceed by providing a brief introduction to the concept of state-space models, and particle filtering, which is a particular algorithm that can be used to perform inference in state-space models.
-State-space models can be seen as special cases of more general Bayesian hierarchical models [@Bishop:2006ui; @Murphy:2012ua].
-Following the same approach as @Penny:2013iv, we first reframe the generative model used for sensory inference as a dynamic probabilistic graphical model, which is suitable for modelling the time-varying behaviour of a system in terms of latent, unobservable processes and noisy observations.
+State-space models can be seen as special cases of more general dynamic probabilistic graphical models [@Bishop:2006ui; @Murphy:2012ua], and are suitable for modelling the time-varying behaviour of a system in terms of a latent, unobservable process and noisy measurements.
 
 For example, consider trying to estimate the velocity and position during a leftward head turn.
-The afferent SCC signals are considered to be proportional to the angular velocity of the head with a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv].
-The measurements are therefore observations of the true, unobservable angular velocity; however, they are contaminted by Gaussian noise, which is determined by characteristics of the sensory organs.
-Figure \@ref(fig:head-turn-data) shows a sequence of noisy measurements during a two-second turn of the head.
+In this case, the latent process provides a model of the kinematics of the system, and the measurement model describes how the afferent SCC signals are generated, dependent on the kinematics.
+
+The SCC signals are considered to be proportional to the angular velocity of the head within a frequency band of approximately 0.1 – 10 Hz, which constitutes the range of natural head movements [@Grabherr:2008kk; @Carriot:2014fv].
+These measurements, however, are contaminated by Gaussian noise, which is determined by characteristics of the sensory organs.
+Figure \@ref(fig:head-turn-data) shows a sequence of noisy measurements during a two-second leftward turn of the head.
 The acceleration generating the movement is sinusoidal, given for each time step $t$ by $Asin(2\pi t/T)$, where $A$ is the amplitude of the movement, $T = 1/f$ is the duration of the movement, and $f$ is the frequency.
-This acceleration results in a peak velocity $\omega_{peak} = AT/\pi$ and a final angular position $\theta_{final} = AT^2/2\pi$.
+This acceleration results in a half-cycle cosine velocity profile, with a peak velocity $\omega_{peak} = AT/\pi$ and a final angular position $\theta_{final} = AT^2/2\pi$.
 Particle filtering provides a means of filtering out the noise inherent in the measurements, and allows estimation of evolution of the unobservable, latent state variables.
@@
 ```{r head-turn-data, echo=FALSE, fig.cap= "Sensory measurements obtained during a short leftward head turn.
-The goal is to infer the latent angular velocity and position, based on the noisy measurements.
-\\label{fig:head-turn}", message=FALSE, warning=FALSE, dpi=100}
+The goal is to infer the latent angular velocity and position, based on the noisy measurements.", message=FALSE, warning=FALSE, dpi=100}
 dt <- 0.1
@@
                              sensor_sd = 2.0)
-plot_trajectories(motiondata = motiondata)
+# plot_trajectories(motiondata = motiondata)
 
@@
                                     "acceleration")))
+df_tmp <- data_frame(time = motiondata$time,
+                     observations = motiondata$observations)
+df_tmp$param <- as.factor("Measurements")
 
@@
                        fill = "white", size = 6, stroke = 2) +
-            geom_point(data = motiondata,
-                       aes(x = time, y = observations),
-                       shape = 21, colour = "black",
+            geom_point(data = df_tmp,
+                       aes(x = time, y = observations, shape = param),
+                       colour = "black",
                        fill = "white", size = 4, stroke = 2) +
             guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+            scale_shape_manual(name = element_blank(),
+                               labels = "Measurements",
+                               values = 21) +
             xlab("Time (sec)") +
@@
             theme_apa() +
-            theme(legend.title=element_blank())
+            theme(legend.title=element_blank(), 
+                  panel.border = element_blank(), 
+                  axis.line = element_line(colour="black", 
+                                           size = 0.75, lineend="square"))
 p1
@@
 
-The key idea is that we can model the dynamics of a system by the evolution of the latent state variables, representing the velocity and position, as a first-order Markov process, where the state at each time point $t$ depends only on the previous state at time $t-1$.
+The dynamics of the system can be modelled by the evolution of the 3-d latent state variable, representing the position, velocity and acceleration, as a first-order Markov process, where the state at each time point $t$ depends only on the previous state at time $t-1$.
 This is referred to as the process model.
-Estimating the velocity $\omega$ and position $\theta$ in discrete time requires writing the kinematic equations as a set of first order difference equations.
+Estimating the position $\theta$, velocity $\omega$ and acceleration $\alpha$ in discrete time requires writing the kinematic equations as a set of first order difference equations.
 
-$$\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t$$
-$$\omega_{t} = \omega_{t-1}$$
+$$\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2} \Delta t^2$$
+$$\omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t$$
+$$\alpha_t = \alpha_{t-1}$$
 
-Note that we have omitted the acceleration from the process model, for the sake of simplicity; this amounts to assuming a constant velocity model, which just means that we expect the current velocity to be same as the previous velocity.
+This model assumes a constant acceleration.
 These kinematic equations form a linear model, and could optionally be written in matrix notation.
-For further simplicity, we retain the current notation.
+For simplicity, we retain the current notation.
 Next, we consider the measurements at each time step $t$ as depending only on the current state.
@@
 The above equations represent the expected values of the state and observation variables.
-In order to complete the generative model, we require a description of the stochastic nature dependancies.
-We model both process and measurement noise a being normally distributed, leading to the following linear Gaussian state-space model:
+In order to complete the generative model, we require a description of the stochastic nature of the dependencies.
+We model both process and measurement noise as being normally distributed, leading to the following linear Gaussian state-space model:
 
@@
 $$y_{t} = g(x_t, \delta)$$
-$$\epsilon \sim N(0, \sigma^2_\epsilon) $$
-$$\delta \sim N(0, \sigma^2_\delta) $$
-
-where the state variable $x_t$ consists of both the velocity $\omega_t$ and position $\theta_t$, and $y_{t}$ denotes the sensory measurement.[^sensormodel] $\epsilon$ and $\delta$ are zero-mean Gaussian noise terms with variances $\sigma^2_\epsilon$ and $\sigma^2_\delta$.
+$$\epsilon \sim N(0, \sigma^2_\epsilon)$$
+$$\delta \sim N(0, \sigma^2_\delta)$$
+<!-- TODO: it is better to start with the generative model, f() and g(), and then implement the state space description -->
+where the state variable $x_t$ consists of the position $\theta_t$, velocity $\omega_t$ and acceleration $\alpha_t$, and $y_{t}$ denotes the sensory measurement.
+<!-- [^sensormodel]  -->
+$\epsilon$ and $\delta$ are zero-mean Gaussian noise terms with variances $\sigma^2_\epsilon$ and $\sigma^2_\delta$.
 These refer to the process and measurement noise, respectively.
-The process noise represents the uncertainty in the dependance of the current state on the previous state; under the constant velocity assumption, the process noise must be large enough if the model is to allow for change in velocity due to unmodelled acceleration.
+The process noise represents the uncertainty in the dependence of the current state on the previous state; under the constant velocity assumption, the process noise must be large enough if the model is to allow for change in velocity due to unmodelled acceleration.
 Note that the measurement noise is a model of the noise characteristics of the sensor, and not a physical property of the sensor itself.
 The function $f$ and $g$ represent the process transition function and the measurement function, respectively.
-In the case of a linear dynamical system, the transition function can be written as:
-$$x_t \sim N(Ax_{t-1}), \epsilon^2)$$,
-where the matrix $A$ replaces the transition function $f$.
+<!-- In the case of a linear dynamical system, the transition function can be written as: -->
+<!-- $$x_t \sim N(Ax_{t-1}), \epsilon^2)$$, -->
+<!-- where the matrix $A$ replaces the transition function $f$. -->
 
-[^sensormodel]: Note that we although both @Laurens:2006be, @Karmali:2012cv consider a realistic sensor model, for simplicity, we implement the sensor as being a noisy realisation of the velocity.
+<!-- [^sensormodel]: Note that we although both @Laurens:2006be, @Karmali:2012cv consider a realistic sensor model, for simplicity, we implement the sensor as being a noisy realisation of the velocity.
 Our model is not intended to be a realistic model of the SCC; rather, we focus on the probabilistic computations relevant for higher-level cognition.
+-->
 
@@
 Figure \@ref(fig:graphical-model-inference) shows a representation of a state-space model as a graphical model, unrolled over multiple time steps.
-Starting at time 0, the state evolves according to the process model, with the measurements at each time step being generated according to the measurement model.
-The 2-dimensiontal latent state variable is unshaded, to indicate that it cannot be observed, wheres the observed sensory measurements are shaded.
-Whereas previous models were discussed in terms of the particular filtering algorithm used, the graphical notation highlights the fact that a state-space model can be viewed as a dynamic latent factor model [@Blei:2014cp].
+The state evolves according to the process model, with the measurements at each time step being generated according to the measurement model.
+The 3-dimensional latent state variable is unshaded, to indicate that it cannot be observed, wheres the observed sensory measurements are shaded.
+<!-- Whereas previous models were discussed in terms of the particular filtering algorithm used, the graphical notation highlights the fact that a state-space model can be viewed as a dynamic latent factor model [@Blei:2014cp].
 This opens up the possibility of extending the graphical model to include higher-level variables.
+-->
 Given a generative model, we can apply various sampling algorithms, either in order to run a Monte Carlo simulation by sampling from the prior without incorporating any observed variables, or to perform inference, by conditioning on observed variables.
@@
 
-We give a brief overview of particle filtering, an sequential Monte Carlo algorithm for performing approximate Bayesian inference in a graphical model.
-We descibe a simple bootstrap particle filter, concentrating on those aspects most relevant for this study.
+We give a brief overview of particle filtering, a sequential Monte Carlo algorithm for performing approximate Bayesian inference in a graphical model.
+We describe a simple bootstrap particle filter, concentrating on those aspects most relevant for this study.
 For a more thorough treatment, see @Doucet:2000bh and @Doucet:2009us.
 @Speekenbrink:2016kc provides a recent review with applications in psychological research.
-Particle filtering is .
-It proceeds by recursively estimating the distribution of the current state, $P(x_t | y_{1:t})$, as a function of its parent nodes, conditional upon the sequence of measurements $y_{1:t}$.
+Particle filtering proceeds by recursively estimating the distribution of the current state, $P(x_t | y_{1:t})$, as a function of its parent nodes, conditional upon the sequence of measurements $y_{1:t}$.
 The distributions are represented by a collection of discrete weighted samples, known as particles.
@@
 In other words: first, a hypothesis about the expected state is generated, and then this hypothesis is tested by evaluating how well it predicts the data.
-The algorithm starts with a prior distribution of the particles, and at each subsequent time step, proceeds by first sampling from the so-called proposal distribution $P(x_t | y_{1:t-1})$, which is based on the specific choice of transition function, and therefore reflects the probabilistic causal model of the sensory measurements.
+The algorithm starts with a prior distribution of the particles, and at each subsequent time step, proceeds by first sampling from the proposal distribution $P(x_t | y_{1:t-1})$, which is based on the specific choice of transition function.
 Next, during the updating step, the current measurement is incorporated by evaluating the likelihood, $P(y_t | x_t)$, which gives the probability of observing the measurement for each individual particle.
@@
 They are equally weighted, indicated by their diameters.
-During the update step, the particle are reweighted, with each particle’s weights reflecting how well it explains the measurement.
+During the update step, the particle are re-weighted, with each particle’s weights reflecting how well it explains the measurement.
 This is visualized by the change in diameter; those particles lying close to the observation (sensory input from the horizontal canal) have a larger diameter.
@@
 The uncertainty of the belief, as indicated by the particle cloud's standard deviation (light grey lines), increases at every prediction step, and subsequently decreases after updating.
-This cycle of prediction, updating and resampling ensures that the collection of particles comes to represent the 'true' filtering distribution, by eliminating those particles with low weights.
+This cycle of prediction, updating and resampling ensures that the collection of particles comes to represent the 'true' posterior distribution, by eliminating those particles with low weights.
 The particle filter represents all beliefs in the form of probability distributions; the mean and variance of the belief distribution can be computed at each time step, and the logarithm of the weights $log(\frac{1}{N}\sum^N_{i=1}w_i)$, averaged over all $N$ particles, may be summed over all time steps in order to compute the log-likelihood $\sum^T_{t=1}\Big[log(\sum^N_{i=1}w_i) - log(N)\Big]$, which can be used in order to evaluate how well the model was able to fit the sensory evidence.
@@
 
-Figure \@ref(fig:passive) shows the result of applying the particle filtering algorithm to the graphical model depicted in Figure \@ref(fig:graphical-model-inference), in order to infer the velocity and position during a two-second leftward head turn.
+Figure \@ref(fig:3d-statespace-fits) (a) shows the result of applying the particle filtering algorithm to the graphical model depicted in Figure \@ref(fig:graphical-model-inference) and simulated measurements, in order to infer the velocity and position during a two-second leftward head turn.
+The acceleration is also inferred, but is not of interest.
 
-```{r passive, cache=TRUE, fig.cap="Particle filter for inferring velocity and position during a two-second head turn.", echo=FALSE, message=FALSE, warning=FALSE}
+```{r 3d-statespace, fig.cap="Particle filter with a 3-d state for inferring velocity and position during a two-second head turn.", message=FALSE, warning=FALSE, include=FALSE}
+
+load("simulations/model-fits/m1_3d_head_turn.Rda")
+
+df_w <- with(m1_3d_head_turn$summ_smc, {
+    data_frame(t = 1:length(w$f$mean),
+               mean = w$f$mean,
+               lower = w$f$quant$`0.025`,
+               upper = w$f$quant$`0.975`)
+})
+df_w$param <- "Velocity"
+df_w$true <- m1_3d_head_turn$model$data()$w_true
+df_w$observations <- m1_3d_head_turn$model$data()$y
+
+df_p <- with(m1_3d_head_turn$summ_smc, {
+    data_frame(t = 1:length(p$f$mean),
+               mean = p$f$mean,
+               lower = p$f$quant$`0.025`,
+               upper = p$f$quant$`0.975`)
+})
+df_p$param <- "Position"
+df_p$true <- m1_3d_head_turn$model$data()$p_true
+df_p$observations <- m1_3d_head_turn$model$data()$y
+
+df <- bind_rows(df_p, df_w)
+
+p1 <- df %>%
+    filter(param == "Position") %>%
+    ggplot(data = ., aes(x = t)) +
+    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
+
+    geom_line(aes(y = true), alpha = 0.9,
+              linetype = 4, size = 1.2) +
+
+    geom_line(aes(y = mean), size = 1.6) +
+    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
+    scale_x_continuous(limits = c(1, 20),
+                       breaks = c(0, 5, 10, 15, 20),
+                       labels = c("0", "0.5", "1", "1.5", "2")) +
+    # scale_y_continuous(limits = c(-5, 20)) +
+    # xlab("Time (sec)") + 
+    ylab("Angular position (deg)") +
+    papaja::theme_apa(base_size = 18) +
+    theme(axis.title.x=element_blank()) +
+    theme(panel.border = element_blank(), 
+          axis.line = element_line(colour="black", size= 0.75, lineend="square"))
+
+p2 <- df %>%
+    filter(param == "Velocity") %>%
+    ggplot(data = ., aes(x = t)) +
+    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
+
+    geom_line(aes(y = true), alpha = 0.9,
+              linetype = 4, size = 1.2) +
+
+    geom_line(aes(y = mean), size = 1.6) +
+    geom_point(aes(y = observations),
+               alpha = 1, fill = "white",
+               colour = "white",
+               shape = 21, size = 6) +
+    geom_point(aes(y = observations), alpha = 1,
+               fill = "black",
+               colour = "grey40", shape = 21, size = 4) +
+    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
+    scale_x_continuous(limits = c(1, 20),
+                       breaks = c(0, 5, 10, 15, 20),
+                       labels = c("0", "0.5", "1", "1.5", "2")) +
+    # scale_y_continuous(limits = c(-5, 20)) +
+    xlab("Time (sec)") + 
+    ylab("Angular Velocity (deg/s)") +
+    papaja::theme_apa(base_size = 18) +
+    theme(panel.border = element_blank(), 
+          axis.line = element_line(colour="black", size = 0.75, lineend="square"))
+
+p_3d_1 <- cowplot::plot_grid(p1, p2, ncol = 1)
+                   # labels = c("Position", "Velocity"),
+                   # label_size = 12)
+```
+
+```{r passive, eval=FALSE, fig.cap="Particle filter for inferring velocity and position during a two-second head turn.", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
 params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
@@
 
-The measurement are shown as squares, and the true velocity that generated the data is shown as a dotted line.
+The measurements are shown as circles, and the true velocity that generated the data is shown as a dotted line.
 This velocity is the result of an external force being applied.
-The solid lines depict the estimated velocity (left) and position (right), and the shaded areas show the $95\%$ credible intervals.
-The model is able to estimate the latent state despite making the assumption that the velocity is constant.
-It does this by allowing for random changes in velocity by letting the process noise be sufficiently large.
-Note that both velcoity and position estimates lag the true values; this is due the fact that the acceleration is not modelled explicitely, and thus the system is slow to react to changes in velocity.
-
-This graphical model assumes that the velocity is constant, and does not  the explicitly estimate acceleration; however, it allows for random changes in its velocity by in terms of the process noise.
+The solid lines depict the mean estimates, and the shaded areas show the $95\%$ credible intervals.
+The model is able to estimate the latent state despite making the assumption that the acceleration is constant.
+It does this by allowing for random changes in acceleration by letting the process noise be sufficiently large.
+Note that both position and velocity estimates lag the true values; this is due the fact that the acceleration is expected to be constant, and thus the system is slow to react to changes.
 The estimation can be improved by increasing the process noise.
-Figure \@ref(fig:passive-2) shows the resulting estimates.
-Whilst the lag is somewhat reduced, a consequence of increasing the process noise is that the measurement are given more weight, resulting in a tendency to overfit the noise in the measurements.
+Figure \@ref(fig:3d-statespace-fits) (b) shows the resulting estimates.
+Whilst the lag is reduced, a consequence of increasing the process noise is that the measurements are given more weight, resulting in a tendency to over-fit the noise in the measurements.
 The constant velocity (random acceleration) model is unable to capture the underlying smooth trajectory of the velocity and position, instead showing random jumps in the velocity estimate.
 
-```{r passive-2, cache=TRUE, fig.cap="Particle filter for inferring velocity and position during a two-second head turn with increased process noise.", echo=FALSE, message=FALSE, warning=FALSE}
+```{r 3d-statespace-2, fig.cap="Particle filter with increased process noise.", message=FALSE, warning=FALSE, include=FALSE}
+
+load("simulations/model-fits/m1_3d_head_turn_2.Rda")
+
+df_w <- with(m1_3d_head_turn_2$summ_smc, {
+    data_frame(t = 1:length(w$f$mean),
+               mean = w$f$mean,
+               lower = w$f$quant$`0.025`,
+               upper = w$f$quant$`0.975`)
+})
+df_w$param <- "Velocity"
+df_w$true <- m1_3d_head_turn_2$model$data()$w_true
+df_w$observations <- m1_3d_head_turn_2$model$data()$y
+
+df_p <- with(m1_3d_head_turn_2$summ_smc, {
+    data_frame(t = 1:length(p$f$mean),
+               mean = p$f$mean,
+               lower = p$f$quant$`0.025`,
+               upper = p$f$quant$`0.975`)
+})
+df_p$param <- "Position"
+df_p$true <- m1_3d_head_turn_2$model$data()$p_true
+df_p$observations <- m1_3d_head_turn_2$model$data()$y
+
+df <- bind_rows(df_p, df_w)
+
+p1 <- df %>%
+    filter(param == "Position") %>%
+    ggplot(data = ., aes(x = t)) +
+    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
+
+    geom_line(aes(y = true), alpha = 0.9,
+              linetype = 4, size = 1.2) +
+
+    geom_line(aes(y = mean), size = 1.6) +
+    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
+    scale_x_continuous(limits = c(1, 20),
+                       breaks = c(0, 5, 10, 15, 20),
+                       labels = c("0", "0.5", "1", "1.5", "2")) +
+    # scale_y_continuous(limits = c(-5, 20)) +
+    # xlab("Time (sec)") + 
+    ylab("Angular position (deg)") +
+    papaja::theme_apa(base_size = 18) +
+    theme(axis.title.x=element_blank()) +
+    theme(panel.border = element_blank(), 
+          axis.line = element_line(colour="black", size= 0.75, lineend="square"))
+
+p2 <- df %>%
+    filter(param == "Velocity") %>%
+    ggplot(data = ., aes(x = t)) +
+    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
+
+    geom_line(aes(y = true), alpha = 0.9,
+              linetype = 4, size = 1.2) +
+
+    geom_line(aes(y = mean), size = 1.6) +
+    geom_point(aes(y = observations),
+               alpha = 1, fill = "white",
+               colour = "white",
+               shape = 21, size = 6) +
+    geom_point(aes(y = observations), alpha = 1,
+               fill = "black",
+               colour = "grey40", shape = 21, size = 4) +
+    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
+    scale_x_continuous(limits = c(1, 20),
+                       breaks = c(0, 5, 10, 15, 20),
+                       labels = c("0", "0.5", "1", "1.5", "2")) +
+    # scale_y_continuous(limits = c(-5, 20)) +
+    xlab("Time (sec)") + 
+    ylab("Angular Velocity (deg/s)") +
+    papaja::theme_apa(base_size = 18) +
+    theme(panel.border = element_blank(), 
+          axis.line = element_line(colour="black", size = 0.75, lineend="square"))
+
+p_3d_2 <- cowplot::plot_grid(p1, p2, ncol = 1) 
+                   # labels = c("Position", "Velocity"),
+                   # label_size = 12)
+
+```
+```{r passive-2, eval=FALSE, fig.cap="Particle filter for inferring velocity and position during a two-second head turn with increased process noise.", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
 params <- list(sdx = c(1.8, 1.8) * eye(2), sdy = 2.0,
@@
 
-```{r eval=FALSE, include=FALSE}
-# plot_pf_estimates(object = passive, predict = FALSE)
+```{r 3d-statespace-fits, echo=FALSE, fig.cap="(a) Particle filter with a 3-d state space model for inferring velocity and position during a two-second head turn.
+(b) The same particle filter with increased process noise.", message=FALSE, warning=FALSE}
+cowplot::plot_grid(p_3d_1, p_3d_2, ncol = 2)
 ```
 
-In order to imporove the model, the acceleration can be included in the latent state, resulting in a 3-dimensional latent state variable:
-
-$$ \theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2}\Delta t^2 $$
-
-$$ \omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t $$
-$$ \alpha_t = \alpha_{t-1} $$
-
-The effect that this has in terms of the inference algorithm is that the proposal distribution $P(x_t | y_{1:t})$, from which samples are drawn from during the 'predict' step, is improved.
-By including the acceleration in the latent state, the the changes in velocity are modelled.
-Of course, the acceleration itself is not modelled, and the changes in acceleration (jerk) are thus treated as random, resuling in a random jerk model.
-
-> QUESTION: shall we show a model that includes acceleration in its state?
-
-How do we explain higher-level anticipation/expectations in such a model?
-This form of anticipation is not the same as the 1-step ahead predictions made during inference, but rather implement higher level prior beliefs about the direction of movement, for example.
-
-An obvious way of including this is the mechanism that allows one to predict the sensory consequences of self-initiated movement via an efference copy.
-
-<!-- ## Performance for various types of natural head movements -->
-
-```{r shaking-head, eval=FALSE, fig.cap="Particle filter applied to ", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
-dt <- 0.1
-head_shaking_data <- generate_data(T = 2, frequency = 2.0,
-                             amplitude = 20,
-                             sensor_sd = 2.0)
-
-# plot_trajectories(motiondata = abrupt_stop_data)
-params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
-                x_init = c(0, 0),
-                sdx_init = 0.5*eye(2),
-                A = 20, N = 1000,
-                dt = 0.1, freq = 0.5,
-                transfun = transfun)
-
-head_shaking <- run_particle_filter(motiondata = head_shaking_data, 
-                                    params = params,
-                                    selfmotion = "passive")
-
-plot(head_shaking, panel_label = "", print_loglik = FALSE, predict = FALSE, imagine = FALSE)
-```
-
-```{r two-external-forces, eval=FALSE, fig.cap="Particle filter applied to ", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
-dt <- 0.1
-ext_data_1 <- generate_data(T = 4, frequency = 0.7,
-                             amplitude = 20,
-                             sensor_sd = 2.0)
-ext_data_2 <- generate_data(T = 4, frequency = 0.5,
-                             amplitude = -5,
-                             sensor_sd = 2.0)
-superposition_data <- ext_data_1 + ext_data_2
-
-# plot_trajectories(motiondata = abrupt_stop_data)
-params <- list(sdx = c(0.8, 0.8) * eye(2), sdy = 2.0,
-                x_init = c(0, 0),
-                sdx_init = 0.5*eye(2),
-                A = 20, N = 1000,
-                dt = 0.1, freq = 0.5,
-                transfun = transfun)
-
-superposition <- run_particle_filter(motiondata = superposition_data, 
-                                    params = params,
-                                    selfmotion = "passive")
-
-plot(superposition, panel_label = "", print_loglik = FALSE, predict = FALSE, imagine = FALSE)
-```
+<!-- A well known shortcoming of the model described above is that the proposal distribution is equal to the prior distribution [@VanDerMerwe:2000vx] -->
 
 # Extension allowing the prediction of active movements
-While the model described above can be used in order to track the behaviour of a dynamic system by including the acceleration as a latent state, it does not include any means of incorporating an active prediction of the kinematics.
+
 During sensory inference the measurements are incorporated, and and a posterior estimate of the latent state is obtained by updating the prior with the likelihood.
 The underlying kinematic model thus relies on the sensory data to inform its estimates.
-While this model is perfectly suitable for sensory inference, for the task of predicting the sensory consequences of self-initiated movemenents, the model is inadequate, as it does not allow including further knowledge about kinematics, which can be used to improve the proposal distribution.
+This model can be used in order to track the behaviour of a dynamic system by including the acceleration as a latent state, it is purely data-driven, and does not include any means of incorporating an active prediction of the kinematics.
+Thus, it cannot explain top-down expectations or biases.
+This form of anticipation is not the same as the 1-step ahead predictions made during inference, but rather implements higher level prior beliefs about the direction of movement.
+For example, imagine seeing that someone is going to push you.
+You can use this information, derived from the visual modality, in order prepare yourself in order to compensate for this.
+However, you do not want to act too early, but rather wait until you are actually moving.
+
+> While this model is perfectly suitable for sensory inference, for the task of predicting the sensory consequences of self-initiated movemenents, the model is inadequate, as it does not allow including further knowledge about kinematics, which can be used to improve the proposal distribution.
 This knowledge is obtained by means of an efference copy of executed or planned motor commands, which is sent to a forward model.
@@
 
-__A further issue is that the sensory inference model cannot incorporate any top-down expectations or biases.
-For example, imagine seeing that someone is going to push you.
-You can use this information, derived from the visual modality, in order prepare yourself in order to compensate for this.
-However, you do not want to act too early, but rather wait until you are actually moving.__
-
-In terms of state-space models, this is implemented as a control input $u$, which extends the graphical model shown in Figure \@ref(fig:graphical-model-inference).
+An obvious way of including this is the mechanism that allows one to predict the sensory consequences of self-initiated movement via an efference copy.
+In state-space models, this is implemented as a control input $u$, which extends the graphical model shown in Figure \@ref(fig:graphical-model-inference).
 In this model, the estimated acceleration is replaced by an known acceleration signal.
@@
 Assuming the model has full knowledge about the onset of movement, and the type of movement (sinusoidal head turn), the model requires information about the frequency, $f = 1/T$, amplitude, $A$, and direction, which is given by the sign of the amplitude.
-These are
 
@@
 
+# __BIS HIER.__
+
 # Modelling questions
@@
 
-# An Investigation of Various Priors
+# An investigation of various priors and predictions
 Here, we identify the various forms of predictive processing in the context of real-time sensory inference using SMC
 
-- Related to Filtering Algorithm
-- Higher-level Priors: The process model determines the kind of motion that is expected. Call this _hypothesis-driven_ sensory inference. It implements an expectation and is a strong prior.
+- Related to filtering algorithm
+- Higher-level priors: The process model determines the kind of motion that is expected. Call this _hypothesis-driven_ sensory inference. It implements expectation on longer time-scales.
 
@@
 - parameters of the motion
-
-control input allows inferring the consequences of one's own actions.
 
@@
 # Simulations
-
+> Should we have a separate simulations section, or should the simulations be distributed?
 ## Model specifications
-
-## 3-d state space model
-
-Figure \@ref(fig:3d-statespace) show a 3-d state space model
-
-```{r 3d-statespace, echo=FALSE, fig.cap="3-d state space model.", message=FALSE, warning=FALSE}
-
-load("simulations/model-fits/m1-3d-statespace.Rda")
-
-df_w <- with(m1_3d_statespace$summ_smc, {
-    data_frame(t = 1:length(w$f$mean),
-               mean = w$f$mean,
-               lower = w$f$quant$`0.025`,
-               upper = w$f$quant$`0.975`)
-})
-df_w$param <- "Velocity"
-df_w$true <- m1_3d_statespace$model$data()$w_true
-df_w$observations <- m1_3d_statespace$model$data()$y
-
-df_p <- with(m1_3d_statespace$summ_smc, {
-    data_frame(t = 1:length(p$f$mean),
-               mean = p$f$mean,
-               lower = p$f$quant$`0.025`,
-               upper = p$f$quant$`0.975`)
-})
-df_p$param <- "Position"
-df_p$true <- m1_3d_statespace$model$data()$p_true
-df_p$observations <- m1_3d_statespace$model$data()$y
-
-df <- bind_rows(df_p, df_w)
-
-p1 <- df %>%
-    filter(param == "Position") %>%
-    ggplot(data = ., aes(x = t)) +
-    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
-
-    geom_line(aes(y = true), alpha = 0.9,
-              linetype = 4, size = 1.2) +
-
-    geom_line(aes(y = mean), size = 1.6) +
-    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
-    scale_x_continuous(limits = c(1, 20),
-                       breaks = c(0, 5, 10, 15, 20),
-                       labels = c("0", "0.5", "1", "1.5", "2")) +
-    # xlab("Time (sec)") + 
-    ylab("Angular position (deg)") +
-    papaja::theme_apa(base_size = 18) +
-    theme(axis.title.x=element_blank()) +
-    theme(panel.border = element_blank(), 
-          axis.line = element_line(colour="black", size= 0.75, lineend="square"))
-
-p2 <- df %>%
-    filter(param == "Velocity") %>%
-    ggplot(data = ., aes(x = t)) +
-    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey90") +
-
-    geom_line(aes(y = true), alpha = 0.9,
-              linetype = 4, size = 1.2) +
-
-    geom_line(aes(y = mean), size = 1.6) +
-    geom_point(aes(y = observations),
-               alpha = 1, fill = "white",
-               colour = "white",
-               shape = 21, size = 6) +
-    geom_point(aes(y = observations), alpha = 1,
-               fill = "black",
-               colour = "grey40", shape = 21, size = 4) +
-    geom_hline(yintercept = 0, linetype = 2, color = "grey70") +
-    scale_x_continuous(limits = c(1, 20),
-                       breaks = c(0, 5, 10, 15, 20),
-                       labels = c("0", "0.5", "1", "1.5", "2")) +
-    xlab("Time (sec)") + 
-    ylab("Angular Velocity (deg/s)") +
-    papaja::theme_apa(base_size = 18) +
-    theme(panel.border = element_blank(), 
-          axis.line = element_line(colour="black", size = 0.75, lineend="square"))
-
-cowplot::plot_grid(p1, p2, ncol = 1) 
-                   # labels = c("Position", "Velocity"),
-                   # label_size = 12)
-
-```
 
@@
 
-- turing left
+- turning left
 - shaking head quickly
@@
 ## Expectations based on higher level priors
-In our model, there is no difference between modelling active self-motion, and an anticpation of passive self-motion due to predictable external forces. This is a testable hypothesis.
+In our model, there is no difference between modelling active self-motion, and an anticipation of passive self-motion due to predictable external forces. This is a testable hypothesis.
 
@@
 
-```{r active-1, cache=TRUE, fig.cap="Mis-specified proposal distribution.", echo=FALSE, message=FALSE, warning=FALSE}
+```{r active-1, eval=FALSE, fig.cap="Mis-specified proposal distribution.", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
 
@@
 - directional signal is not just a simple integration of velocity signals from the vestibular system
-- signals from motor efferecence copy or proprioceptive systems appear to be crucial for upating the preferred firing direction over time
+- signals from motor efferecence copy or proprioceptive systems appear to be crucial for updating the preferred firing direction over time
 

### 2017-01-28 manuscript.Rmd
word export fixed

@@
 output:
-     # papaja::apa6_word
-     papaja::apa6_pdf
+     papaja::apa6_word
+     # papaja::apa6_pdf
         # pandoc_args:
@@
 
-```{r preliminary, cache=FALSE, message = FALSE, warning = FALSE}
+```{r preliminary, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE}
 library(papaja)
@@
 
-```{r head-turn-data, echo=FALSE, fig.cap= "Sensory measurements obtained during a short leftward head turn.
-The goal is to infer the latent angular velocity and position, based on the noisy measurements.", message=FALSE, warning=FALSE, dpi=100}
+```{r head-turn-data, fig.cap="Sensory measurements obtained during a short leftward head turn.
+The goal is to infer the latent angular velocity and position, warning=FALSE, based on the noisy measurements.", message=FALSE, dpi=100, include=FALSE}
 dt <- 0.1
@@
 
-$$\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2} \Delta t^2$$
-$$\omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t$$
-$$\alpha_t = \alpha_{t-1}$$
+$$ \theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2} \Delta t^2 $$
+$$ \omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t $$
+$$ \alpha_t = \alpha_{t-1} $$
 
@@
 
-$$x_t = f(x_{t-1}, \epsilon)$$
-$$y_{t} = g(x_t, \delta)$$
-$$\epsilon \sim N(0, \sigma^2_\epsilon)$$
-$$\delta \sim N(0, \sigma^2_\delta)$$
+$$ x_t = f(x_{t-1}, \epsilon) $$
+$$ y_{t} = g(x_t, \delta) $$
+$$ \epsilon \sim N(0, \sigma^2_\epsilon) $$
+$$ \delta \sim N(0, \sigma^2_\delta) $$
+
 <!-- TODO: it is better to start with the generative model, f() and g(), and then implement the state space description -->
@@
 
-```{r 3d-statespace-fits, echo=FALSE, fig.cap="(a) Particle filter with a 3-d state space model for inferring velocity and position during a two-second head turn.
-(b) The same particle filter with increased process noise.", message=FALSE, warning=FALSE}
+```{r 3d-statespace-fits, fig.cap="(a) Particle filter with a 3-d state space model for inferring velocity and position during a two-second head turn.
+(b) The same particle filter with increased process noise.", message=FALSE, warning=FALSE, include=FALSE}
 cowplot::plot_grid(p_3d_1, p_3d_2, ncol = 2)

### 2017-01-28 manuscript.Rmd
converted to word and sent to Fred

@@
 
-title: "Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion"
+# title: "Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion"
 
-# title: "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
+title: "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
 # title: "Probabilistic Computations Underlying Simulated Self-Rotations"
@@
 
-```{r head-turn-data, fig.cap="Sensory measurements obtained during a short leftward head turn.
-The goal is to infer the latent angular velocity and position, warning=FALSE, based on the noisy measurements.", message=FALSE, dpi=100, include=FALSE}
+```{r head-turn-data, echo=FALSE, fig.cap="Sensory measurements obtained during a short leftward head turn.
+The goal is to infer the latent angular velocity and position, message=FALSE, warning=FALSE, based on the noisy measurements.", message=FALSE, dpi=100}
 dt <- 0.1
@@
 
-<!-- ![Particle filter computations performed at every time step.
+```{r particle-filter-explained, echo=FALSE, fig.cap="Particle filter computations performed at every time step.
 The cloud of particles (grey circles) represents the current belief state.
-Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model.
-This is equivalent to sampling from the prior distribution.
-Then, during the updating step, an observation  is obtained (shown as a black diamond).
-Each particle is assigned an importance weight, according to how well it explains the observation.
-Each particle's diameter reflects its importance weight.
-Subsequently, a new set of particles is formed by weighted multinomial resampling, with each particle's probability of being chosen proportional to its importance weight.
-The particle cloud's mean and standard deviation are shown as light grey circles and lines.
-This resulting cloud of particles represents the posterior distribution.
-This posterior subsequently forms the prior distribution for the next time step.](diagrams/figure-particle-filter-explained.pdf){#fig:particle-filter-explained width=100%} -->
-
-```{r particle-filter-explained, fig.cap="Particle filter computations performed at every time step.
-The cloud of particles (grey circles) represents the current belief state.
-Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model.
+Initially, message=FALSE, warning=FALSE, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model.
 This is equivalent to sampling from the prior distribution.
@@
 
-```{r 3d-statespace-fits, fig.cap="(a) Particle filter with a 3-d state space model for inferring velocity and position during a two-second head turn.
-(b) The same particle filter with increased process noise.", message=FALSE, warning=FALSE, include=FALSE}
-cowplot::plot_grid(p_3d_1, p_3d_2, ncol = 2)
+```{r 3d-statespace-fits, echo=FALSE, fig.cap="(a) Particle filter with a 3-d state space model for inferring velocity and position during a two-second head turn.
+(b) The same particle filter with increased process noise.", message=FALSE, warning=FALSE}
+cowplot::plot_grid(p_3d_1, p_3d_2, ncol = 2, labels = c("a", "b"))
 ```

### 2017-02-28 manuscript.Rmd
Ampersand filter now working
papaja bug fixed

@@
 ---
-# title: "Modeling Cognitive Effects Vestibular Sensory Inference: From Sensory Inference To Imagined Self-Motion"
-
-# title: "Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion"
-
-title: "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
-# title: "Probabilistic Computations Underlying Simulated Self-Rotations"
-# title: Probabilistic Computations Underlying Vestibular Cognition
-shorttitle        : "Vestibular Contributions to SPT"
-
+title: 'A Computational Perspective On Spatial Perspective Transformations: From Sensory
+  Inference To Imagined Self-motion'
 author:
-  - name          : "Andrew W. Ellis"
-    # affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Fabrikstrasse 8, 3012 Bern, Switzerland"
-    email         : "andrew.ellis@psy.unibe.ch"
-  - name          : "Fred W. Mast"
-    # affiliation   : "1"
-
+- address: Fabrikstrasse 8, 3012 Bern, Switzerland
+  corresponding: yes
+  email: andrew.ellis@psy.unibe.ch
+  name: Andrew W. Ellis
+- name: Fred W. Mast
 affiliation:
-  # - id            : "1"
-  - institution   : "Department of Psychology, University of Bern, Switzerland"
-
-author_note: >
+- institution: Department of Psychology, University of Bern, Switzerland
+output:
+  '# papaja': default
+  keep_md: yes
+  papaja::apa6_word: null
+bibliography: bibliography.bib
+class: man
+fig_caption: yes
+figsintext: yes
+figurelist: yes
+footnotelist: no
+keywords: probabilistic graphical models, particle filter, embodied cognition, vestibular
+  cognition, spatial perspective taking, mental imagery, mental simulation
+lang: english
+lineno: yes
+linkcolor: blue
+number_sections: no
+author_note: |
   Andrew W. Ellis & Fred W. Mast, Department of Psychology.
-
   AE and FM planned and conducted the research; AE and FM performed the simulations and wrote the manuscript.
-
   Code for simulations is available at https://osf.io/u26mq.
-
-abstract: >
+shorttitle: Vestibular Contributions to SPT
+tablelist: yes
+abstract: |
   Mental spatial transformations are considered to be an essential cognitive ability.
@@
 Our model is able to explain various source of interference between simulations of self-rotation and sensory inference.
-
-keywords: probabilistic graphical models, particle filter, embodied cognition, vestibular cognition, spatial perspective taking, mental imagery, mental simulation
-
-wordcount: "X"
-
-bibliography: "bibliography.bib"
-
-figsintext        : yes
-figurelist        : yes
-tablelist         : yes
-footnotelist      : no
-number_sections   : no
-lineno            : yes
-linkcolor         : "blue"
-fig_caption       : true
-
-# citeproc          : yes
-# csl               : apa.csl
-
-lang              : "english"
-class             : "man"
-output:
-     papaja::apa6_word
-     # papaja::apa6_pdf
-        # pandoc_args:
-        # - --filter
-        #   pandoc-fignos
-
+wordcount: X
 ---
@@
 
+### Dynamics/kinematics
 For example, consider trying to estimate the velocity and position during a leftward head turn.
@@
 ```{r head-turn-data, echo=FALSE, fig.cap="Sensory measurements obtained during a short leftward head turn.
-The goal is to infer the latent angular velocity and position, message=FALSE, warning=FALSE, based on the noisy measurements.", message=FALSE, dpi=100}
+The goal is to infer the latent angular velocity and position, based on the noisy measurements.", message=FALSE, warning=FALSE, dpi=100}
 dt <- 0.1
@@
                        shape = 21, colour = "white",
-                       fill = "white", size = 6, stroke = 2) +
+                       fill = "white", size = 8, stroke = 2) +
             geom_point(data = df_tmp,
@@
                        colour = "black",
-                       fill = "white", size = 4, stroke = 2) +
+                       fill = "white", size = 6, stroke = 2) +
             guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
@@
 
-The dynamics of the system can be modelled by the evolution of the 3-d latent state variable, representing the position, velocity and acceleration, as a first-order Markov process, where the state at each time point $t$ depends only on the previous state at time $t-1$.
+### State-space notation
+The dynamics of the system can be modelled by the evolution of the 3-d latent state variable[^statespace], representing the position and its derivatives, velocity and acceleration, as a first-order Markov process, where the state at each time point $t$ depends only on the previous state at time $t-1$.
 This is referred to as the process model.
 Estimating the position $\theta$, velocity $\omega$ and acceleration $\alpha$ in discrete time requires writing the kinematic equations as a set of first order difference equations.
+
+[^statespace]: A state-space model reduces an n\textsuperscript{th} order difference equations (in discrete time) to n first-order difference equations.
 
@@
 (b) The same particle filter with increased process noise.", message=FALSE, warning=FALSE}
-cowplot::plot_grid(p_3d_1, p_3d_2, ncol = 2, labels = c("a", "b"))
+cowplot::plot_grid(p_3d_1, p_3d_2, ncol = 2, labels = c("a", "b"), label_size = 24)
 ```
@@
 
-# Modelling questions
-Experimental results to be explained: @Wertheim:2001cb report a cognitive suppression of otolith responses, based on participants' prior beliefs about possible motions.
+# TODO: Modelling questions
+
+- Connecting higher cognition to sensory inference: 
+    - imagery uses a similar generative model (copied from sensory model), then what is it doing? 
+    - __how does mental simulation work?
+When we propose a mental simulation, what does this actually mean?
+When we propose that simulation re-uses the circuit for sensory inference, which parts are actually required?
+Do we need the data?
+Surely not.
+NO, we need the generative model that is used for on-line inference, but we should not perform inference with it.
+ON the contrary, we should avoid inference at all costs, instead running a Monte Carlo simulation with the generative model, i.e.
+just evolving the evolution of the dynamical system according to the transition prior (sampling from the proposal distribution).__
+    - mental simulation allows us to answer "what if" (counterfactual) queries in a generative model.
+Why should this model be close to the sensory model?
+What if the senspry model changes?
+The twin model needs updating.
+    - If mental simulation interferes with sensory inference: how?
+    - __If we want to model higher level cognition as being connected to (grounded in) lower level sensory circuits, we can join both in a hierarchical graphical model.__
+
+- Explain what tasks sensory inference must perform.
+Tasks the are particular to the modality that senses movements of the body.
+It needs to detect non-linearities, i.e.
+non-stationary process.
+This means ist needs to detect the sudden onset of movements, or sudden direction changes.
+When these are passive and cannot be predicted , the best the system can hope for is to detect the changes as quickly as possible.
+When they are the result of active movement, they can be predicted.
+But they can also be anticipated due to knowledge from other sources, such as visual or verbal clues; if the knowledge is available, how does the system make use of it?
+Proposal distribution use for sequential importance sampling are notoriously difficult to design.
+The approach of sampling from the prior can be very inefficient, as when the process is not stationary, i.e.
+in this case when there are sudden changes in acceleration, the likelihood can change abruptly, and the weights would become neglible for all particles.
+In certain cases, the filter cannot recover from this, and would completely misrepresent the data.
+We propose that the ability to anticipate must result in the construction (flexible) of a proposal distribution, and that interference between mental simulation and inference involves the proposal distribution.
+
+> Control input for predictions is used to 'design' better proposal distributions for inference.
+
+- Experimental results to be explained: @Wertheim:2001cb report a cognitive suppression of otolith responses, based on participants' prior beliefs about possible motions.
+
+## TODO: Contrast model with Merfeld (2016)
+Merfeld's model claims: filtering is performed at a low level in order to extract velocity (maybe peak velocity?) and then this estimate is processed sub-optimally by low-pass filtering.
+The decision variable is computed at a higher level and separate from  the model that is invloved in inference.
+
+We propose that there is an alternaitve way of looking at this.
+What if there a second system that "interprets" the data when sensory inference using the 3-d state space representation alone is insufficent.
+This model "steps in" when the uncertainty is high, and it is this system that enables anticpation of non-stationarities, i.e.
+onset of sudden changes.
+This is derived from the system that allows us to predict consequences of own actions through the use of a control input obtained via a forward model from the cerebellum.
+This system would traditionally be viewed as depending on the motor commands from commands that are actually performed.
+Maybe this is more general, allowing the brain to include an exogenous variable in order to predict as a control input, i.e.
+this can also include "covert" actions.
+This flexible control of the exogenous input is actually necessary for planning, etc.
+Flexible control of input means: the ability to clamp nodes, i.e.
+to fix nodes at certain values.
+Without the ability to clamp higher-level node, the inference model can only be used inflexibly for inference.
+The whole idea is based on the fact that higher level knowledge and beliefs can influence our perception, and this is especially true for vestibular signals, which are highly ambiguous, and require prior knowledge for disambiguation and interpretation.
+Actually, we should implement something similar to [@Hamrick:8oFH5gP_]: a metacontroller, which chooses either the low-level module, or when this is unable to deal with the problem, a higher level module.
+This higher level module introduces additional information about the problem, e.g.
+information obtained by verbal instructions, or a bias.
+FOr example, the instructions: "choose right when uncertain" should lead participants to implement use this strategy.
+In other words, when you are certain, base you choice on your low level sensory inference, but when uncertain use other information, such as task demands.
+Surely we can design a better experiment than this, though?
+
+Implement the above ideas in a figure, showing the two approaches, and differences, especially for predictions about future data.
+
+## TODO: Simulations
+Show that when fit to the same data, a model implementing a control provides substantially better fit, as measured by log-likelihood, when compared to 3-d state space model (basic model) [think of better names for models].
+This is almost trivial; but show it like this: under the assumption that the agent knows the precise timing (onset) of motion, and the functional form (sinusiodal acceleration), all the agent needs to assume is: {A, direction, frequency}.
+Show that performance is substantially better if agent has priors for A, direction (maybe ignoring f, or vice versa: ignore A, and put priors on f, might be better).
+Then draw A (or f) from a distribution, and run inference using control model.
+Sampling A from prior distribution means: the agent has imperfect knowledge about its amplitude of motion.
+
+This improvement in inference is only better when the proposal distribution is well-specified; for a mis-specified model, the performance can be dramatically worse.
+
+We claim: interference effects arise when participant's wrongly use a control (predictive model or __anticipation model__) using motion paramters used for mental simulation during subsequent inference.
+This use of higher level priors has the same (similar) effect as anticipation.
 
@@
 # Various sources of interference
+
+Interference is interesting in its own right.
+It is important to understand how and why this should occur.
+Furthermore, we propose that gaining a better understanding of this will allow us to better understand the nature of "thinking" as mental simulation.
 
@@
 
+- Forward model for computing anticpation-related control input may reside in the cerebellum.
+THe cerebellum may be involved in imagined movemeent, or is it just for actions that are actually performed?
+
 ## Testable hypotheses

### 2017-03-01 manuscript.Rmd
restore title ampersand filter (papaja) now working

@@
 ---
-title: 'A Computational Perspective On Spatial Perspective Transformations: From Sensory
-  Inference To Imagined Self-motion'
+title: "A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion"
+shorttitle        : "Vestibular Contributions to SPT"
+
 author: 
-- address: Fabrikstrasse 8, 3012 Bern, Switzerland
-  corresponding: yes
-  email: andrew.ellis@psy.unibe.ch
-  name: Andrew W. Ellis
-- name: Fred W. Mast
+  - name          : "Andrew W. Ellis"
+    corresponding : yes    # Define only one corresponding author
+    address: Fabrikstrasse 8, 3012 Bern, Switzerland
+    email         : andrew.ellis@psy.unibe.ch
+  - name          : Fred W. Mast
+
 affiliation:
-- institution: Department of Psychology, University of Bern, Switzerland
-output:
-  '# papaja': default
-  keep_md: yes
-  papaja::apa6_word: null
-bibliography: bibliography.bib
-class: man
-fig_caption: yes
-figsintext: yes
-figurelist: yes
-footnotelist: no
-keywords: probabilistic graphical models, particle filter, embodied cognition, vestibular
-  cognition, spatial perspective taking, mental imagery, mental simulation
-lang: english
-lineno: yes
-linkcolor: blue
-number_sections: no
+    institution: Department of Psychology, University of Bern, Switzerland
+
 author_note: |
     Andrew W. Ellis & Fred W. Mast, Department of Psychology. 
-  AE and FM planned and conducted the research; AE and FM performed the simulations and wrote the manuscript.
-  Code for simulations is available at https://osf.io/u26mq.
-shorttitle: Vestibular Contributions to SPT
-tablelist: yes
+    AE and FM planned and conducted the research; AE and FM performed the simulations and wrote the manuscript. Code for simulations is available at https://osf.io/u26mq.
+
 abstract: |
@@
 Our model is able to explain various source of interference between simulations of self-rotation and sensory inference.
-wordcount: X
+
+keywords: probabilistic graphical models, particle filter, embodied cognition, vestibular cognition, spatial perspective taking, mental imagery, mental simulation
+wordcount         : "X"
+
+bibliography      : ["bibliography.bib"]
+
+figsintext        : yes
+figurelist        : yes
+tablelist         : yes
+footnotelist      : no
+lineno            : yes
+
+lang              : "english"
+class             : "man"
+output            : papaja::apa6_pdf
 ---

### 2017-03-24 draft-manuscript.Rmd
converted to word and sent to Fred
restarted working on manuscript

--- 2017-01-22 draft-manuscript.Rmd
@@
 class             : "man"
-output            : papaja::apa6_word
-# output            : papaja::apa6_pdf
+# output            : papaja::apa6_word
+output            : papaja::apa6_pdf
 ---
@@
 Two tasks thought to constitute vestibular cognition are : SPT and imagined self-motion.
-We show that both processes are thought to involve a simulated rotation about the body's vertical (yaw) axis.
-Describe both processes and show figure.
+= = = = =  Describe both processes and show figure.
 Both can be interpreted as mentally simulated rotations.

### 2017-03-24 manuscript.Rmd
restarted working on manuscript

@@
 
+Note the the vestibular senspry input may be the result of a superposition of multiple generative source; the afferent signals, measuring head velocity, may be caused by a multitude of processes (active head turn, unintentional body sway, perturbative external forcec (wind), etc.).
+The brain must treat these resulting head movements differently; active head turn may be goal-directed, with the aim of directing the gaze toward a target.
+In this case, no compensatory action os desired in order to counteract the movement.
+This would be counterproductive.
+However, in order to maintian postural stability, the brain must compensate for all movement that is not gol-directed, i.e.
+does not occur as a result of "planned" active motion.
+These components may be separated either at the level of the outcome variable (sensory signal), or at the level of the latent state variables.
+The latter amounts ot a latent decomposition.
+
+> Estimated velocity due to active/passive movement must be trated differently.
+Sensory signals due to active motion (re-afference) can be viewed as being cause by known explanatory variables, or as intervention effects.
+
+> discuss @Sanders_2015 (artificial velcoity vector)
+
 - Connecting higher cognition to sensory inference: 

### 2017-03-24 manuscript.tex
restarted working on manuscript

@@
             pdfauthor={},
-            pdftitle={Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion},
+            pdftitle={A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion},
             colorlinks=true,
@@
             urlcolor=blue,
-            linkcolor=blue,
+            linkcolor=black,
             pdfborder={0 0 0}}
@@
 % Essential manuscript parts
-  \title{Vestibular Contributions to Spatial Perspective Transformations: From
-Sensory Inference To Imagined Self-Motion}
+  \title{A Computational Perspective On Spatial Perspective Transformations: From
+Sensory Inference To Imagined Self-motion}
 
@@
     Andrew W. Ellis \& Fred W. Mast, Department of Psychology. AE and FM
-    planned and conducted the research; AE and FM analyzed the data and
-    wrote the manuscript. Code for simulations is available at
+    planned and conducted the research; AE and FM performed the simulations
+    and wrote the manuscript. Code for simulations is available at
     \url{https://osf.io/u26mq}.
@@
 
-  \abstract{Recent research has indicated that the vestibular network is involved in
-various cognitive tasks involving mental spatial transformations.
-However, the nature of this involvement remains unclear. Here, we
-provide a computational framework for vestibular cognition and discuss
-this in the context of perspective transformations, which require a
-simulated rotation of the self. We explore how the ability to model the
-effects of body rotation using vestibular input might enable the ability
-to imagine body rotations. We approach this problem by extending
-previous models of vestibular sensory processing to include the ability
-to model the sensory consequences of self-initiated movements, and
-discuss how this can be further extended to allow mentally simulated
-rotations of the self. We \ldots{} Predicting the sensory input is key
-to understanding imagined rotations. Mental imagery goes beyond
-prediction, rather, it involves the use of counterfactual queries in a
-probabilistic model.}
+  \abstract{Mental spatial transformations are considered to be an essential
+cognitive ability. Recent research has indicated that the vestibular
+network is involved in such mental spatial transformations. However, the
+nature of this involvement remains unclear. We propose a computational
+framework for the involvement of the vestibular system in cognition and
+discuss this in the context of perspective transformations, which
+require a simulated rotation of the self. We explore how the ability to
+anticipate the sensory consequences of self-initiated actions might
+underlie the ability to imagine body rotations. We approach this problem
+by extending previous models of vestibular sensory inference to include
+the ability to model the sensory consequences of self-initiated
+movements, and discuss how this can be further extended to allow
+mentally simulated rotations of the self. Our model is able to explain
+various source of interference between simulations of self-rotation and
+sensory inference.}
   \keywords{probabilistic graphical models, particle filter, embodied cognition,
-vestibular cognition, spatial perspective taking, mental imagery \\
+vestibular cognition, spatial perspective taking, mental imagery, mental
+simulation \\
 
@@
 \settowidth{\cfttabnumwidth}{Table 10.\qquad}
-
-  \usepackage{caption}
 
@@
 
-% pandoc-xnos: macro to create a caption without a prefix
-\makeatletter
-\long\def\@makenoprefixcaption#1#2{
-  \vskip\abovecaptionskip
-  \sbox\@tempboxa{#2}
-  \ifdim \wd\@tempboxa >\hsize
-    #2\par
-  \else
-    \global \@minipagefalse
-    \hb@xt@\hsize{\hfil\box\@tempboxa\hfil}
-  \fi
-  \vskip\belowcaptionskip}
-\makeatother
-
-% pandoc-fignos: save original macros
-\makeatletter
-\let\@oldmakecaption=\@makecaption
-\let\oldthefigure=\thefigure
-\let\oldtheHfigure=\theHfigure
-\makeatother
-
-% pandoc-fignos: environment disables figure caption prefixes
-\makeatletter
-\newcounter{figno}
-\newenvironment{no-prefix-figure-caption}{
-  \let\@makecaption=\@makenoprefixcaption
-  \renewcommand\thefigure{x.\thefigno}
-  \renewcommand\theHfigure{x.\thefigno}
-  \stepcounter{figno}
-}{
-  \let\thefigure=\oldthefigure
-  \let\theHfigure=\oldtheHfigure
-  \let\@makecaption=\@oldmakecaption
-  \addtocounter{figure}{-1}
-}
-\makeatother
-
-% pandoc-xnos: cleveref fakery
-\newcommand{\plusnamesingular}{}
-\newcommand{\starnamesingular}{}
-\newcommand{\xrefname}[1]{\protect\renewcommand{\plusnamesingular}{#1}}
-\newcommand{\Xrefname}[1]{\protect\renewcommand{\starnamesingular}{#1}}
-\providecommand{\cref}{\plusnamesingular~\ref}
-\providecommand{\Cref}{\starnamesingular~\ref}
-\providecommand{\crefformat}[2]{}
-\providecommand{\Crefformat}[2]{}
-
-% pandoc-xnos: cleveref formatting
-\crefformat{figure}{Figure~#2#1#3}
-\Crefformat{figure}{Figure~#2#1#3}
-
 \section{Introduction}\label{introduction}
 
-Adopting the spatial perspective of another person is generally
-considered an essential cognitive ability, and plays a vital role in our
-ability to determine what another person can see and in order to
-coordinate joint actions (Creem-Regehr, Gagnon, Geuss, \& Stefanucci,
-2013, Pezzulo, Iodice, Ferraina, \& Kessler (2013)). Many studies have
-also pointed out that perspective taking is of fundamental importance
-for language comprehension and communication (Beveridge \& Pickering,
-2013), and social cognition in general (Deroualle \& Lopez, 2014).
-
-A number of recent studies have demonstrated that the vestibular system,
-which deals with sensory signal relating to movements of the head, is
-involved in such spatial transformations (Lenggenhager, Lopez, \&
-Blanke, 2008, Deroualle, Borel, Devèze, \& Lopez (2015), Falconer \&
-Mast (2012), Gardner, Stent, Mohr, \& Golding (2016)). This is
-demonstrated in the form of interference effects between mental
-transformations and concurrent sensory stimulation. However, to date,
-there has little attempt to understand the involvement of the vestibular
-system in higher-level cognitive abilities from a computational
-perspective. In this paper, we focus specifically on the role of the
-vestibular system for spatial perspective transformations, and we draw
-the link between probabilistic computations performed by the vestibular
-system in the service of real-time interaction with the world, and the
-brain's ability to run mental simulations for the purpose of adopting
-the perspective of another person.
-
-Consider the use of deictic pronouns, such as \enquote{this} or
-\enquote{that} and \enquote{left} or \enquote{right}. These terms only
-assume a their meaning when placed within a certain frame of reference
-(FOR). For example, in a dialogue between two speakers, the statement
-\enquote{I would like that one}, or \enquote{I would like the one on the
-right} can only be understood under the assumption that the deictic
-terms \enquote{that} and \enquote{right} are used in an egocentric FOR.
-In addition, both interlocutors must be able to access the other's FOR
-with respect to an external FOR. In other words: In order to give
-instructions, the speaker must be able to mentally assume the
-addressee's position in space. A further example that requires a similar
+Adopting the spatial perspective of another person is an essential
+cognitive ability, and plays a vital role in our ability to determine
+possible actions of another person, and in order to coordinate joint
+actions (Creem-Regehr, Gagnon, Geuss, \& Stefanucci, 2013, Pezzulo,
+Iodice, Ferraina, and Kessler (2013)). Perspective taking is also
+considered to be of fundamental importance for language comprehension
+and communication (Beveridge \& Pickering, 2013), and social cognition
+in general (Deroualle \& Lopez, 2014).\\
+ A number of recent studies have demonstrated that the vestibular
+system, which processes sensory signal relating to angular and linear
+movements of the head, is involved in such spatial transformations
+(Lenggenhager, Lopez, \& Blanke, 2008, Deroualle, Borel, Devèze, and
+Lopez (2015), Falconer and Mast (2012), Gardner, Stent, Mohr, and
+Golding (2016)). This is demonstrated in the form of interference
+effects between mental transformations and concurrent sensory
+stimulation. To date, however, there has little attempt to understand
+the involvement of the vestibular system in higher-level cognitive
+abilities in terms of the computational mechanisms. Consider the use of
+deictic pronouns, such as \enquote{this} or \enquote{that} and
+\enquote{left} or \enquote{right}. These terms only assume a meaning
+relative to a certain frame of reference (FOR). For example, in a
+dialogue between two speakers, the statement \enquote{I would like that
+one}, or \enquote{I would like the one on the right} can only be
+understood if both interlocutors are able to access the other's FOR with
+respect to an external FOR. A further example that requires a similar
 spatial transformation is when two people want to perform a coordinated
@@
 of the other's potential actions in order to adjust one's own actions,
-and this again requires knowledge of the other's position with an
+and this again requires knowledge of the other's position within an
 external FOR. A judgement about what kind of actions the other person
 can perform requires an embodied simulation using a representation of
-one's own body schema (Pezzulo et al., 2013). Giving an instruction to
-another person thus requires running an embodied simulation in order to
-infer what kind of motor commands the other person should perform
+one's own body schema (Pezzulo et al., 2013) and giving an instruction
+to another person thus requires running an embodied simulation in order
+to infer what kind of motor commands the other person should perform
 \textbf{{[}Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod,
 2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo,
-2011a,b{]}}. This embodied simulation is performed using one's own body
-schema. Communicating this information to the other actor requires
-knowing the position of the addressee's FOR relative to the external
-reference frame, and the difference between the speaker and the
-addressee's egocentric reference frames. Perspective taking thus can be
-conceived of as a aligning a representation of one's body with that of
-another person in order to access spatial information from a viewpoint
-other than one's own egocentric viewpoint. In order to achieve this
-alignment, one must perform translations and rotations of one's own
-egocentric reference frame. Processing translations and rotations of the
-head and body is the core domain of the vestibular system, and thus is
-of particular interest for research into mental spatial transformations.
-Spatial transformations in 6 dimensions (translation along 3 axes, and
-rotations around 3 axes) are highly complex, and non-linear, and we
-propose that the brain adopts the strategy of simulating motion of the
-body through space in order achieve a desired target position. This
-effectively amounts to a running a simulation of dynamical system,
-consisting of simulated limb movements and the resulting sensory
-consequences. \autoref{fig:spt-schematic} (a) shows a simple scenario in
-which a speaker wants to give an addressee, standing opposite, verbal
-instructions about an action to perform on an external object. The
-speaker must determine the most suitable action, given the addressee's
-position relative to the external object, and in order to achieve this
-goal, the speaker must align himself with the addressee, so that he can
-use his own body-centred reference frame to quickly simulate the best
-action. In this example, alignment with the addressee's reference frame
-consists of a forwards translation and a rotation around the vertical
-axis. This can be described as an imagined movement, in which the goal
-is achieve a specific position.
+2011a,b{]}}. But this again requires information about the addressee's
+FOR relative to the external reference frame, and the difference between
+the speaker and the addressee's egocentric reference frames. Perspective
+taking can be conceived of as a aligning a representation of one's body
+with that of another person in order to access spatial information from
+a viewpoint other than one's own egocentric viewpoint. In order to
+achieve this alignment, one must perform translations and rotations of
+one's own egocentric reference frame. Processing translations and
+rotations of the head and body is the core domain of the vestibular
+system, and thus is of particular interest for research into mental
+spatial transformations. Spatial transformations in 6 dimensions
+(translation along 3 axes, and rotations around 3 axes) are highly
+complex, and non-linear, and it has been proposed that the brain adopts
+the strategy of simulating motion of the body through space in order
+achieve a desired target position {[}CITATION{]}. This effectively
+amounts to a running a simulation of dynamical system, consisting of
+simulated limb movements and the resulting sensory consequences. Figure
+\ref{fig:spt-schematic} (a) shows a simple scenario in which a speaker
+wants to tell an addressee, standing opposite, to pick something up from
+the floor. The speaker must determine the most suitable action by
+embodied simulation, given the addressee's position relative to the
+external object, and in order to achieve this goal, the speaker must
+align himself with the addressee, so that he can use his own
+body-centred FOR. In this example, alignment with the addressee's
+reference frame consists of a forwards translation and a rotation around
+the vertical axis. This can be described as an imagined movement, in
+which the goal is achieve a specific position.
 
@@
 \centering
-\includegraphics[width=0.50000\textwidth]{diagrams/fig-1-common-process.pdf}
-\caption{Perspective taking and imagined self-motion involve spatial
-transformations of a representation of the self. (a) If a speaker's goal
-is to take the perspective of an addressee, he must first solve the
-inverse problem of determining the appropriate action, and then run a
-forwards simulation in order to perform a translation and rotation of
-his own egocentric reference frame. These operations can be viewed as
-imagined movements.(b) Focussing on imagined rotation around the
-vertical axis allows us to simplify the problem, and highlights the fact
-that perspective taking and mental imagery involve common
-computations.\label{fig:spt-schematic}}
+\includegraphics{diagrams/fig-1-common-process.pdf}
+\caption{\label{fig:spt-schematic}Perspective taking and imagined
+self-motion involve spatial transformations of a representation of the
+self. (a) If a speaker's goal is to take the perspective of an
+addressee, he must first solve the inverse problem of determining the
+appropriate action, and then run a forwards simulation in order to
+perform a translation and rotation of his own egocentric reference
+frame. These operations can be viewed as imagined movements.(b)
+Focussing on imagined rotation around the vertical axis allows us to
+simplify the problem, and highlights the fact that perspective taking
+and mental imagery involve common computations.}
 \end{figure}
@@
 In order to do this, the speaker must first determine what type of
-movement to perform, and the direction, velocity and duration that lead
-to the desired outcome. This means that in order to know what to
-simulate, the speaker must first work backwards from the desired
-outcome, and infer the actions that are most likely to lead to that
-outcome. This is known in the field of motor control as the inverse
-problem (Wolpert \& Kawato, 1998). In order to achieve the desired
-position, the motor commands can then be simulated, and the expected
-sensory consequences can be estimated; this is known as the forward
-problem. \xrefname{Figure}\cref{fig:spt-schematic} (b) shows that if the
-goal is to simply simulate the perceptual experience of rotation around
-the body's vertical (yaw) axis, rather than project a representation of
-one's body onto a target, the problem is simplified; the actor can
-perform forward inference without the need to determine the optimal
-motor action. As demonstrated by Penny, Zeidman, \& Burgess (2013),
-these computations can be performed in a common probabilistic model.
-Interestingly, this has been discussed in the context of reinforcement
-learning, and planning in humans has been described as reverse inference
-(Botvinick \& Toussaint, 2012). Further, (Pickering \& Clark, 2014)
-discuss the relationship between control-theoretic notions of
-inverse/forward models, known collectively as internal models, and and
-more general concepts of a generative model in a hierarchical Bayesian
-framework.
-
-Performing a simulation of body movements and estimating the sensory
-consquences amounts to simulating the behaviour of a dynamical system;
-this requires a generative model of the system. The aim of this paper is
-to specifically investigate the generative model that may be used for
-mental spatial transformations. The ability to run embodied simulations
-is often claimed to be grounded in the relevant sensorimotor circuits,
-and we claim that the vestibular system is ideally suited to investigate
-this connection, because the generative models used by the vestibular
-system for dynamic sensory inference have been extensively investigated
-using approaches from optimal control theory, and more recently, dynamic
-Baysian inference. The vestibular system faces a number of challenges in
-interpreting the sensory measurements provided by the semi-circular
-canals (SCC) and otolith organs. On the one hand, these signals are very
-noisy, and ambiguous (MacNeilage, Ganesan, \& Angelaki, 2008); the
-otoliths respond both to translations and tilt of the head relative to
-gravity, and in order to disambiguate between translation and tilt, the
-brain must perform sophisticated computations at the level of the brain
-stem. It has become clear that the nature of these computations is
+movement to perform. This means that in order to know what to simulate,
+the speaker must first work backwards from the desired outcome, and
+infer the actions that are most likely to lead to that outcome (Hamrick
+\& Griffiths, 2014). In the field of motor control, this is known as the
+inverse problem (Wolpert \& Kawato, 1998). In order to achieve the
+desired position, the motor commands can then be simulated, and the
+expected sensory consequences can be estimated; this is known as the
+forward problem. Figure \ref{fig:spt-schematic} (b) shows that if the
+goal is to deliberatively simulate the perceptual experience of rotation
+around the body's vertical (yaw) axis, rather than project a
+representation of one's body onto a target, the problem is simplified;
+the actor can perform forward inference without the need to determine
+the optimal motor action. As demonstrated by Penny, Zeidman, and Burgess
+(2013), these computations can be performed in a common probabilistic
+model. Interestingly, this has been discussed in the context of
+reinforcement learning, and planning in humans has been described as
+reverse inference (Botvinick \& Toussaint, 2012). Further, (Pickering \&
+Clark, 2014) discuss the relationship between control-theoretic notions
+of inverse/forward models, known collectively as internal models, and
+and more general concepts of a generative model in a hierarchical
+Bayesian framework. Simulating the behaviour of a dynamical system using
+a forward model requires a generative model of the system.
+
+The aim of this paper is to specifically investigate the generative
+model that may be used for mental spatial transformations. The ability
+to run embodied simulations is often claimed to be grounded in the
+relevant sensorimotor circuits, and we claim that the vestibular system
+is ideally suited to investigate this connection, because the generative
+models used by the vestibular system for dynamic sensory inference have
+been extensively investigated using approaches from optimal control
+theory, and more recently, dynamic Bayesian inference (Oman, CITATIONS,
+MacNeilage, Ganesan, \& Angelaki, 2008).
+
+The vestibular system faces a number of challenges in interpreting the
+sensory measurements provided by the semi-circular canals (SCC) and
+otolith organs. On the one hand, these signals are very noisy, and
+ambiguous; the otoliths respond both to translations and tilt of the
+head relative to gravity, and in order to disambiguate between
+translation and tilt, the brain must perform sophisticated computations.
+The emerging picture is that the nature of these computations is
 probabilistic, and that the brain makes use of strong prior beliefs in
-order to interpret these sensory signals. A further challenge faced by
-the vestibular system is that it needs to distinguish between sensory
-signals that are the result of active motion, and those resulting from
-external forces (Cullen, Brooks, \& Sadeghi, 2009, Cullen (2014)). The
-ability to distinguishing re-afferent from ex-afferent sensory signals
-is an essential capability of all organisms that move actively (Crapse
-\& Sommer, 2008).
+order to interpret these sensory signals {[}Mittelstaedt, etc{]}. A
+further challenge faced by the vestibular system is that it needs to
+distinguish between sensory signals that are the result of active
+motion, and those resulting from external forces (Cullen, Brooks, \&
+Sadeghi, 2009, Cullen (2014)). The ability to distinguishing re-afferent
+from ex-afferent sensory signals is an essential capability of all
+organisms that move actively (Crapse \& Sommer, 2008).
 
@@
 predictive processing has repeatedly been alluded to (Moulton \&
-Kosslyn, 2009, Clark (2012), Gambi \& Pickering (2015)), and Grush
+Kosslyn, 2009, Clark (2012), Gambi and Pickering (2015)), and Grush
 (2004) claims that mental imagery is performed by emulator circuits
@@
 different time scales. This was recently discussed by (Tian \& Poeppel,
-2012) in the context of imagined speech.
+2012) in the context of imagined speech. The putative link between
+predictive processing and higher cognitive processing has a firm basis
+in predictive processing accounts of sensory inference {[}Friston{]}.
+The predictive processing framework has been applied to almost all
+sensory modalities {[}CITATIONS: pain, proprioception, auditory ,
+visual, etc.{]}, and the picture that has emerged in recent years is
+that the brain employs sophisticated inference algorithms in order to
+\enquote{interpret} the stream of incoming sensory data. The brain seems
+to be actively engaged in formulating hypotheses, and testing these on
+sensory data; this has been extended to the concept of active inference
+{[}Friston{]}, in which an agent attempts to either change it's internal
+model of world, or to intervene on the world in order to fulfill sensory
+predictions.
+
+We claim that the vestibular modality is particularly well-suited to
+investigating predictive processing at multiple hierarchical scales,
+from the lowest levels of sensory inference to higher levels, where
+goals, intentions and beliefs of the organism become relevant. A case in
+point is the study by Wertheim, Mesland, and Bles (2001), who provided
+evidence that participants' perceptions of tilt induced by vestibular
+stimulation are shaped to a large extent by the beliefs about possible
+motions they formed prior to taking part in the experiment.
+
+Interference effects between higher-level cognitive processes, such as
+SPT and mental imagery, and sensory inference, are thought to occur as a
+result of shared circuitry (Deroualle \& Lopez, 2014, Nigmatullina et
+al. (2015)). However, the precise nature of such interactions, and what
+exactly might be shared, is left unspecified. At first glance, it appear
+unlikely that such interference should be possible, as low levels of
+vestibular processing appear to be cognitively impenetrable
+{[}CITATION{]}. In particular, Nigmatullina et al. (2015) found that the
+onset of the vestibulo-ocular reflex (VOR) was earlier when the
+direction of an imagined self-rotation was congruent with the direction
+of a subsequently experienced rotation. The VOR, however, must be
+susceptible to the influence of intentions and goals; for example, if a
+person intends to fixate a target during a movement, the VOR should be
+cancelled. Thus, seemingly reflexive, and impenetrable, motor actions
+require a large amount of flexibility based on top-down information.
+Even accepting the fact that low level sensorimotor processing may be
+more cognitively penetrable than previously thought, the fact that
+imagining a rotation of the self should have an effect on a person's
+ability to detect rotations, or to discriminate between directions of
+rotation, still requires an explanation.
+
+\begin{quote}
+NOTE: we need to include all studies demonstrating interference,
+especially in the direction cognition -\textgreater{} inference
+(Nigmatullina effect)
+\end{quote}
+
+Ellis and Mast (2017) recently claimed that the types of interference
+effects might arise as a result of cognitive processing exerting an
+influence on circuits performing sensory inference via shared
+higher-level priors. In particular, the interference described by
+Nigmatullina et al. (2015) looks like an effect of a bias in the sensory
+discrimination task. Although vestibular perceptual thresholds have been
+extensively investigate for yaw rotations (Grabherr, Nicoucar, Mast, \&
+Merfeld, 2008; Merfeld, 2011), the methods used for estimating
+thresholds require unbiased responding by the participants. There have
+no investigations of biasing effects on yaw rotations.
+
+In this paper, we pursue the claim that interference should be
+investigated as a bias, and we attempt to explain the grounding of
+higher level cognitive abilities in the computations performed by
+low-level sensory circuits in the service of dynamic real-time
+interactions with the world.
+
+Kwisthout, Bekkering, and Rooij (2016) make the point that predictive
+processing accounts are situated at low levels in the processing
+hierarchy, usually at the level of sensory inference or at the level
+describing kinematics.
+
+In this paper, we address this issue by starting from a model developed
+for low-level sensory inference, incorporating basic abilities to
+predict sensory consequences of movement based on knowledge about the
+kinematics, and extending this to account for interactions between
+thought processes, operationalized as mental simulations of sensorimotor
+experience, and sensory processing.
+
+Battaglia, Hamrick, and Tenenbaum (2013) addressed the idea that the
+higher level cognitive ability to perform mental rotation are grounded
+in the circuits performing sensory inference. We claim that if
+vestibular cognition is grounded in sensory inference, and that this is
+to be experimentally studied by analyzing interactions between thought
+and sensory inference, then it is vital that this be carried within the
+constraints of a theoretical (computational) framework. A suitable
+formalism for this is the framework of probabilistic graphical models
+(Pearl, 1988, Koller and Friedman (2009)), and within this framework,
+interactions between thought and perception can be modelled as effects
+of higher level priors and hyperpriors on inference at a computational
+level. Investigating inference at the level of algorithmic
+implementations might provide further insight.
+
+Probabilistic graphical models are able to represent hierarchically
+structured knowledge (Kwisthout et al., 2016, Griffiths, Vul, and
+Sanborn (2012)), which is required for Bayesian models of cognition
+(Tenenbaum, Kemp, Griffiths, \& Goodman, 2011, Lucas and Griffiths
+(2010)). This approach has been used to model concept learning (Kemp,
+Perfors, \& Tenenbaum, 2007), inductive biases in reasoning, in
+cognitive models of development (A. Perfors, Tenenbaum, Griffiths, \&
+Xu, 2011), language (Goodman \& Frank, 2016) and social cognition
+(Baker, Saxe, \& Tenenbaum, 2009).
+
+In this paper, we focus specifically on the role of the vestibular
+system for spatial perspective transformations, and we draw the link
+between probabilistic computations performed by the vestibular system,
+and the brain's ability to run mental simulations for the purpose of
+adopting the perspective of another person.
 
@@
 real-time, dynamic sensory processing of vestibular data, based on
-particle filters (Laurens \& Droulez, 2007, Karmali \& Merfeld (2012)),
-and we reframe these as probabilistic graphical models. This allows us
+particle filters (Laurens \& Droulez, 2007, Karmali and Merfeld (2012)),
+and we re-frame these as probabilistic graphical models. This allows us
 to investigate the generative model separately from the algorithm used
 for inference, and we then extend the basic model for sensory inference
-in order to provide the ability to model the consequences of
-self-initiated, active movements. Our aims are to:
+in order to focus specifically on the ability to model the consequences
+of self-initiated, active movements. This is an implementation of the
+idea that information about motor commands are projected to sensory
+areas by means of an efference copy, based on which a forward model then
+computes the sensory consequences. Our aims are to:
 
@@
 
+Following Karmali and Merfeld (2012), we focus on one specific type of
+head movement, a rotation of the head about the yaw axis, and discuss
+this in the context of the probabilistic computations that are performed
+when the goal is to:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\tightlist
+\item
+  infer motion caused by external events
+\item
+  infer the consequences of one's own actions.
+\end{enumerate}
+
 \section{Dynamic Bayesian models for sensory
@@
 Whereas early models of vestibular sensory processing were inspired by
-control theory (Borah, Young, \& Curry, 1988, Merfeld, Zupan, \& Peterka
-(1999)), several recent studies have described vestibular sensory
-processing as dynamic Bayesian inference (Laurens \& Droulez, 2007,
-MacNeilage et al. (2008), Karmali \& Merfeld (2012)). Selva \& Oman
-(2012) provide a recent review of the relationship between optimal
-observer models and models which employ Kalman filtering, and show that
-these are largely equivalent. Both Karmali \& Merfeld (2012) and Laurens
-\& Droulez (2007) described vestibular sensory inference in terms of
-particle filtering, which is a particular inference algorithm that can
-be used to perform inference in state-space models. (Doucet, Godsill, \&
-Andrieu, 2000, Doucet \& Johansen (2009)). Using their model, Laurens \&
-Droulez (2007) were able to simulate perceptual responses to
-centrifugation and off-vertical axis rotations.
-
-A further well-known phenomenon of vestibular perception is velocity
-storage, which is defined as the slower decay of perceptual and
-oculomotor responses, relative to the SCC afferent responses (Raphan,
-Matsuo, \& Cohen, 1979), \textbf{{[}Robinson 1977{]}}. This is generally
-seen as evidence for the existence of an internal model in the
-processing of vestibular signals. Karmali \& Merfeld (2012) focussed on
-one-dimensional angular velocity during yaw rotations; their model was
-able to explain the phenomenon of velocity storage as a consequence of
-particle filtering, with the time constant of velocity storage depending
-on the variability of the sensory afferent signals.
+control theory (Borah, Young, \& Curry, 1988, Merfeld, Zupan, and
+Peterka (1999)), several recent studies have described vestibular
+sensory processing as dynamic Bayesian inference (Laurens \& Droulez,
+2007, MacNeilage et al. (2008), Karmali and Merfeld (2012)). Selva and
+Oman (2012) provide a recent review of the relationship between optimal
+observer models and models that employ Kalman filtering, and show that
+these are largely equivalent. Both Karmali and Merfeld (2012) and
+Laurens and Droulez (2007) described vestibular sensory inference in
+terms of particle filtering, which is a particular algorithm that can be
+used to perform inference in state-space models. (Doucet, Godsill, \&
+Andrieu, 2000, Doucet and Johansen (2009)). Using their model, Laurens
+and Droulez (2007) were able to simulate perceptual responses to
+centrifugation and off-vertical axis rotations. A well-known phenomenon
+of vestibular perception is velocity storage, which is defined as the
+slower decay of perceptual and oculomotor responses, relative to the SCC
+afferent responses (Raphan, Matsuo, \& Cohen, 1979), \textbf{{[}Robinson
+1977{]}}. This is generally seen as evidence for the existence of an
+internal model in the processing of vestibular signals. Karmali and
+Merfeld (2012) focused on one-dimensional angular velocity during yaw
+rotations; their model was able to explain the phenomenon of velocity
+storage as a consequence of particle filtering, with the time constant
+of velocity storage depending on the noise of the sensory afferent
+signals.
 
@@
 state-space models, and particle filtering, which is a particular
-algorithm that can perform inference in state-space models. While Kalman
-filters are restricted to linear Gaussian State-space models, particle
-filtering can be applied to arbitrary nonlinear models. State-space
-models can be seen as special cases of more general Bayesian
-hierarchical models (Bishop, 2006; Murphy, 2012). Following the same
-approach as Penny et al. (2013), we first describe the generative model
-used for sensory inference as a dynamic probabilistic graphical model.
-
-A state-space is special case of a dynamic latent variable model which
-is suitable for modelling dynamical system in terms of latent,
-unobservable processes and noisy observations. As an example, consider
-the case where we want to model the velocity and position during a
-leftward head turn. While the SCC measures angular acceleration, the
-afferent signals are thought to be proportional to the angular velocity
-of the head with a frequency band of approximately 0.1 -- 10 Hz, which
-constitutes the range of natural head movements (Carriot, Jamali,
-Chacron, \& Cullen, 2014; Grabherr, Nicoucar, Mast, \& Merfeld, 2008).
-The measurements are observations of the true, unobservable angular
-velocity; however, they are contaminted by Gaussian noise, which is
-determined by characteristics of the sensory organs.
-\autoref{fig:head-turn-data} shows a sequence of noisy measurements
-during a two-second turn of the head. The acceleration generating the
+algorithm that can be used to perform inference in state-space models.
+State-space models can be seen as special cases of more general dynamic
+probabilistic graphical models (Bishop, 2006; Murphy, 2012), and are
+suitable for modelling the time-varying behaviour of a system in terms
+of a latent, unobservable process and noisy measurements.
+
+\subsubsection{Dynamics/kinematics}\label{dynamicskinematics}
+
+For example, consider trying to estimate the velocity and position
+during a leftward head turn. In this case, the latent process provides a
+model of the kinematics of the system, and the measurement model
+describes how the afferent SCC signals are generated, dependent on the
+kinematics.
+
+The SCC signals are considered to be proportional to the angular
+velocity of the head within a frequency band of approximately 0.1 -- 10
+Hz, which constitutes the range of natural head movements (Carriot,
+Jamali, Chacron, \& Cullen, 2014; Grabherr et al., 2008). These
+measurements, however, are contaminated by Gaussian noise, which is
+determined by characteristics of the sensory organs. Figure
+\ref{fig:head-turn-data} shows a sequence of noisy measurements during a
+two-second leftward turn of the head. The acceleration generating the
 movement is sinusoidal, given for each time step \(t\) by
@@
 \(T = 1/f\) is the duration of the movement, and \(f\) is the frequency.
-This acceleration results in a peak velocity \(\omega_{peak} = AT/\pi\)
-and a final angular position \(\theta_{final} = AT^2/2\pi\).
-
-\begin{no-prefix-figure-caption}
+This acceleration results in a half-cycle cosine velocity profile, with
+a peak velocity \(\omega_{peak} = AT/\pi\) and a final angular position
+\(\theta_{final} = AT^2/2\pi\). Particle filtering provides a means of
+filtering out the noise inherent in the measurements, and allows
+estimation of evolution of the unobservable, latent state variables.
 
@@
 \centering
-\includegraphics{../generated-figures/head-turn-data-1.pdf}
+\includegraphics{generated-figures/head-turn-data-1.pdf}
 \caption{\label{fig:head-turn-data}Sensory measurements obtained during a
-short head turn. The goal is to infer the latent angular velocity and
-position, based on the noisy measurements. \label{fig:head-turn}}
+short leftward head turn. The goal is to infer the latent angular
+velocity and position, based on the noisy measurements.}
 \end{figure}
 
-\end{no-prefix-figure-caption}
-
-The key idea is that we can write model the dynamics of a system by the
-evolution of the latent state variables in discrete time steps as a
-Markov process, where the state at each time point \(t\) depends only on
-the previous state at time \(t-1\). This is referred to as the process
-model. Estimating the velocity \(\omega\) and position \(\theta\) in
-discrete time therefore requires writing the kinematic equations as a
-set of first order difference equations.
-
-\[\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t\]
-\[\omega_{t} = \omega_{t-1}\]
-
-Note that we have omitted the acceleration from the process model, for
-the sake of simplicity; this amounts to assuming a constant velocity
-model. These kinematic equations form a linear model, and could
-optionally be written in matrix notation. For further simplicity, we
-retain the current notation. Next, we consider the measurements at each
-time step \(t\) as depending only on the current state. This simply
-implements the idea that the sensory signal depends only on the current
-head velocity. The above equations represent the expected values of the
-state and observation variables. In order to complete the generative
-model, we require a description of the stochastic nature dependancies.
-We model both process and measurement noise a being normally
-distributed, leading to the following linear Gaussian state-space model:
-
-\[x_t = f(x_{t-1}, \epsilon)\] \[y_{t} = g(x_t, \delta)\]
-\[\epsilon \sim N(0, \sigma^2_\epsilon) \]
-\[\delta \sim N(0, \sigma^2_\delta) \]
-
-where the state variable \(x_t\) consists of both the velocity
-\(\omega_t\) and position \(\theta_t\), and \(y_{t}\) denotes the
-sensory measurement.\footnote{Note that we although both Laurens \&
-  Droulez (2007), Karmali \& Merfeld (2012) consider a realistic sensor
-  model, for simplicity, we implement the sensor as being a noisy
-  realisation of the velocity. Our model is not intended to be a
-  realistic model of the SCC; rather, we focus on the probabilistic
-  computations.} \(\epsilon\) and \(\delta\) are zero-mean Gaussian
-noise terms with variances \(\sigma^2_\epsilon\) and
+\subsubsection{State-space notation}\label{state-space-notation}
+
+The dynamics of the system can be modelled by the evolution of the 3-d
+latent state variable\footnote{A state-space model reduces an
+  n\textsuperscript{th} order difference equations (in discrete time) to
+  n first-order difference equations.}, representing the position and
+its derivatives, velocity and acceleration, as a first-order Markov
+process, where the state at each time point \(t\) depends only on the
+previous state at time \(t-1\). This is referred to as the process
+model. Estimating the position \(\theta\), velocity \(\omega\) and
+acceleration \(\alpha\) in discrete time requires writing the kinematic
+equations as a set of first order difference equations.
+
+\[ \theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2} \Delta t^2 \]
+\[ \omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t \]
+\[ \alpha_t = \alpha_{t-1} \]
+
+This model assumes a constant acceleration. These kinematic equations
+form a linear model, and could optionally be written in matrix notation.
+For simplicity, we retain the current notation. Next, we consider the
+measurements at each time step \(t\) as depending only on the current
+state. This simply implements the idea that the sensory signal depends
+only on the current head velocity. The above equations represent the
+expected values of the state and observation variables. In order to
+complete the generative model, we require a description of the
+stochastic nature of the dependencies. We model both process and
+measurement noise as being normally distributed, leading to the
+following linear Gaussian state-space model:
+
+\[ x_t = f(x_{t-1}, \epsilon) \] \[ y_{t} = g(x_t, \delta) \]
+\[ \epsilon \sim N(0, \sigma^2_\epsilon) \]
+\[ \delta \sim N(0, \sigma^2_\delta) \]
+
+where the state variable \(x_t\) consists of the position \(\theta_t\),
+velocity \(\omega_t\) and acceleration \(\alpha_t\), and \(y_{t}\)
+denotes the sensory measurement. \(\epsilon\) and \(\delta\) are
+zero-mean Gaussian noise terms with variances \(\sigma^2_\epsilon\) and
 \(\sigma^2_\delta\). These refer to the process and measurement noise,
-respectively. The process noise
-
-\autoref{fig:graphical-model-inference} shows a representation of
+respectively. The process noise represents the uncertainty in the
+dependence of the current state on the previous state; under the
+constant velocity assumption, the process noise must be large enough if
+the model is to allow for change in velocity due to unmodelled
+acceleration. Note that the measurement noise is a model of the noise
+characteristics of the sensor, and not a physical property of the sensor
+itself. The function \(f\) and \(g\) represent the process transition
+function and the measurement function, respectively.
+
+Figure \ref{fig:graphical-model-inference} shows a representation of a
 state-space model as a graphical model, unrolled over multiple time
-steps. Starting at time 0, the state variables evolve according to the
-process model, with the measurements at each time step being generated
-according to the measurement model. The 2-dimensiontal latent state
-variable is unshaded, to indicate that it cannot be observed, wheres the
-observed sensory measurements are shaded. Whereas previous models were
-discussed in terms of the particular filtering algorithm used, the
-graphical notation highlights the fact that a state-space model can be
-viewed as a dynamic latent factor model (Blei, 2014). This opens up the
-possibility of extending the graphical model to include higher-level
-variables. After discussing particle filtering, an inference algorithm
-often used to infer the values of the latent state variables, we will
-introduce our extension to the model that will allow a prediction of the
-consequences of active movements.
+steps. The state evolves according to the process model, with the
+measurements at each time step being generated according to the
+measurement model. The 3-dimensional latent state variable is unshaded,
+to indicate that it cannot be observed, wheres the observed sensory
+measurements are shaded. Given a generative model, we can apply various
+sampling algorithms, either in order to run a Monte Carlo simulation by
+sampling from the prior without incorporating any observed variables, or
+to perform inference, by conditioning on observed variables. After a
+brief discussion particle filtering, an inference algorithm often used
+to infer the values of the latent state variables, we will introduce our
+extension to the model that will allow a prediction of the consequences
+of active movements.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{diagrams/dynamic-belief-network-inference-1} \caption{Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.}\label{fig:graphical-model-inference}
+\centering
+\includegraphics{diagrams/dynamic-belief-network-inference-1.pdf}
+\caption{\label{fig:graphical-model-inference}Graphical model for inferring
+the 2-dimensional state \(x\), consisting of angular velocity and
+position, based on noisy sensory measurements \(y\).}
 \end{figure}
 
-\subsection{Particle filtering}\label{particle-filtering}
-
-We give a brief overview of particle filtering, concentrating on those
-aspects most relevant for this study. For a more thorough treatment, see
-Doucet et al. (2000) and Doucet \& Johansen (2009). Speekenbrink (2016)
-provides a recent review with applications in psychological research.
-Particle filtering is an algorithm for performing approximate Bayesian
-inference in a graphical model. It proceeds by recursively computing the
-distribution of the current state \(x_t\) as a a function of its parent
-nodes, consisting in our example of the state at the previous time step.
-
-The general procedure for recursive Bayesian inference is to perform
-alternating prediction and updating steps.
+\subsection{Bayesian inference using particle
+filtering}\label{bayesian-inference-using-particle-filtering}
+
+We give a brief overview of particle filtering, a sequential Monte Carlo
+algorithm for performing approximate Bayesian inference in a graphical
+model. We describe a simple bootstrap particle filter, concentrating on
+those aspects most relevant for this study. For a more thorough
+treatment, see Doucet et al. (2000) and Doucet and Johansen (2009).
+Speekenbrink (2016) provides a recent review with applications in
+psychological research. Particle filtering proceeds by recursively
+estimating the distribution of the current state, \(P(x_t | y_{1:t})\),
+as a function of its parent nodes, conditional upon the sequence of
+measurements \(y_{1:t}\). The distributions are represented by a
+collection of discrete weighted samples, known as particles.
+
+The general procedure for recursive Bayesian inference is to start from
+an initial prior distribution, and the to perform alternating prediction
+and updating steps. In other words: first, a hypothesis about the
+expected state is generated, and then this hypothesis is tested by
+evaluating how well it predicts the data. The algorithm starts with a
+prior distribution of the particles, and at each subsequent time step,
+proceeds by first sampling from the proposal distribution
+\(P(x_t | y_{1:t-1})\), which is based on the specific choice of
+transition function. Next, during the updating step, the current
+measurement is incorporated by evaluating the likelihood,
+\(P(y_t | x_t)\), which gives the probability of observing the
+measurement for each individual particle. The particle are then assigned
+an importance weight, based on how well they predicted the measurement.
+A common problem in sequential Monte Carlo methods is that the particle
+weights tend to become very small over time; this is known as the
+degeneracy problem. In order to counteract this, multinomial resampling
+is performed when effective sample size falls under a threshold. A new
+set of particles is then chosen by weighted sampling with replacement,
+which each particle's probability of being selected given by its
+importance weight. Resampling ensures that only particles with a
+relatively large probability of \enquote{explaining} the observations
+are maintained. This introduces flexibility in the algorithm's method
+for incorporating evidence into its beliefs about the latent states.
+
+Figure \ref{fig:particle-filter-explained} shows a visualization of the
+computations performed by a particle filter at each time step, in order
+to track the angular velocity. Initially, the particles are propagated
+forwards in time according to the proposal distribution. They are
+equally weighted, indicated by their diameters. During the update step,
+the particle are re-weighted, with each particle's weights reflecting
+how well it explains the measurement. This is visualized by the change
+in diameter; those particles lying close to the observation (sensory
+input from the horizontal canal) have a larger diameter. After
+resampling, the new set of particles is pulled towards the measurement,
+and the weights are reset. This adjustment toward the observation
+depends on both the process noise, and the modelled sensor noise. The
+uncertainty of the belief, as indicated by the particle cloud's standard
+deviation (light grey lines), increases at every prediction step, and
+subsequently decreases after updating. This cycle of prediction,
+updating and resampling ensures that the collection of particles comes
+to represent the \enquote{true} posterior distribution, by eliminating
+those particles with low weights. The particle filter represents all
+beliefs in the form of probability distributions; the mean and variance
+of the belief distribution can be computed at each time step, and the
+logarithm of the weights \(log(\frac{1}{N}\sum^N_{i=1}w_i)\), averaged
+over all \(N\) particles, may be summed over all time steps in order to
+compute the log-likelihood
+\(\sum^T_{t=1}\Big[log(\sum^N_{i=1}w_i) - log(N)\Big]\), which can be
+used in order to evaluate how well the model was able to fit the sensory
+evidence. The summed log-likelihood can be used for parameter estimation
+in extensions of the basic particle filtering algorithm (Andrieu,
+Doucet, \& Holenstein, 2010).
+
+\begin{figure}
+\centering
+\includegraphics{diagrams/figure-particle-filter-explained.pdf}
+\caption{\label{fig:particle-filter-explained}Particle filter computations
+performed at every time step. The cloud of particles (grey circles)
+represents the current belief state. Initially, message=FALSE,
+warning=FALSE, a prediction is made for each particle by sampling from
+the proposal distribution, which is determined by the agent's kinematic
+model. This is equivalent to sampling from the prior distribution. Then,
+during the updating step, an observation is obtained (shown as a black
+diamond). Each particle is assigned an importance weight, according to
+how well it explains the observation. Each particle's diameter reflects
+its importance weight. Subsequently, a new set of particles is formed by
+weighted multinomial resampling, with each particle's probability of
+being chosen proportional to its importance weight. The particle cloud's
+mean and standard deviation are shown as light grey circles and lines.
+This resulting cloud of particles represents the posterior distribution.
+This posterior subsequently forms the prior distribution for the next
+time step.}
+\end{figure}
+
+Figure \ref{fig:3d-statespace-fits} (a) shows the result of applying the
+particle filtering algorithm to the graphical model depicted in Figure
+\ref{fig:graphical-model-inference} and simulated measurements, in order
+to infer the velocity and position during a two-second leftward head
+turn. The acceleration is also inferred, but is not of interest.
+
+The measurements are shown as circles, and the true velocity that
+generated the data is shown as a dotted line. This velocity is the
+result of an external force being applied. The solid lines depict the
+mean estimates, and the shaded areas show the \(95\%\) credible
+intervals. The model is able to estimate the latent state despite making
+the assumption that the acceleration is constant. It does this by
+allowing for random changes in acceleration by letting the process noise
+be sufficiently large. Note that both position and velocity estimates
+lag the true values; this is due the fact that the acceleration is
+expected to be constant, and thus the system is slow to react to
+changes. The estimation can be improved by increasing the process noise.
+Figure \ref{fig:3d-statespace-fits} (b) shows the resulting estimates.
+Whilst the lag is reduced, a consequence of increasing the process noise
+is that the measurements are given more weight, resulting in a tendency
+to over-fit the noise in the measurements. The constant velocity (random
+acceleration) model is unable to capture the underlying smooth
+trajectory of the velocity and position, instead showing random jumps in
+the velocity estimate.
+
+\begin{figure}
+\centering
+\includegraphics{generated-figures/3d-statespace-fits-1.pdf}
+\caption{\label{fig:3d-statespace-fits}(a) Particle filter with a 3-d state
+space model for inferring velocity and position during a two-second head
+turn. (b) The same particle filter with increased process noise.}
+\end{figure}
+
+\section{Extension allowing the prediction of active
+movements}\label{extension-allowing-the-prediction-of-active-movements}
+
+During sensory inference the measurements are incorporated, and and a
+posterior estimate of the latent state is obtained by updating the prior
+with the likelihood. The underlying kinematic model thus relies on the
+sensory data to inform its estimates. This model can be used in order to
+track the behaviour of a dynamic system by including the acceleration as
+a latent state, it is purely data-driven, and does not include any means
+of incorporating an active prediction of the kinematics. Thus, it cannot
+explain top-down expectations or biases. This form of anticipation is
+not the same as the 1-step ahead predictions made during inference, but
+rather implements higher level prior beliefs about the direction of
+movement. For example, imagine seeing that someone is going to push you.
+You can use this information, derived from the visual modality, in order
+prepare yourself in order to compensate for this. However, you do not
+want to act too early, but rather wait until you are actually moving.
+
+\begin{quote}
+While this model is perfectly suitable for sensory inference, for the
+task of predicting the sensory consequences of self-initiated
+movemenents, the model is inadequate, as it does not allow including
+further knowledge about kinematics, which can be used to improve the
+proposal distribution. This knowledge is obtained by means of an
+efference copy of executed or planned motor commands, which is sent to a
+forward model. The forward model must then translate the information
+from the efference copy in a format that is suitable for the sensory
+system.
+\end{quote}
+
+An obvious way of including this is the mechanism that allows one to
+predict the sensory consequences of self-initiated movement via an
+efference copy. In state-space models, this is implemented as a control
+input \(u\), which extends the graphical model shown in Figure
+\ref{fig:graphical-model-inference}. In this model, the estimated
+acceleration is replaced by an known acceleration signal. This
+capability to predict the consequences of self-initiated movements is
+essential in order to distinguish re-afference from ex-afference - the
+system has knowledge about how it moves, and can use this to predict the
+sensory consequences. A further benefit of including a control input is
+that the model can now be used in order to run counterfactual
+simulations of the dynamic behaviour of the system for various types of
+movements.
+
+We implement the control in the following manner: during an active head
+movement, the model generates a sequence of control signals \(u_{1:T}\).
+For example, during a leftward head turn, using the sinusoidal
+acceleration profile shown in Figure \ref{fig:head-turn-data}, the
+acceleration is given by \(Asin(2\pi t/T)\). Assuming the model has full
+knowledge about the onset of movement, and the type of movement
+(sinusoidal head turn), the model requires information about the
+frequency, \(f = 1/T\), amplitude, \(A\), and direction, which is given
+by the sign of the amplitude.
+
+Figure \ref{fig:graphical-model-control} shows the resulting graphical
+model.
+
+\begin{figure}
+\centering
+\includegraphics{diagrams/dynamic-belief-network-inference-2.pdf}
+\caption{\label{fig:graphical-model-control}Graphical model extended with
+hierarchical model for self-initiated movements.}
+\end{figure}
+
+\section{\texorpdfstring{\textbf{BIS
+HIER.}}{BIS HIER.}}\label{bis-hier.}
+
+\section{TODO: Modelling questions}\label{todo-modelling-questions}
+
+Note the the vestibular senspry input may be the result of a
+superposition of multiple generative source; the afferent signals,
+measuring head velocity, may be caused by a multitude of processes
+(active head turn, unintentional body sway, perturbative external forcec
+(wind), etc.). The brain must treat these resulting head movements
+differently; active head turn may be goal-directed, with the aim of
+directing the gaze toward a target. In this case, no compensatory action
+os desired in order to counteract the movement. This would be
+counterproductive. However, in order to maintian postural stability, the
+brain must compensate for all movement that is not gol-directed,
+i.e.~does not occur as a result of \enquote{planned} active motion.
+These components may be separated either at the level of the outcome
+variable (sensory signal), or at the level of the latent state
+variables. The latter amounts ot a latent decomposition.
+
+\begin{quote}
+Estimated velocity due to active/passive movement must be trated
+differently. Sensory signals due to active motion (re-afference) can be
+viewed as being cause by known explanatory variables, or as intervention
+effects.
+\end{quote}
+
+\begin{quote}
+discuss ({\textbf{???}}) (artificial velcoity vector)
+\end{quote}
+
+\begin{itemize}
+\tightlist
+\item
+  Connecting higher cognition to sensory inference:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    imagery uses a similar generative model (copied from sensory model),
+    then what is it doing?
+  \item
+    \textbf{how does mental simulation work? When we propose a mental
+    simulation, what does this actually mean? When we propose that
+    simulation re-uses the circuit for sensory inference, which parts
+    are actually required? Do we need the data? Surely not. NO, we need
+    the generative model that is used for on-line inference, but we
+    should not perform inference with it. ON the contrary, we should
+    avoid inference at all costs, instead running a Monte Carlo
+    simulation with the generative model, i.e.~just evolving the
+    evolution of the dynamical system according to the transition prior
+    (sampling from the proposal distribution).}
+  \item
+    mental simulation allows us to answer \enquote{what if}
+    (counterfactual) queries in a generative model. Why should this
+    model be close to the sensory model? What if the senspry model
+    changes? The twin model needs updating.
+  \item
+    If mental simulation interferes with sensory inference: how?
+  \item
+    \textbf{If we want to model higher level cognition as being
+    connected to (grounded in) lower level sensory circuits, we can join
+    both in a hierarchical graphical model.}
+  \end{itemize}
+\item
+  Explain what tasks sensory inference must perform. Tasks the are
+  particular to the modality that senses movements of the body. It needs
+  to detect non-linearities, i.e.~non-stationary process. This means ist
+  needs to detect the sudden onset of movements, or sudden direction
+  changes. When these are passive and cannot be predicted , the best the
+  system can hope for is to detect the changes as quickly as possible.
+  When they are the result of active movement, they can be predicted.
+  But they can also be anticipated due to knowledge from other sources,
+  such as visual or verbal clues; if the knowledge is available, how
+  does the system make use of it? Proposal distribution use for
+  sequential importance sampling are notoriously difficult to design.
+  The approach of sampling from the prior can be very inefficient, as
+  when the process is not stationary, i.e.~in this case when there are
+  sudden changes in acceleration, the likelihood can change abruptly,
+  and the weights would become neglible for all particles. In certain
+  cases, the filter cannot recover from this, and would completely
+  misrepresent the data. We propose that the ability to anticipate must
+  result in the construction (flexible) of a proposal distribution, and
+  that interference between mental simulation and inference involves the
+  proposal distribution.
+\end{itemize}
+
+\begin{quote}
+Control input for predictions is used to \enquote{design} better
+proposal distributions for inference.
+\end{quote}
+
+\begin{itemize}
+\tightlist
+\item
+  Experimental results to be explained: Wertheim et al. (2001) report a
+  cognitive suppression of otolith responses, based on participants'
+  prior beliefs about possible motions.
+\end{itemize}
+
+\subsection{TODO: Contrast model with Merfeld
+(2016)}\label{todo-contrast-model-with-merfeld-2016}
+
+Merfeld's model claims: filtering is performed at a low level in order
+to extract velocity (maybe peak velocity?) and then this estimate is
+processed sub-optimally by low-pass filtering. The decision variable is
+computed at a higher level and separate from the model that is invloved
+in inference.
+
+We propose that there is an alternaitve way of looking at this. What if
+there a second system that \enquote{interprets} the data when sensory
+inference using the 3-d state space representation alone is insufficent.
+This model \enquote{steps in} when the uncertainty is high, and it is
+this system that enables anticpation of non-stationarities, i.e.~onset
+of sudden changes. This is derived from the system that allows us to
+predict consequences of own actions through the use of a control input
+obtained via a forward model from the cerebellum. This system would
+traditionally be viewed as depending on the motor commands from commands
+that are actually performed. Maybe this is more general, allowing the
+brain to include an exogenous variable in order to predict as a control
+input, i.e.~this can also include \enquote{covert} actions. This
+flexible control of the exogenous input is actually necessary for
+planning, etc. Flexible control of input means: the ability to clamp
+nodes, i.e.~to fix nodes at certain values. Without the ability to clamp
+higher-level node, the inference model can only be used inflexibly for
+inference. The whole idea is based on the fact that higher level
+knowledge and beliefs can influence our perception, and this is
+especially true for vestibular signals, which are highly ambiguous, and
+require prior knowledge for disambiguation and interpretation. Actually,
+we should implement something similar to (Hamrick et al., n.d.): a
+metacontroller, which chooses either the low-level module, or when this
+is unable to deal with the problem, a higher level module. This higher
+level module introduces additional information about the problem,
+e.g.~information obtained by verbal instructions, or a bias. FOr
+example, the instructions: \enquote{choose right when uncertain} should
+lead participants to implement use this strategy. In other words, when
+you are certain, base you choice on your low level sensory inference,
+but when uncertain use other information, such as task demands. Surely
+we can design a better experiment than this, though?
+
+Implement the above ideas in a figure, showing the two approaches, and
+differences, especially for predictions about future data.
+
+\subsection{TODO: Simulations}\label{todo-simulations}
+
+Show that when fit to the same data, a model implementing a control
+provides substantially better fit, as measured by log-likelihood, when
+compared to 3-d state space model (basic model) {[}think of better names
+for models{]}. This is almost trivial; but show it like this: under the
+assumption that the agent knows the precise timing (onset) of motion,
+and the functional form (sinusiodal acceleration), all the agent needs
+to assume is: \{A, direction, frequency\}. Show that performance is
+substantially better if agent has priors for A, direction (maybe
+ignoring f, or vice versa: ignore A, and put priors on f, might be
+better). Then draw A (or f) from a distribution, and run inference using
+control model. Sampling A from prior distribution means: the agent has
+imperfect knowledge about its amplitude of motion.
+
+This improvement in inference is only better when the proposal
+distribution is well-specified; for a mis-specified model, the
+performance can be dramatically worse.
+
+We claim: interference effects arise when participant's wrongly use a
+control (predictive model or \textbf{anticipation model}) using motion
+paramters used for mental simulation during subsequent inference. This
+use of higher level priors has the same (similar) effect as
+anticipation.
+
+\section{An investigation of various priors and
+predictions}\label{an-investigation-of-various-priors-and-predictions}
+
+Here, we identify the various forms of predictive processing in the
+context of real-time sensory inference using SMC
+
+\begin{itemize}
+\tightlist
+\item
+  Related to filtering algorithm
+\item
+  Higher-level priors: The process model determines the kind of motion
+  that is expected. Call this \emph{hypothesis-driven} sensory
+  inference. It implements expectation on longer time-scales.
+\end{itemize}
+
+Show a table demonstrating the difference between:
+
+\begin{itemize}
+\tightlist
+\item
+  prediction: 1-step ahead during inference
+\item
+  expectation: higher level priors
+\end{itemize}
+
+The brain implements multiple priors:
+
+\begin{itemize}
+\tightlist
+\item
+  ability to select type of motion
+\item
+  parameters of the motion
+\end{itemize}
+
+\section{Various sources of
+interference}\label{various-sources-of-interference}
+
+Interference is interesting in its own right. It is important to
+understand how and why this should occur. Furthermore, we propose that
+gaining a better understanding of this will allow us to better
+understand the nature of \enquote{thinking} as mental simulation.
+
+Interference is rather surprising; if the goal is to perform simulation
+without incorporating the data, then we should run a Monte Carlo
+simulation. Why should this interfere with sensory inference perform by
+a particle filter? We propose that the answer is in the shared
+generative model.
+
+Here, we identify various sources of interference (2-way) between
+thought (higher-level cognition) and lower-level sensory processing.
+This interference can be either due to low-level influences in sensory
+inference, due to priors on higher-level parameters, choice of process
+model (false expectations of certain type of motion). This is shown by
+model simulations.
+
+\begin{itemize}
+\item
+  Model mis-specification: If an observer uses a strong higher level
+  prior, i.e.~the wrong direction, this means that the data may be
+  completely mis-interpreted. (call this: the effect of inappropriate
+  priors (Schwabe \& Blanke, 2008))
+\item
+  Re-use of same generative model: This may lead to leakage,
+  i.e.~inadvertently incorporating sensory evidence
+\item
+  Decision-making: The interference may be explained in terms of
+  higher-level perceptual decision-making mechanisms
+
+  \begin{itemize}
+  \tightlist
+  \item
+    evidence accumulation
+  \item
+    bias (Starting point)
+  \end{itemize}
+\end{itemize}
+
+\subsection{Summary}\label{summary}
+
+Interference may the result of either:
 
@@
 \def\labelenumi{\arabic{enumi})}
-\item
-  Predict step: At time t, the current state is predicted according to
-  the process model. This prediction is maintained in the form of
-  probability distributions over all the values.
-\item
-  Update step: a measurement is obtained, and this measurement is used
-  in order to update the state estimate.
+\tightlist
+\item
+  mis-specified proposal distribution: inappropriate priors, or wrong
+  type of motion or
+\item
+  biased decision-making
 \end{enumerate}
 
-By recursively estimating the current state, the observations can be
-processed continuously; thus, the agent does not need to store the
-entire history of observations -- only the previous belief state
-estimate and the current observation are needed in order to estimate the
-current state. At time step t, the agent observes the value of , and
-this observation is used to update the previous belief:
-
-\autoref{fig:particle-filter-explained} shows a diagram \ldots{}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{diagrams/figure-particle-filter-explained} \caption{Particle filter computations performed at every time step.
-The cloud of particles (grey circles) represents the current belief state.
-Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model.
-This is equivalent to sampling from the prior distribution.
-Then, during the updating step, an observation  is obtained (shown as a black diamond).
-Each particle is assigned an importance weight, according to how well it explains the observation.
-Each particle's diameter reflects its importance weight.
-Subsequently, a new set of particles is formed by weighted multinomial resampling, with each particle's probability of being chosen proportional to its importance weight.
-The particle cloud's mean and standard deviation are shown as light grey circles and lines.
-This resulting cloud of particles represents the posterior distribution.
-This posterior subsequently forms the prior distribution for the next time step.}\label{fig:particle-filter-explained}
-\end{figure}
-
-\autoref{fig:passive} shows the estimated velocity \ldots{}
-
-\begin{no-prefix-figure-caption}
-
-\begin{figure}
-\centering
-\includegraphics{../generated-figures/passive-1.pdf}
-\caption{\label{fig:passive}Particle filter applied\ldots{}}
-\end{figure}
-
-\end{no-prefix-figure-caption}
-
-\subsection{Summary}\label{summary}
-
-\section{Extend previous models with control
-input}\label{extend-previous-models-with-control-input}
-
-For sensory inference of a passive disturbance, the algorithm can be run
-with fixed parameters. Here we extend the basic particle filtering
-(algorithm) model, in order to:
+We should be able to investigate which by applying cognitive process
+models that explain response times and choices.
+
+\section{Simulations}\label{simulations}
+
+\begin{quote}
+Should we have a separate simulations section, or should the simulations
+be distributed? \#\# Model specifications
+\end{quote}
+
+\subsection{Monte Carlo simulation}\label{monte-carlo-simulation}
+
+Show dynamics when simulating using various proposal distributions
+implementing various natural head motions:
 
@@
 \item
-  \textbf{provide the capability to model the sensory consequences of
-  self-initiated actions} (i.e.~predict/infer the consequences of one's
-  own actions)
-\item
-  provide the capability to run a Monte Carlo simulation using the same
-  generative model as is used for sensory inference
+  turning left
+\item
+  shaking head quickly
+\item
+  shaking head slowly (communicative gesture)
 \end{itemize}
 
-\subsection{Generative Model}\label{generative-model}
-
-The generative model is a state-space model, which itself is a dynamic
-model consisting of latent (state variables) and observed (measured)
-variables.
-
-\subsection{Inference Algorithm}\label{inference-algorithm}
-
-Sequential Monte Carlo
-
-\section{Modelling questions}\label{modelling-questions}
-
-Experimental results to be explained: Wertheim, Mesland, \& Bles (2001)
-report a cognitive suppression of otolith responses, based on
-participants' prior beliefs about possible motions.
-
-\section{An Investigation of Various
-Priors}\label{an-investigation-of-various-priors}
-
-Here, we identify the various forms of predictive processing in the
-context of real-time sensory inference using SMC \#\# Related to
-Filtering Algorithm \#\# Higher-level Priors control input
--\textgreater{} infer the consequences of one's own actions.
-
-\section{Various sources of
-interference}\label{various-sources-of-interference}
-
-Here, we identify various sources of intereference (2-way) between
-thought (higher-level cognition) and lower-level sensory processing
+\subsection{Expectations based on higher level
+priors}\label{expectations-based-on-higher-level-priors}
+
+In our model, there is no difference between modelling active
+self-motion, and an anticipation of passive self-motion due to
+predictable external forces. This is a testable hypothesis.
 
@@
 
-If an observer uses a strong higher level prior, i.e.~the wrong
-direction, this means that the data may be completely mis-interpreted.
-(call this: the effect of inappropriate priors (Schwabe \& Blanke,
-2008))
-
-\subsection{Re-use of same generative
-model}\label{re-use-of-same-generative-model}
-
-This may lead to leakage, i.e.~inadvertently incorporating sensory
-evidence
-
-\subsection{Decision-making}\label{decision-making}
-
-The interference may be explained in terms of higher-level perceptual
-decision -making mechanisms \#\# evidence accumulation \#\# bias
-(Starting point)
-
-\subsection{Summary}\label{summary-1}
-
-Interference may the result of either inappropriate priors, leading to a
-mis-specified filtering algorithm, or biased decision-making. We should
-be able to investigate which by applying cognitive process models that
-explain response times and choices.
-
-\section{Simulations}\label{simulations}
+\subsubsection{Inappropriate priors}\label{inappropriate-priors}
+
+\subsubsection{Wrong type of motion
+model}\label{wrong-type-of-motion-model}
+
+For example, expect shaking instead of head turn
 
@@
 
-The main result of this study is - place the emerging field of
-vestibular cognition with a framewotk for dynamic probabilistic
-inference and draw links to established models of sensory processing -
-interpret result from vestibular cognition studies showing interference
-between thought and perception as being the result of biased sensory
-processing
+The main results of this study are:
+
+\begin{itemize}
+\item
+  place the emerging field of vestibular cognition with a framework for
+  dynamic probabilistic inference and draw links to established models
+  of sensory processing
+\item
+  interpret result from vestibular cognition studies showing
+  interference between thought and perception as being the result of
+  biased sensory processing. This interference can be either due to
+  low-level influences in sensory inference, due to priors on
+  higher-level parameters, choice of process model (false expectations
+  of certain type of motion). This is shown by model simulations.
+\item
+  A further possibility for interference is a bias introduced in the
+  decision-making process (discuss potential model of decision-making
+  based on sequential Bayes factor and contrast with (Merfeld, Clark,
+  Lu, \& Karmali, 2016).)
+\end{itemize}
 
@@
 
-\subsection{Support for our model:
-Nigmatullina}\label{support-for-our-model-nigmatullina}
+\subsection{What is the difference between predicting and
+imagining?}\label{what-is-the-difference-between-predicting-and-imagining}
+
+Imagining can be described as counterfactual inference (Krishnan,
+Shalit, \& Sontag, 2015). Imagined motion (for planning,
+decision-making, perspective transformations) is performed under
+counterfactual assumptions. Predictions/anticipation is performed under
+factual assumptions.
+
+\subsection{Support for our model:}\label{support-for-our-model}
+
+\begin{itemize}
+\tightlist
+\item
+  Nigmatullina
+\item
+  Wertheim (Grabherr, Cuffel, Guyot, \& Mast, 2011) showed that patients
+  with bilateral or unilateral vestibular loss show impaired ability to
+  mentally transform images of bodies and body parts.
+\end{itemize}
 
@@
 
-Why dynamic simulation? Maybe in order to provide vestibular in put to
-head direction cells (Taube, 2007)
+Head direction cells are thought to integrate signals of vestibular
+origin to maintain a signal of cumulative rotation (reviewed in Taube
+(2007) and are found in many brain areas, including the postsubiculum,
+retrosplenial cortex, thalamus, lateral mammillary nucleus, dorsal
+tegmental nucleus, striatum and entorhinal cortex (Cullen, 2014). This
+may be necessary in order to mentally simulate being in another
+position, i.e.~the brain apparently needs to simulate the n-step
+dynamics of its motion through space. Why is this? Because vestibular
+connections make a major contribution to spatial orientation;
+information about the position in space obtained through the vestibular
+sensory system is computed by integrating velocity over time.
+
+\begin{itemize}
+\item
+  mental simulation cannot just re-use the state used in sensory
+  inference, because the velocity is fed to the head direction cells.
+  The brain would get confused
+\item
+  directional signal is not just a simple integration of velocity
+  signals from the vestibular system
+\item
+  signals from motor efferecence copy or proprioceptive systems appear
+  to be crucial for updating the preferred firing direction over time
+\item
+  Forward model for computing anticpation-related control input may
+  reside in the cerebellum. THe cerebellum may be involved in imagined
+  movemeent, or is it just for actions that are actually performed?
+\end{itemize}
+
+\subsection{Testable hypotheses}\label{testable-hypotheses}
+
+Show a table of testable hypotheses
 
@@
 
-\section{Motion Data}\label{motion-data}
-
-\section{Tables}\label{tables}
-
-\begin{longtable}[]{@{}rlcl@{}}
-\caption{Demonstration of simple table syntax.}\tabularnewline
-\toprule
-Right & Left & Center & Default\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Right & Left & Center & Default\tabularnewline
-\midrule
-\endhead
-12 & 12 & 12 & 12\tabularnewline
-123 & 123 & 123 & 123\tabularnewline
-1 & 1 & 1 & 1\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}clrl@{}}
-\caption{Here's the caption. It, too, may span multiple
-lines.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.15\columnwidth}\centering\strut
-Centered Header\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Default Aligned\strut
-\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\raggedleft\strut
-Right Aligned\strut
-\end{minipage} & \begin{minipage}[b]{0.31\columnwidth}\raggedright\strut
-Left Aligned\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.15\columnwidth}\centering\strut
-Centered Header\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Default Aligned\strut
-\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\raggedleft\strut
-Right Aligned\strut
-\end{minipage} & \begin{minipage}[b]{0.31\columnwidth}\raggedright\strut
-Left Aligned\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.15\columnwidth}\centering\strut
-First\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-row\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedleft\strut
-12.0\strut
-\end{minipage} & \begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
-Example of a row that spans multiple lines.\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.15\columnwidth}\centering\strut
-Second\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-row\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedleft\strut
-5.0\strut
-\end{minipage} & \begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
-Here's another one. Note the blank line between rows.\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
 \newpage
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Andrieu:2010gc}{}
+Andrieu, C., Doucet, A., \& Holenstein, R. (2010). Particle Markov chain
+Monte Carlo methods. \emph{Journal of the Royal Statistical Society:
+Series B (Statistical Methodology)}, \emph{72}(3), 269--342.
+
+\hypertarget{ref-Baker:2009kx}{}
+Baker, C. L., Saxe, R., \& Tenenbaum, J. B. (2009). Action understanding
+as inverse planning. \emph{Cognition}, \emph{113}(3), 329--349.
+
+\hypertarget{ref-Battaglia:2013fm}{}
+Battaglia, P. W., Hamrick, J. B., \& Tenenbaum, J. B. (2013). Simulation
+as an engine of physical scene understanding. \emph{Proceedings of the
+National Academy of Sciences of the United States of America},
+\emph{110}(45), 18327--18332.
+
 \hypertarget{ref-Beveridge:2013gx}{}
@@
 
-\hypertarget{ref-Blei:2014cp}{}
-Blei, D. M. (2014). Build, Compute, Critique, Repeat: Data Analysis with
-Latent Variable Models. \emph{Annual Review of Statistics and Its
-Application}, \emph{1}(1), 203--232.
-
 \hypertarget{ref-Borah:1988fd}{}
@@
 Botvinick, M. M., \& Toussaint, M. (2012). Planning as inference.
-\emph{Trends In Cognitive Sciences}, \emph{16}(10), 485--488.
+\emph{Trends in Cognitive Sciences}, \emph{16}(10), 485--488.
 
@@
 
+\hypertarget{ref-Ellis:2017fg}{}
+Ellis, A. W., \& Mast, F. W. (2017). Toward a Dynamic Probabilistic
+Model for Vestibular Cognition. \emph{Frontiers in Psychology},
+\emph{8}.
+
 \hypertarget{ref-zuEulenburg:2012bh}{}
@@
 
+\hypertarget{ref-Goodman:2016go}{}
+Goodman, N. D., \& Frank, M. C. (2016). Pragmatic Language
+Interpretation as Probabilistic Inference. \emph{Trends in Cognitive
+Sciences}, \emph{20}(11), 818--829.
+
+\hypertarget{ref-Grabherr:2011ej}{}
+Grabherr, L., Cuffel, C., Guyot, J.-P., \& Mast, F. W. (2011). Mental
+transformation abilities in patients with unilateral and bilateral
+vestibular loss. \emph{Experimental Brain Research}, \emph{209}(2),
+205--214.
+
 \hypertarget{ref-Grabherr:2008kk}{}
@@
 
+\hypertarget{ref-Griffiths:2012ii}{}
+Griffiths, T. L., Vul, E., \& Sanborn, A. N. (2012). Bridging levels of
+analysis for probabilistic models of cognition. \emph{Current Directions
+in Psychological Science}.
+
 \hypertarget{ref-Grush:2004wx}{}
@@
 \emph{27}(3), 377--96--discussion396--442.
+
+\hypertarget{ref-Hamrick:2014wq}{}
+Hamrick, J. B., \& Griffiths, T. L. (2014). What to simulate? Inferring
+the right direction for mental rotation. \emph{CogSci}.
+
+\hypertarget{ref-Hamrick:8oFH5gP_}{}
+Hamrick, J. B., Ballard, A. J., Pascanu, R., Vinyals, O., Hees, N., \&
+Battaglia, P. W. (n.d.). Metacontrol For Adaptive Imagination-Based
+Optimization. Retrieved from
+\url{https://openreview.net/pdf?id=Bk8BvDqex}
 
@@
 
+\hypertarget{ref-Kemp:2007eo}{}
+Kemp, C., Perfors, A., \& Tenenbaum, J. B. (2007). Learning
+overhypotheses with hierarchical Bayesian models. \emph{Developmental
+Science}, \emph{10}(3), 307--321.
+
 \hypertarget{ref-Klingner:2016ia}{}
@@
 
+\hypertarget{ref-Koller:2009ty}{}
+Koller, D., \& Friedman, N. (2009). \emph{Probabilistic Graphical
+Models}. MIT Press.
+
+\hypertarget{ref-Krishnan:2015wr}{}
+Krishnan, R. G., Shalit, U., \& Sontag, D. (2015). Deep Kalman Filters.
+Retrieved from \url{http://arxiv.org/abs/1511.05121}
+
+\hypertarget{ref-Kwisthout:2016fx}{}
+Kwisthout, J., Bekkering, H., \& Rooij, I. van. (2016). To be precise,
+the details don't matter: On predictive processing, precision, and level
+of detail of predictions. \emph{Brain and Cognition}.
+
 \hypertarget{ref-Laurens:2006be}{}
@@
 
+\hypertarget{ref-Lucas:2010tu}{}
+Lucas, C. G., \& Griffiths, T. L. (2010). Learning the form of causal
+relationships using hierarchical Bayesian models. \emph{Cognitive
+Science}.
+
 \hypertarget{ref-PaulRMacNeilage:2008bv}{}
@@
 
+\hypertarget{ref-Merfeld:2011fs}{}
+Merfeld, D. (2011). Signal detection theory and vestibular thresholds:
+I. Basic theory and practical considerations. \emph{Experimental Brain
+Research. Experimentelle Hirnforschung. Expérimentation cérébrale},
+\emph{210}(3-4), 389--405.
+
+\hypertarget{ref-Merfeld:2016il}{}
+Merfeld, D., Clark, T. K., Lu, Y. M., \& Karmali, F. (2016). Dynamics of
+individual perceptual decisions. \emph{Journal of Neurophysiology},
+\emph{115}(1), 39--59.
+
 \hypertarget{ref-Merfeld:1999cg}{}
@@
 
+\hypertarget{ref-Nigmatullina:2015gh}{}
+Nigmatullina, Y., Arshad, Q., Wu, K., Seemungal, B. M., Bronstein, A.
+M., \& Soto, D. (2015). How imagery changes self-motion perception.
+\emph{Neuroscience}, \emph{291}, 46--52.
+
+\hypertarget{ref-Pearl:1988wz}{}
+Pearl, J. (1988). \emph{Probabilistic Reasoning in Intelligent Systems}.
+Morgan Kaufmann.
+
 \hypertarget{ref-Penny:2013iv}{}
@@
 \emph{9}(12), e1003383.
+
+\hypertarget{ref-Perfors:2011ex}{}
+Perfors, A., Tenenbaum, J. B., Griffiths, T. L., \& Xu, F. (2011). A
+tutorial introduction to Bayesian models of cognitive development.
+\emph{Cognition}, \emph{120}(3), 302--321.
 
@@
 Pickering, M. J., \& Clark, A. (2014). Getting ahead: forward models and
-their place in cognitive architecture. \emph{Trends In Cognitive
+their place in cognitive architecture. \emph{Trends in Cognitive
 Sciences}, \emph{18}(9), 451--456.
@@
 Raphan, T., Matsuo, V., \& Cohen, B. (1979). Velocity storage in the
-vestibulo-ocular reflex arc (VOR). \emph{Experimental Brain Research}.
+vestibulo-ocular reflex arc (VOR). \emph{Experimental Brain Research},
+\emph{35}(2).
 
@@
 
+\hypertarget{ref-Tenenbaum:2011hd}{}
+Tenenbaum, J. B., Kemp, C., Griffiths, T. L., \& Goodman, N. D. (2011).
+How to grow a mind: statistics, structure, and abstraction.
+\emph{Science}, \emph{331}(6022), 1279--1285.
+
 \hypertarget{ref-Tian:2012ui}{}

### 2017-03-24 simulations/01-3d-statespace.R

@@ -77,7 +77,7 @@ data {
   for (t in 2:t_max) {
     # p: position
     # w: velocity
-    a_true[t-1] <- A * sin(two_pi*(t-1) * freq) #+ A * sin(two_pi*(t-1) * 4 * freq)
+    a_true[t-1] <- A * sin(two_pi*(t-1) * freq) + A * sin(two_pi*(t-1) * 4 * freq)
     p_true[t] ~ dnorm(p_true[t-1] + dt * w_true[t-1] + dt^2/2 * a_true[t-1], prec_p)
     w_true[t] ~ dnorm(w_true[t-1] + dt * a_true[t-1], prec_w)
 
@@ -96,7 +96,7 @@ model {
     # p: position
     # w: velocity
     # a[t-1] <- A * sin(two_pi*(t-1)/t_max)
-    p[t] ~ dnorm(p[t-1] + dt * w[t-1] + dt^2/2 * A * a[t-1], prec_p)
+    p[t] ~ dnorm(p[t-1] + dt * w[t-1] + dt^2/2 * 1 * a[t-1], prec_p)
     w[t] ~ dnorm(w[t-1] + dt * A * a[t-1], prec_w)
     a[t] ~ dnorm(a[t-1], prec_a)
 
@@ -317,5 +317,5 @@ p4 <- df %>%
 p4
 
 p5 <- cowplot::plot_grid(p3, p4)
-
-cowplot::plot_grid(p5, p5, ncol = 1)
+p5
+# cowplot::plot_grid(p5, p5, ncol = 1)

### 2017-03-24 R/playground.R
added particle filter algorithm diagram

@@ -1,0 +1,50 @@
+dt <- 0.1
+motiondata <-  generate_data(T = 2, frequency = 0.5,
+                             amplitude = 20,
+                             sensor_sd = 2.0)
+# plot_trajectories(motiondata = motiondata)
+
+motiondata_hidden <- motiondata %>%
+    gather(type, value = value, -time) %>%
+    filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration")))
+motiondata_hidden <- motiondata %>%
+    gather(type, value = value, -time) %>%
+    # filter(type != "observations") %>%
+    mutate(type = factor(type,
+                         levels = c("position",
+                                    "velocity",
+                                    "acceleration",
+                                    "observations")))
+# df_tmp <- data_frame(time = motiondata$time,
+#                      observations = motiondata$observations)
+# df_tmp$param <- as.factor("Measurements")
+
+p1 <- motiondata_hidden %>%
+    filter(type != "observations") %>%
+    ggplot(aes(x = time, y = value)) +
+    geom_line(aes(linetype = type), size = 1.4) +
+    geom_point(data = filter(motiondata_hidden, type == "observations"),
+               aes(x = time, y = value),
+               shape = 21, colour = "white",
+               fill = "white", size = 6, stroke = 2) +
+    geom_point(data = filter(motiondata_hidden, type == "observations"),
+               aes(x = time, y = value),
+               colour = "black",
+               fill = "white", size = 4, stroke = 2) +
+    # scale_shape_manual(name = element_blank(),
+    #                    labels = "Measurements",
+    #                    values = 21) +
+    guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
+    xlab("Time (sec)") +
+    ylab("Angular position (deg)") +
+    facet_wrap(~ type, ncol = 1) +
+    theme_apa() +
+    theme(legend.title=element_blank(),
+          panel.border = element_blank(),
+          axis.line = element_line(colour="black",
+                                   size = 0.75, lineend="square"))
+p1
